### 2024-06-03 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@VSPEC-collab}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2024-06-03 src/tex/ms.tex
Copy intro from overleaf

@@
 \usepackage{showyourwork}
+
+\newcommand{\teff}{$T_{\rm eff}$}
+\definecolor{tedcommentcolor}{HTML}{e17701}
+\newcommand{\TJ}[1]{\textcolor{tedcommentcolor}{#1}}
+
+\shorttitle{\sc VSPEC}
+\shortauthors{Johnson et al.}
 
@@
 % Title
-\title{An open source scientific article}
+\title{The {\sc VSPEC} code: Spectroscopic phase curves of 3D exoplanet models in the presence of stellar variability}
 
-% Author list
-\author{@VSPEC-collab}
+\author{Ted Johnson}
+\affiliation{NASA Goddard Space Flight Center \\
+8800 Greenbelt Rd \\
+Greenbelt, MD 20771, USA}
+\affiliation{Nevada Center for Astrophysics, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
+\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
+% Also CRESST, SURA, SEEC!
+
+\author{Cameron Kelahan}
+\affiliation{NASA Goddard Space Flight Center \\
+8800 Greenbelt Rd \\
+Greenbelt, MD 20771, USA}
+
+\author{Avi M. Mandell}
+\affiliation{NASA Goddard Space Flight Center \\
+8800 Greenbelt Rd \\
+Greenbelt, MD 20771, USA}
+
+\author{Tom Barclay}
+\affiliation{NASA Goddard Space Flight Center \\
+8800 Greenbelt Rd \\
+Greenbelt, MD 20771, USA}
+
+\author{Veselin B. Kostov}
+\affiliation{NASA Goddard Space Flight Center \\
+8800 Greenbelt Rd \\
+Greenbelt, MD 20771, USA}
+
+\author{Geronimo L. Villanueva}
+\affiliation{NASA Goddard Space Flight Center \\
+8800 Greenbelt Rd \\
+Greenbelt, MD 20771, USA}
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+    We present the Variable Star Phase Curve (\href{https://github.com/VSPEC-collab/VSPEC}{ \sc VSPEC}) code,
+    a python package to simulate spectroscopic observations of rocky exoplanets in the presence of stellar variability.
+    {\sc VSPEC} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG GlobES) application along with a custom-built
+    stellar model to produce spectroscopic light curves of the planet-host system. We also describe associated codes. \TJ{These codes should be named here and described a little}
 \end{abstract}
+
+\keywords{Exoplanet, M dwarf}
 
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+In the era of high-sensitivity exoplanet characterization missions such as the James Webb Space Telescope (JWST)
+and the Atmospheric Remote-sensing Infrared Exoplanet Large-survey (ARIEL), spectral analysis of exoplanet atmospheres
+is increasingly sensitive to contamination due to stellar inhomogenies (spots, granulation, etc.).
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+Recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between
+atmospheric absorption by water and water-rich spots on the stellar surface. This ``transit light source effect''
+\citep[TLS,][see also \citet{apai2018,barclay2021,garcia2022,barclay2023}]{rackham2018}
+occurs when the occulted region of the stellar surface is not representitive of the disk-integrated spectrum
 
-\bibliography{bib}
+Similarly, the Mid-IR Exoplanet CLimate Explorer \citep[MIRECLE,][]{mandell2022} mission concept
+will employ the Planetary Infrared Excess \citep[PIE][]{stevenson2020} technique to extract the planetary
+contribution from combined-light observations. Uncertainties on the stellar spectrum will dominate the analysis if it is not removed appropriately.
+
+To adequately prepare for future observations and future missions it will be necessary to demonstrate a method to mitigate these effects.
+This task requires a flexible tool that combines models of exoplanet atmospheres and stellar variability in a robust way.
+
+We present {\sc VSPEC}: Variable Star PhasE Curve\footnote{\url{https://github.com/VSPEC-collab/VSPEC}},
+an open-source Python 3 package to simulate observations of exoplanet systems with variable host stars.
+{\sc VSPEC} combines NASA's Planetary Spectrum Generator \citep[PSG,][]{villanueva2018} with a grid of
+custom PHOENIX models \citep{husser2013} to produce multiwavelength phase curves of the system, accounting
+for a 3D planetary atmosphere, transits and eclipses, and multiple sources of stellar variability.
+
+In this paper we will describe the science behind {\sc VSPEC} and demonstrate examples of its use.
+Section \ref{sec:star} describes the stellar model and the available sources of variability.
+Section \ref{sec:pl_model} discusses modeling the planetary atmosphere, including PSG/GlobES and the
+built-in GCM. In Section \ref{sec:examples} we provide examples of {\sc VSPEC} use cases and demonstrate
+its value to the exoplanet community. Finally, in Section \ref{sec:conclusion} we will discuss the future of the code and issues it might have.
+
+\bibliography{syw,VSPEC}
 

### 2024-06-03 src/tex/ms.tex
Coppied through section 2.0

@@
 
+{Stellar Model \label{sec:star}}
+
+The {\sc VSPEC} star model is designed modularly to allow
+for both simple and complex behaviors. Currently, it
+is represented by a rectangular grid of points on the stellar surface,
+each assigned an effective temperature. At any given time, the model computes
+the surface coverage fractions of each temperature visible to the observer, accounting for the
+spherical geometry, limb darkening, and any occultation by a transiting planet.
+
+The attributes of the \texttt{Star} class describe the bulk properties of the star, including radius,
+period, and the effective temperature of quiet photosphere. Herein we refer to this temperature as the
+photosphere temperature to differentiate it from the temperature of spots, faculae, or other sources of variability.
+
+\begin{figure}
+    \centering
+        \includegraphics[width=0.5\textwidth]{figures/surface_map_and_lc.pdf}
+    \script{surface_map_and_lc.py}
+    \caption{{\bf Top}: The surface map of a star with spots and faculae during a transit.
+    The large circles are spots. Faculae display their 3D structure, as the hot wall is more prevalent near the limbs.
+    {\bf Bottom}: An example lightcurve of a spotted star.}
+    \label{fig:surface_map}
+\end{figure}
+
 \bibliography{syw,VSPEC}

### 2024-06-03 src/scripts/surface_map_and_lc.py

@@ -1,0 +1,74 @@
+
+
+
+
+from astropy import units as u
+import numpy as np
+import matplotlib.pyplot as plt
+from pathlib import Path
+
+from vspec_vsm.spots import SpotCollection, StarSpot
+from vspec_vsm.faculae import FaculaCollection, Facula
+from vspec_vsm.granules import Granulation
+from vspec_vsm.flares import FlareGenerator
+from vspec_vsm.star import Star
+from vspec_vsm.config import MSH
+
+import pypsg
+
+pypsg.docker.set_url_and_run()
+
+from VSPEC import ObservationModel,PhaseAnalyzer
+
+from cartopy import crs as ccrs
+
+import paths
+
+outfile = paths.figures / 'surface_map_and_lc.pdf'
+
+fig = plt.figure(figsize=(4,6))
+ax1 = fig.add_subplot(2,1,1,projection=ccrs.Orthographic(0,90))
+ax2 = fig.add_subplot(2,1,2)
+
+
+CFG_PATH = Path(__file__).parent / 'spot_lightcurve.yaml'
+
+model = ObservationModel.from_yaml(CFG_PATH)
+model.build_star()
+model.warm_up_star(facula_warmup_time=10*u.hr)
+
+
+
+lon0 = 0*u.deg
+lat0 = 0*u.deg
+pl_radius = 1*u.R_earth
+pl_orbit = 0.05*u.AU
+inclination = 89.8*u.deg
+phase = 180.4*u.deg
+
+model.star.plot_surface(lat0, lon0, ax1, pl_orbit, pl_radius, phase, inclination)
+
+# LC
+
+
+model.build_planet()
+model.build_spectra()
+
+data = PhaseAnalyzer(model.directories['all_model'])
+
+wl_pixels = [0,300,500,700]
+time = data.time.to(u.day)
+for i in wl_pixels:
+    wl = data.wavelength[i]
+    lc = data.lightcurve(
+        source='star',
+        pixel=i,
+        normalize=0
+    )
+    ax2.plot(time,lc,label=f'{wl:.1f}')
+ax2.legend()
+ax2.set_xlabel(f'time ({time.unit})')
+ax2.set_ylabel('Flux (normalized)')
+
+fig.tight_layout()
+fig.savefig(outfile)

### 2024-06-03 src/tex/ms.tex
Copied through section 2.3 (faculae)

@@
 
+\subsection{Stellar Spectra \label{subsec:spectra}}
+Once the surface coverage is calculated, a composite spectrum is computed from a grid of PHOENIX stellar spectra
+\citep{husser2013} As of \texttt{VSPEC 0.1}, we have spectra between 2300 K and 3900 K, with steps of
+100 K. Each spectrum has $\log{g} = 5$ and solar metalicity.
+The raw spectra span from 0.1 to 20 \micron~ with $5 \times 10^{-6}$ \micron~ steps, however binned
+versions are pre-computed for faster runtimes.
+
+\subsection{Spots \label{subsec:spots}}
+Our star spot model is nearly entirely based on observations of the sun. Sun spots can be resolved and are well-studied,
+whereas spots on other stars (e.g., M dwarfs) can only be observed indirectly. We therefore designed our spot model to mimic
+sun spots but with parameterized values for spot temperature and lifetime that can be matched to observations of other stellar types and ages.
+
+On the Sun, spots have two regions: the dark, central umbra and the lighter, surrounding penumbra. A detailed review of sunspot behavior,
+including sizes and lifetimes, is \citet{solanki2003}, however the specific solar values are not relevant to this description. However,
+the spot area as a function of time is
+\begin{equation}
+    A(t) = \left\{
+    \begin{array}{lr}
+        A_0 e^{(t-t_0)/\tau}, & \text{if } t \leq t_0 \\
+        A_0 - W(t-t_0), & \text{if } t > t_0
+    \end{array}
+    \right\}
+\end{equation}
+where $A_0$ is the maximum area reached, $t_0$ is the time of the maximum, $\tau$ is the exponential growth rate, and $W$ is the linear decay rate.
+
+{\sc VSPEC} also takes in parameters describing the population of spots, such as the mean of the (lognormal) peak area distribution
+\citep{bogdan1988}, the equilibrium surface coverage, and the distribution style.
+
+\subsection{Faculae \label{subsec:faculae}}
+Faculae are magnetically-generated regions of the solar surface that usually appear as bright points near the limb; we employ the ``hot wall''
+model \citep{spruit1976} where faculae are described as three-dimensional pores in the stellar surface with a hot, bright wall, and a
+cool, dark floor, as shown in Figure \ref{fig:fac_struct}.
+
+Their three-dimensional structure causes faculae to behave differently depending on their angle from disk-center. Close to the limb,
+the hot wall is visible to the observer, and faculae appear as bright points. Near the center, however, the cool floor is exposed and
+faculae appear dark. To consider this effect in the faculae lightcurve, we compute the fraction of the facula's normalized area -- the
+area on the disk it would occupy as a flat spot -- that is occupied by each the hot wall and cool floor. This is done via numerical integral
+along the radius of the spot.
+
+\begin{equation}
+    f_{wall} = \frac{\int _{-R}^{R} Z_{\rm eff} dr}{\int _{-R}^{R} Z_{\rm eff} dr + \int _{-R}^{R} R_{\rm eff} dr}
+\end{equation}
+where $Z_{\rm eff}$ is the apparent height of the wall
+\begin{equation}
+    Z_{\rm eff} = \left\{
+    \begin{array}{lr}
+        Z_w \sin{\alpha}, & \text{if } \alpha \leq \alpha_{\rm crit} \\
+        2\sqrt{R^2 - r^2}\cos{\alpha}, & \text{if } \alpha > \alpha_{\rm crit}
+    \end{array}
+    \right\}
+\end{equation}
+and $R_{\rm eff}$ is the apparent width of the floor
+\begin{equation}
+    R_{\rm eff} = \left\{
+    \begin{array}{lr}
+        2\sqrt{R^2 - r^2} - Z_w\sin{\alpha}, & \text{if } \alpha \leq \alpha_{\rm crit} \\
+        0, & \text{if } \alpha > \alpha_{\rm crit}
+    \end{array}
+    \right\}
+\end{equation}
+for facula radius $R$, depth $Z_w$, and angle from disk-center $\alpha$. $r$ in this numerical scheme
+is defined as the distance from the center of the facula along the radial line connecting the facula center to the
+disk center. $\alpha_{\rm crit}$ is the value of alpha at which the floor is no longer visible and is defined to be $\arctan{\frac{2\sqrt{R^2-r^2}}{Z_w}}$.
+
+According to \citet{1997ApJ...484..479T}, faculae temperatures (of both the floor and wall) are dependent on the facula radius,
+while depth appears to be constant. They also find that the smallest faculae have no visible floor, and that even at disk center they
+appear as bright points. We parameterize the floor temperature to be
+\begin{equation}
+    \Delta T_{\rm eff, floor} = \left\{
+    \begin{array}{lr}
+    \text{Not visible}, & \text{if } r<r_{\rm min} \\
+    m_{\rm floor}(r-r_{\rm min}) + \Delta T_{\rm eff, floor, 0}, & \text{if } r \ge r_{\rm min}
+    \end{array}
+    \right\}
+\end{equation}
+
+Where $r$ is the radius of the facula, $r_{\rm min}$ is the minimum radius where the floor is visible, $\Delta  T_{\rm eff, floor, 0}$ is
+the difference between the floor and photosphere \teff at $r_{\rm min}$, and $m_{\rm floor}$ is the slope of the relationship with
+units of [temperature] [length]$^{-1}$.
+
+Similarly, the wall temperature is parameterized
+
+\begin{equation}
+    \Delta T_{\rm eff,wall} = m_{\rm wall}r + \Delta T_{\rm eff,wall,0}
+\end{equation}
+
+Where $\Delta T_{\rm eff,wall,0}$ is the temperature of a zero-radius facula and and $m_{\rm wall}$ is the slope of the radius-temperature
+relationship with units of [temperature] [length]$^{-1}$.
+These relationships can be defined by the user, for example setting $m=0$ for constant temperatures.
+
+Facula lifetimes are defined as the time it takes to decay by $e^{-2}$. Because faculae grow and decay exponentially
+at the same rate, each facula spends 1 lifetime with a radius greater than $e^{-1}$ of it's maximum. Each facula is born
+and dies at $e^{-2}$ of its maximum, effectively existing in the code for two lifetimes.
+
+\citet{hovis-afflerbach2022} suggest a typical facula lifetime to be on the order of 6 hours,
+with a distribution that resembles a Poisson function. However, we choose a lognormal distribution
+for both lifetime and maximum radius because it does not allow for these values to be 0. We choose to
+correlate lifetime and radius so that they are determined by the normally distributed random variable $\mu$;
+for each new facula, $\mu$ is randomly drawn and determines both the faclue lifetime and maximum radius.
+
+\begin{figure}
+    \centering
+    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_model.png}}
+    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_depth.pdf}}
+    \script{facula_depth.py}
+    \caption{
+        ``Hot Wall'' model of faculae. Faculae structure causes their contrast to be dependent on their distance
+        from the center of the disk. {\bf Top}: Depiction of a facula on the limb of a star. The hot wall is exposed
+        to the observer causing the pore to appear bright. At disk center, the cool floor is most visible. {\bf Bottom}:
+        The effects of depression depth and viewing angle on facula brightness. The 3D structure of faculae is most apparent
+        when radius $\sim$ depth. A toy flux model was used to demonstrate the shape of these curves, but their magnitudes
+        in practice depend on stellar spectral models.
+        }
+    \label{fig:fac_struct}
+\end{figure}
+
 \bibliography{syw,VSPEC}

### 2024-06-03 src/scripts/facula_depth.py

@@ -1,0 +1,89 @@
+"""
+Facula depth and the lightcurve.
+
+https://vspec-vsm.readthedocs.io/en/latest/auto_examples/plot_effect_facula_depth.html#sphx-glr-auto-examples-plot-effect-facula-depth-py
+"""
+
+from astropy import units as u
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import numpy as np
+import os
+import shutil
+
+from vspec_vsm.faculae import Facula
+from vspec_vsm.helpers import round_teff
+
+from pathlib import Path
+
+import paths
+
+fig,ax = plt.subplots(1,1,figsize=(4,3))
+
+outfig = paths.figures / 'facula_depth.pdf'
+
+
+def make_fac(_depth_over_rad: float) -> Facula:
+    """Return a default facula"""
+    radius = 0.005*u.R_sun
+    return Facula(
+        lat=0*u.deg,
+        lon=0*u.deg,
+        r_max=radius,
+        r_init=radius,
+        depth=radius*_depth_over_rad,
+        # None of the below parameters will affect this
+        # example, but we must set them to something
+        lifetime=1*u.day,
+        floor_teff_slope=0*u.K/u.km,
+        floor_teff_min_rad=1*u.km,
+        floor_teff_base_dteff=-100*u.K,
+        wall_teff_intercept=100*u.K,
+        wall_teff_slope=0*u.K/u.km
+    )
+
+
+rad_star = 0.15*u.R_sun
+WALL_BRIGHTNESS = 1.1
+FLOOR_BRIGNTNESS = 0.9
+
+
+def relative_flux(_facula: Facula, angle: u.Quantity) -> float:
+    """Get the contrast from a toy flux model"""
+    effective_area = _facula.effective_area(angle)
+    area_floor = effective_area[round_teff(_facula.floor_dteff)]
+    area_wall = effective_area[round_teff(_facula.wall_dteff)]
+    area_of_disk = np.pi*rad_star**2
+    floor_fraction = (
+        area_floor/area_of_disk).to_value(u.dimensionless_unscaled)
+    wall_fraction = (area_wall/area_of_disk).to_value(u.dimensionless_unscaled)
+    return 1 + floor_fraction*(FLOOR_BRIGNTNESS - 1) + wall_fraction*(WALL_BRIGHTNESS-1)
+
+angles = np.linspace(0, 90, 50)*u.deg
[…]

### 2024-06-04 src/tex/ms.tex
Copy up to examples section

@@
 
+\subsection{Flares \label{subsec:flares}}
+Flares are an important source of stellar variability on short timescales. We use the {\sc xoflares}
+package \citep{barclay2020} to generate realistic lightcurve shapes and assume a blackbody spectrum.
+{\sc xoflares} uses a flare profile function empirically derived by \citet{davenport2016} to
+describe a flare lightcurve. Temperature, $E$, FWHM, and the peak time together completely describe a flare's multiwavelength lightcurve.
+
+Before simulating an observation, {\sc VSPEC} pre-computes a population of flares based on the frequency-energy relationship from \citet{gao2022}:
+\begin{equation} \label{eq:flare_freq}
+    \log{(f{\rm [day]})} = \beta + \alpha~\log{(E/{\rm [erg]})}
+\end{equation}
+
+Where $f$ is the frequency of flares with energies $\ge E$.
+We compute the number of expected flares over a time duration and determine the number to create $N$ by a random Poisson draw.
+We then generate $N$ flares with energies determined by the quantile function:
+\begin{equation}
+    E = E_{\text{min}} (1-X)^{1/\alpha}
+\end{equation}
+
+where $E_{\text{min}}$ is the minimum considered flare energy and $X$ is a random number on the interval $[0,1)$.
+We set the default values $\alpha=-0.829$, $\beta=26.87$ from \citet{gao2022}. Figure \ref{fig:flare_freq}
+shows our simulation results compared to their powerlaw.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/flare_freq.pdf}
+    \caption{
+        {\bf Top}: Generated flare frequencies compared to those expected from \citet{gao2022}, generated using a 10,000 day simulation.
+        The $y$-axis shows the frequency of flares with energies {\em greater than or equal to} the value of the $x$-axis. Error bars
+        based on the square root of the number of observed flares as this is a Poisson process.
+        {\bf Bottom}: Lightcurve of a star flaring with the same powerlaw slope as the top panel, but the intercept ($\beta$) has been increased by 0.3
+        for visual effect. The mean flare temperature is $9000$ K and the mean FWHM is $3$ hours.
+        }
+    \script{flare_freq.py}
+    \label{fig:flare_freq}
+\end{figure}
+
+\subsection{Granulation}
+Granulation is a source of stellar variability that arises from convection near the surface of the star.
+The result is a stellar surface that is not constant in temperature and that changes on very short timescales.
+Hydrodynamic simulations of stellar atmospheres \citep[e.g.][]{magic2014} describe hot granules of rising gas surrounded by cooler,
+sinking regions with a \teff a few percent lower than the nominal value.
+
+We model granulation as a global process, and its effects are computed after the effects of spots and faculae.
+Of the remaining ``quiet'' photosphere (i.e. the regions not covered in spots of faculae), a fraction is computed to be part of the cool
+inter-granule surface. The surface coverage of the cool region at any given time is computed by a Gaussian process (GP) using
+the \sc{tinygp} package \citep{foreman-mackey2024} following the methodology of \citet{gordon2020}.
+The GP uses a custom kernel function based on a power spectrum \citep{anderson1990,kallinger2014} to produce random changes in the coverage.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/granulation_lightcurve.pdf}
+    \caption{
+        Effect of granulation on the lightcurve of a star. This simulation assumes a mean
+        10\% coverage by inter-granule region whose temperature is 300 K lower than the surrounding photosphere.
+        It also assumes that the coverage value varies with a magnitude of 1\% on a timescale of 6 hours.
+        }
+    \script{granule_lc.py}
+    \label{fig:gran_lc}
+a\end{figure}
+
+\section{Planet Model \label{sec:pl_model}}
+Our model uses the Planetary Spectrum Generator\footnote{\href{https://psg.gsfc.nasa.gov/}{https://psg.gsfc.nasa.gov/}}
+\citep[PSG,][]{villanueva2018} to calculate the planetary spectrum. We upload a 3D Global Circulation Model (GCM) to PSG's
+Global Emission Spectra (GlobES) application to retrieve thermal and reflected spectra for various points in the planet's phase curve. \TJ{cite Vincent's paper}
+
+\subsection{Interfacing through {\sc pypsg}}
+\ref{subsec:pypsg}
+{\sc VSPEC} interfaces with PSG using the {\sc pypsg} package \TJ{link}, which is essentially a Python wrapper for the PSG API system.
+
+\TJ{Write much more}
+
+\subsection{3D Planet models}
+{\sc VSPEC} allows the user to upload any desired GCM. However, for questions that focus on detectability,
+the specific physics of the GCM are less important than properties like day/night temperature, surface pressure, or
+the major absorbers in the atmosphere. To simplify these studies, {\sc VSPEC} has a built-in GCM that takes in just a
+few parameters and produces a simple-yet-robust exoplanet climate. Simplicity allows the user to know everything about
+the model and reduce surprises, but its basis in thermodynamics and robust energy balance make this model fit for publication-level science \TJ{is this really true? probably not}.
+
+The {\sc VSPEC} GCM centers around a 2D temperature map of the planet's surface. Based on \citet{cowan2011}, this map
+is constructed completely given the incident flux from the star (parameterized by the host \teff and radius
+and the planet's semimajor axis and Bond albedo) and the unitless thermal redistribution efficiency $\epsilon$.
+This efficiency is 0 for a planet that reradiates instantly and $\gg 1$ for an atmosphere that efficiently redistributes heat to its night side.
+Figure \ref{fig:thermal_inertia} shows a temperature map example for an Earth-like planet with $\epsilon = 2\pi$.
+These temperature maps are validated by the energy balance of the planet; {\sc VSPEC} computes the incident and emergent
+flux across the planet and asserts that they agree.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/cowan_gcm.pdf}
+    \caption{
+        Example of a surface temperature map created by {\sc VSPEC}. The solar flux at Earth,
+        a Bond albedo of 0.3, and $\epsilon = 2\pi$ were used in this case.
+        Notice that the hottest point is offset from the sub-stellar point due to thermal inertia.
+    }
+    \script{cowan_gcm.py}
+    \label{fig:thermal_inertia}
+\end{figure}
+
+Now that we have the surface temperature at each GCM coordinate we can build the atmosphere on top of it.
+We choose a surface pressure for our model and generate a pressure profile as a function of layer. It is
+important to note that the choice of the pressure profile -- provided it is scaled to the surface pressure and sampled in a reasonable manner -- is
+arbitrary and will not affect the spectrum of the planet. This is because PSG calculates the altitude and thickness of each layer based on the
+pressure and surface gravity; the pressure profile is essentially a proxy for the sampling of atmosphere layers as a function of altitude.
+
+Next we create an adiabatic temperature profile at each GCM coordinate. An adiabat assumes that
+\begin{equation}
+    T^\gamma P^{1 - \gamma} = \text{const.}
+\end{equation}
+where $\gamma$ is the adiabatic index. Therefore the temperature profile can be computed
+\begin{equation}
+    T = T_{\rm surf} \left(\frac{P}{P_{\rm surf}} \right)^{1 - \frac{1}{\gamma} }
+\end{equation}
+
+Finally, we populate our atmosphere with molecules. The user specifies the species and abundance and atmosphere is filled accordingly with a constant molecular suite.
+
+The caveats of this approach lie in the simplified physics of the atmosphere structure. The heat-redistribution model does not
+capture the effects of latitudinal mixing; an adiabatic temperature profile will never exhibit the temperature inversion
+that we know form in complex atmospheres like Earth's. Similarly, setting a constant abundance does not allow for water vapor to
+vary with temperature as it does on Earth. We must also be careful not to add so much of an absorbing species that the energy balance of
+the atmosphere would change; instead we treat them as trace species in an otherwise transparent atmosphere. Our simplified approach
+allows us to easily examine a large parameter space of atmospheres whose behaviors are easy to predict.
+
 \bibliography{syw,VSPEC}

### 2024-06-04 src/scripts/cowan_gcm.py

@@ -1,0 +1,50 @@
+"""
+https://vspec.readthedocs.io/en/latest/auto_examples/other/gcm/plot_cowan2011.html#sphx-glr-auto-examples-other-gcm-plot-cowan2011-py
+"""
+
+import matplotlib.pyplot as plt
+import numpy as np
+from astropy import units as u
+import cartopy.crs as ccrs
+
+import VSPEC.gcm.heat_transfer as ht
+
+import paths
+
+outfile = paths.figures / 'cowan_gcm.pdf'
+
+epsilon = 2*np.pi
+star_teff = 5800*u.K
+albedo = 0.3
+r_star = 1*u.R_sun
+r_orbit = 1*u.AU
+
+tmap = ht.TemperatureMap.from_planet(
+    epsilon=epsilon,
+    star_teff=star_teff,
+    albedo=albedo,
+    r_star=r_star,
+    r_orbit=r_orbit
+)
+lons = np.linspace(-180,180,90,endpoint=False)*u.deg
+lats = np.linspace(-90,90,46,endpoint=True)*u.deg
+
+longrid,latgrid = np.meshgrid(lons,lats)
+data = tmap.eval(lon=longrid,lat=latgrid,alpha=0)
+
+fig = plt.figure(figsize=(4,3))
+proj = ccrs.Robinson(central_longitude=0)
+ax = fig.add_subplot(projection=proj)
+
+im = ax.pcolormesh(lons.to_value(u.deg),lats.to_value(u.deg),data.to_value(u.K),cmap='gist_heat',transform=ccrs.PlateCarree())
+gl = ax.gridlines(crs=ccrs.PlateCarree(),draw_labels=True,
+    color='grey', alpha=0.8, linestyle='--')
+gl.top_xlabels = False
+gl.right_ylabels = False
+
+
+_=fig.colorbar(im,ax=ax,label='Surface Temperature (K)',pad=0.15,shrink=0.6)
+
+fig.tight_layout()
+
+fig.savefig(outfile)

### 2024-06-04 src/scripts/flare_freq.py

@@ -1,0 +1,85 @@
+"""
+Flare frequency and the lightcurve.
+"""
+
+from astropy import units as u
+import numpy as np
+import matplotlib.pyplot as plt
+
+from vspec_vsm.flares import FlareGenerator
+from VSPEC import ObservationModel, PhaseAnalyzer
+import pypsg
+
+import paths
+
+pypsg.docker.set_url_and_run()
+
+SEED = 10
+outfile = paths.figures / 'flare_freq.pdf'
+cfg_path = paths.scripts / 'flare_lc.yaml'
+
+dt = 10000*u.day  # a long time.
+gen = FlareGenerator(
+    dist_teff_mean=9000*u.K,
+    dist_teff_sigma=1000*u.K,
+    dist_fwhm_mean=3*u.hr,
+    dist_fwhm_logsigma=0.3,
+    min_energy=1e33*u.erg,
+    cluster_size=4,
+    rng=np.random.default_rng(seed=SEED)
+)
+Es = np.logspace(np.log10(gen.min_energy.to_value(u.erg)),
+                 np.log10(gen.min_energy.to_value(u.erg))+4)*u.erg
+
+flares = gen.generate_flare_series(dt)
+
+energies = np.array([flare.energy.to_value(u.erg) for flare in flares])
+
+energies_ge_E = np.array([np.sum(energies >= E) for E in Es.to_value(u.erg)])
+
+measured_freq = energies_ge_E/dt
+measured_freq_err = np.sqrt(energies_ge_E)/dt
+
+beta = gen.beta
+alpha = gen.alpha
+
+expected_log_freq = beta + alpha*np.log10(Es/u.erg)
+expected_freq = 10**expected_log_freq / u.day
+
+ratio = np.where(energies_ge_E > 0, measured_freq/expected_freq, np.nan)
+ratio_err = np.where(
+    energies_ge_E > 0, measured_freq_err/expected_freq, np.nan)
+
+fig, axes = plt.subplots(2, 1,figsize=(4,5),height_ratios=[1,2])
+
+axes[0].plot(Es, expected_freq, c='xkcd:azure', label='Expected')
+axes[0].errorbar(Es, measured_freq, yerr=measured_freq_err, fmt='o',
+                 color='xkcd:rose pink', label='Observed', markersize=3)
+axes[0].set_xlabel('Energy (erg)')
+axes[0].set_ylabel('Frequency (1/day)')
+axes[0].set_xscale('log')
[…]

### 2024-06-04 src/scripts/granule_lc.py

@@ -1,0 +1,131 @@
+"""
+Lightcurve due to granulation
+
+https://vspec.readthedocs.io/en/latest/auto_examples/other/star/plot_granulation_lightcurve.html#sphx-glr-auto-examples-other-star-plot-granulation-lightcurve-py
+"""
+
+from pathlib import Path
+from astropy import units as u
+import matplotlib.pyplot as plt
+import pypsg
+
+from VSPEC import ObservationModel,PhaseAnalyzer
+from VSPEC import params
+from VSPEC.params.gcm import vspec_to_pygcm
+
+
+import paths
+
+outfile = paths.figures / 'granulation_lightcurve.pdf'
+
+SEED = 32
+pypsg.docker.set_url_and_run()
+
+
+header = params.Header(
+    data_path=Path('.vspec/granulation_lightcurve'),
+    teff_min=3000*u.K,
+    teff_max=3400*u.K,
+    seed=SEED,verbose=0
+)
+star = params.StarParameters(
+    psg_star_template='M',
+    teff=3300*u.K,
+    mass = 0.1*u.M_sun,
+    radius=0.15*u.R_sun,
+    period = 10*u.day,
+    misalignment_dir=0*u.deg,
+    misalignment=0*u.deg,
+    ld = params.LimbDarkeningParameters.solar(),
+    faculae=params.FaculaParameters.none(),
+    spots=params.SpotParameters.none(),
+    flares=params.FlareParameters.none(),
+    granulation=params.GranulationParameters(
+        mean=0.2,
+        amp=0.005,
+        period=6*u.hr,
+        dteff=300*u.K
+    ),
+    grid_params=(500, 1000),
+    spectral_grid='default'
+)
+planet = params.PlanetParameters.std(init_phase=180*u.deg,init_substellar_lon=0*u.deg)
+system = params.SystemParameters(
+    distance=1.3*u.pc,
+    inclination=30*u.deg,
+    phase_of_periasteron=0*u.deg
+)
+observation = params.ObservationParameters(
+    observation_time=3*u.day,
+    integration_time=30*u.min
+)
+psg_params = params.psgParameters(
+    gcm_binning=200,
+    phase_binning=1,
+    use_molecular_signatures=True,
+    nmax=0,
+    lmax=0,
+    continuum=['Rayleigh', 'Refraction', 'CIA_all'],
+)
[…]

### 2024-06-04 src/tex/ms.tex
Add moran23 example

@@
 
+\section{Examples \label{sec:examples}}
+
+\subsection{Transit of a spotted star}
+
+In this section we will simulate the transit of a bare rock across a spotted star to demonstrate how an inhomogeneous
+stellar surface can lead to a false atmospheric signal -- a phenomenon known as the ``transit light source effect'' \citep{rackham2018}.
+During transit, the planetary atmosphere is illuminated by the portion of the star directly behind the planet. The depth of the transit
+should be measured with respect for the spectrum of this region of the stellar surface. However, the disk-integrated stellar spectrum is often used instead.
+In the case that the stellar surface is not homogeneous, the transit signal is contaminated by the star.
+\citet{moran2023} first observed this effect using JWST from super-Earth GJ 486b.
+
+Whenever possible, we draw parameter values from \citet{moran2023} and the NExScI Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/GJ1214b}}.
+We use the JWST NIRSpec/G395H instrument setup but set $R=200$ to account for the resolution of the reduction.
+Additionally we observe for 3.53 hours with an 8 minute cadence. Mid-transit is reached exactly halfway through the observation.
+
+We will do three \texttt{VSPEC} simulations:
+\begin{enumerate}
+    \item Bare rock, no spots.
+    \item 1 bar \ce{H2O} atmosphere, no spots.
+    \item Bare rock, 2.5\% coverage by 2700 K spots
+\end{enumerate}
+
+We expect in case 1 to observe a flat spectrum, and in case 2 to observe additional
+absorption from the \ce{H2O} atmosphere. Note also that we disabled \ce{H2O}-\ce{H2O} collision-induced absorption (CIA) for
+these simulations because the effect was not considered in the original analysis of GJ 486b.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/moran23.pdf}
+    \caption{
+        The results from our spotted transit experiment. We find that the addition of
+        spots to a star's surface can give an otherwise flat transit spectrum features that appear to be molecular bands.
+        This is in agreement with the results of \citet{moran2023}.
+        }
+    \label{fig:moran_transit}
+\end{figure}
+
+We find that, as expected, the addition of spots changes the transit depth in a way that depends on wavelength -- mimicking
+light lost to absorbers in a planetary atmosphere. Figure \ref{fig:moran_transit} shows us that, in this case, we observe false absorption;
+this is due to an increase in the spotted fraction of the surface relative to the quiet photosphere. However, if a spot was blocked by
+the planet we might see similar features inverted to mimic emission.
+
 \bibliography{syw,VSPEC}

### 2024-06-04 src/scripts/moran23.py

@@ -1,0 +1,283 @@
+from pathlib import Path
+import numpy as np
+import matplotlib.pyplot as plt
+from astropy import units as u
+import pypsg
+
+from VSPEC import ObservationModel,PhaseAnalyzer
+from VSPEC import params
+from VSPEC.config import MSH
+
+import paths
+
+outfile = paths.figures / 'moran23.pdf'
+
+SEED = 10
+PRE_TRANSIT = 8
+IN_TRANSIT = 13
+pypsg.docker.set_url_and_run()
+
+# Instrument
+
+inst = params.InstrumentParameters(
+    telescope=params.SingleDishParameters.jwst(),
+    bandpass=params.BandpassParameters(
+        wl_blue=2.87*u.um,
+        wl_red=5.14*u.um,
+        resolving_power=200,
+        wavelength_unit=u.um,
+        flux_unit=u.Unit('W m-2 um-1')
+    ),
+    detector=params.DetectorParameters(
+        beam_width=5*u.arcsec,
+        integration_time=0.5*u.s,
+        ccd=params.ccdParameters(
+            pixel_sampling=8,
+            read_noise=16.8*u.electron,
+            dark_current=0.005*u.electron/u.s,
+            throughput=0.3,
+            emissivity=0.1,
+            temperature=50*u.K
+        )
+    )
+)
+
+# Observation
+
+observation = params.ObservationParameters(
+    observation_time=3.53*u.hr,
+    integration_time=8*u.min
+)
+
+# PSG
+
+psg_kwargs = dict(
+    gcm_binning=200,
+    phase_binning=1,
+    nmax=0,
+    lmax=0,
+    continuum=['Rayleigh', 'Refraction'],
+)
+psg_params = params.psgParameters(
+    use_molecular_signatures=True,
+    **psg_kwargs
+)
+psg_no_atm = params.psgParameters(
+    use_molecular_signatures=False,
+    **psg_kwargs
+)
+
+# Star and Planet
+
+star_teff = 3343*u.K
+star_rad = 0.339*u.R_sun
+a_rstar = 11.229 # Moran+23 Table 1
+rp_rstar = 0.03709 # Moran+23 Table 1
+planet_rad = star_rad * rp_rstar
+orbit_rad = star_rad * a_rstar
+orbit_period = 1.467119*u.day
+planet_rot_period = orbit_period
+star_rot_period = np.inf*u.s # assume the star does not change.
[…]

### 2024-06-04 src/tex/ms.tex
Finish copying

@@
         }
+    \script{moran23.py}
     \label{fig:moran_transit}
@@
 
+\subsection{Phase Curve}
+Analysis of JWST MIRI-LRS phase curves of GJ 1214b by \citet{kempton2023} did not consider stellar
+contamination because of the slow rotation rate of the host star \citep[approximately 1/80$^{\text{th}}$ the orbital frequency,][]{cloutier2021}.
+In this example we use {\sc VSPEC} to demonstrate that this is a reasonable assumption.
+
+We initialize {\sc VSPEC} to run a 41.0 hour observation with a cadence of 15 minutes. We use stellar and planetary parameters
+from the NExScI Exoplanet Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/GJ1214b}}.
+The GCM used has a thermal inertia of $\epsilon = 6$ and a 1 bar \ce{CO2} atmosphere.
+
+We initialized two stellar models: one with nothing more than a 3250 K surface, and another with 20\% coverage by 2700 K spots.
+
+After running the {\sc VSPEC} simulations, we bin into 0.5 $\mu m$ wavelength channels and normalize each channel to the two eclipses.
+We assume the star behaves linearly between eclipse measurements and attribute any deviation to the planet.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{sphx_glr_plot_gj1214b_kempton23_006.png}
+    \caption{
+        MIRI phase curves for GJ 1214b binned by wavelength. Compare to \citet{kempton2023} Extended Data Figure 1.
+        While the addition of stellar variability does have a visible effect on the lightcurve, it is much smaller than the thermal emission by the exoplanet.
+        }
+    \script{kempton23.py}
+    \label{fig:gj1214b}
+\end{figure}
+
+We show in Figure \ref{fig:gj1214b} that the stellar properties assumed in this example do not generate enough contamination to
+significantly change the observed phase curve of GJ 1214b. At it's worst (i.e. near transit) the observed spectrum varies by 50 ppm due to stellar
+contamination -- compared to the 700 ppm relative flux of the planet at $8.5 {\mu m}$. Similar studies could be done to assess the potential for
+stellar contamination in other exoplanetary systems in order to judge fitness as a target.
+
+\subsection{MIRECLE Phase Curve}
+In this example we will look at the best-case target for the Mid-IR Exoplanet CLimate Explorer \citep[MIRECLE,][]{mandell2022}
+mission concept: Proxima Centauri b with no stellar variability. We will instead examine the expected noise produced by this observation
+and assess the detectability of the planet in a single orbit. {\sc VSPEC} uses the noise models built into PSG, allowing native support for
+noise due to the instrument, source, and astrophysical background.
+
+We use \texttt{VSPEC}'s built-in MIRECLE instrument template, which includes a 2 m aperture and observes from 1-18 $\mu m$ with $R=50$.
+We draw stellar and bulk planetary parameters from the NExScI Exoplanet Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/proxima\%20cen}}
+and assume $i=85^\circ$ and Earth density. The GCM we use has a Bond albedo of 0.3, $\epsilon=1.5$, $P_{\rm surf} = 1 \text{bar}$ and a 100 ppm \ce{CO2} atmosphere with a \ce{N2} background.
+We ran the simulation for one orbit with an 24 hour cadence. Because this is a best-case scenario, we assumed perfect knowledge of the stellar spectrum.
+We are left with the planetary spectrum and photon noise from the star.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/mirecle.pdf}
+    \caption{
+        Simulated spectra of Proxima Centauri b observed by MIRECLE. Noise and error bars based on 24 hour integrations with a 2 m aperture MIRECLE
+        space telescope.
+        }
+    \script{mirecle.py}
+    \label{fig:mirecle}
+\end{figure}
+
+\section{Conclusion \label{sec:conclusion}}
+\TJ{Talk about what we can do next.}
+
 \bibliography{syw,VSPEC}

### 2024-06-04 src/scripts/kempton23.py

@@ -1,0 +1,228 @@
+from pathlib import Path
+import numpy as np
+import matplotlib.pyplot as plt
+from astropy import units as u
+from cartopy import crs as ccrs
+import pypsg
+from pypsg.globes import PyGCM
+
+from VSPEC import ObservationModel,PhaseAnalyzer
+from VSPEC import params
+from VSPEC.config import MSH
+
+import paths
+
+SEED = 1214
+pypsg.docker.set_url_and_run()
+
+outfile = paths.figures / 'kempton23.pdf'
+
+# Instrument
+inst = params.InstrumentParameters.miri_lrs()
+
+# Observation
+observation = params.ObservationParameters(
+    observation_time=41.0*u.hr,
+    integration_time=15*u.min
+)
+
+# PSG
+psg_params = params.psgParameters(
+    gcm_binning=200,
+    phase_binning=1,
+    nmax=0,
+    lmax=0,
+    continuum=['Rayleigh', 'Refraction', 'CIA_all'],
+    use_molecular_signatures=True
+)
+
+# Star and Planet
+star_teff = 3250*u.K
+star_rad = 0.215*u.R_sun
+planet_rad = 2.742*u.R_earth
+orbit_rad = 0.01490*u.AU
+orbit_period = 1.58040433*u.day
+planet_rot_period = orbit_period
+star_rot_period = 120 * u.day
+planet_mass = 8.17*u.M_earth
+star_mass = 0.178*u.M_sun
+inclination = 88.7*u.deg
+
+start_time_before_eclipse = 2*u.hr
+angle_before_eclipse = (2*np.pi*u.rad * start_time_before_eclipse/orbit_period).to(u.deg)
+initial_phase = 0*u.deg - angle_before_eclipse
+
+planet_params = params.PlanetParameters(
+    name='GJ1214b',
+    radius=planet_rad,
+    gravity=params.GravityParameters('kg',planet_mass),
+    semimajor_axis=orbit_rad,
+    orbit_period=orbit_period,
+    rotation_period=planet_rot_period,
+    eccentricity=0,
+    obliquity=0*u.deg,
+    obliquity_direction=0*u.deg,
+    init_phase=initial_phase,
+    init_substellar_lon=0*u.deg
+)
+
+system_params = params.SystemParameters(
+    distance=14.6427*u.pc,
+    inclination=inclination,
+    phase_of_periasteron=0*u.deg
+)
+
[…]

### 2024-06-04 src/scripts/mirecle.py

@@ -1,0 +1,168 @@
+"""
+Mirecle phase curve
+"""
+
+from pathlib import Path
+import numpy as np
+import matplotlib.pyplot as plt
+from astropy import units as u
+import pypsg
+
+from VSPEC import ObservationModel,PhaseAnalyzer
+from VSPEC import params
+
+import paths
+
+outfile = paths.figures / 'mirecle.pdf'
+
+SEED = 10
+pypsg.docker.set_url_and_run()
+
+# Instrument
+inst = params.InstrumentParameters.mirecle()
+
+# Observation
+
+observation = params.ObservationParameters(
+    observation_time=12*u.day,
+    integration_time=24*u.hr
+)
+
+# PSG
+psg_params = params.psgParameters(
+    use_molecular_signatures=True,
+    gcm_binning=200,
+    phase_binning=1,
+    nmax=0,
+    lmax=0,
+    continuum=['Rayleigh', 'Refraction','CIA_all'],
+    )
+# Star and Planet
+
+star_teff = 2900*u.K
+star_rad = 0.141*u.R_sun
+inclination = 85*u.deg
+planet_mass = 1.07*u.M_earth/np.sin(inclination)
+planet_rad = 1*u.R_earth * planet_mass.to_value(u.M_earth)**3
+orbit_rad = 0.04856*u.AU
+orbit_period = 11.18*u.day
+planet_rot_period = orbit_period
+star_rot_period = 90*u.day
+planet_mass = 1.07*u.M_earth
+star_mass = 0.122*u.M_sun
+
+
+
+initial_phase = 180*u.deg
+
+planet_params = params.PlanetParameters(
+    name='proxcenb',
+    radius=planet_rad,
+    gravity=params.GravityParameters('kg',planet_mass),
+    semimajor_axis=orbit_rad,
+    orbit_period=orbit_period,
+    rotation_period=planet_rot_period,
+    eccentricity=0,
+    obliquity=0*u.deg,
+    obliquity_direction=0*u.deg,
+    init_phase=initial_phase,
+    init_substellar_lon=0*u.deg
+)
+
+system_params = params.SystemParameters(
+    distance=1.3*u.pc,
+    inclination=inclination,
+    phase_of_periasteron=0*u.deg
+)
+
+
+star_dict = {
+    'teff': star_teff,
+    'radius': star_rad
+}
+planet_dict = {'semimajor_axis': orbit_rad}
+
+gcm_dict = {
[…]

### 2024-06-04 src/scripts/setup_psg.py
Load API key from secrets

@@ -1,0 +1,14 @@
+"""
+When running the GHA it is important that the PSG API key is set.
+"""
+
+import os
+
+import pypsg
+
+try:
+    key = os.environ['PSG_API_KEY']
+    pypsg.settings.save_settings(api_key=key)
+
+except KeyError:
+    assert pypsg.docker.is_psg_installed()

### 2024-06-04 src/tex/ms.tex
Many bugfixes. The expectation is the the GHA should now be sucessfull

@@
 \usepackage{showyourwork}
+\usepackage[version=4]{mhchem}
 
@@
 inter-granule surface. The surface coverage of the cool region at any given time is computed by a Gaussian process (GP) using
-the \sc{tinygp} package \citep{foreman-mackey2024} following the methodology of \citet{gordon2020}.
+the {\sc tinygp} package \citep{foreman-mackey2024} following the methodology of \citet{gordon2020}.
 The GP uses a custom kernel function based on a power spectrum \citep{anderson1990,kallinger2014} to produce random changes in the coverage.
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{sphx_glr_plot_gj1214b_kempton23_006.png}
+    \includegraphics[width=0.5\textwidth]{figures/kempton23.pdf}
     \caption{

### 2024-06-04 src/scripts/facula_depth.py

@@ -8,13 +8,11 @@ from astropy import units as u
 import matplotlib.pyplot as plt
 import matplotlib.cm as cm
 import numpy as np
-import os
 import shutil
 
 from vspec_vsm.faculae import Facula
 from vspec_vsm.helpers import round_teff
 
-from pathlib import Path
 
 import paths
 

### 2024-06-04 src/scripts/flare_freq.py

@@ -12,6 +12,7 @@ import pypsg
 
 import paths
 
+import setup_psg;setup_psg.setup_psg()
 pypsg.docker.set_url_and_run()
 
 SEED = 10

### 2024-06-04 src/scripts/granule_lc.py

@@ -19,6 +19,7 @@ import paths
 outfile = paths.figures / 'granulation_lightcurve.pdf'
 
 SEED = 32
+import setup_psg;setup_psg.setup_psg()
 pypsg.docker.set_url_and_run()
 
 

### 2024-06-04 src/scripts/kempton23.py

@@ -2,9 +2,7 @@ from pathlib import Path
 import numpy as np
 import matplotlib.pyplot as plt
 from astropy import units as u
-from cartopy import crs as ccrs
 import pypsg
-from pypsg.globes import PyGCM
 
 from VSPEC import ObservationModel,PhaseAnalyzer
 from VSPEC import params
@@ -13,6 +11,7 @@ from VSPEC.config import MSH
 import paths
 
 SEED = 1214
+import setup_psg;setup_psg.setup_psg()
 pypsg.docker.set_url_and_run()
 
 outfile = paths.figures / 'kempton23.pdf'

### 2024-06-04 src/scripts/mirecle.py

@@ -16,6 +16,7 @@ import paths
 outfile = paths.figures / 'mirecle.pdf'
 
 SEED = 10
+import setup_psg;setup_psg.setup_psg()
 pypsg.docker.set_url_and_run()
 
 # Instrument

### 2024-06-04 src/scripts/moran23.py

@@ -15,6 +15,7 @@ outfile = paths.figures / 'moran23.pdf'
 SEED = 10
 PRE_TRANSIT = 8
 IN_TRANSIT = 13
+import setup_psg;setup_psg.setup_psg()
 pypsg.docker.set_url_and_run()
 
 # Instrument
@@ -253,9 +254,9 @@ filename = paths.static / 'moran2023_fig3.txt'
 df = pd.read_fwf(filename,colspecs=[(0,8),(9,14),(15,20),(21,25),(26,28)],
     header=20,names=['Reduction','Wave','Width','Depth','e_Depth'])
 used_eureka = df['Reduction']=='Eureka'
-moran_x = df.loc[used_eureka,'Wave']
-moran_y = df.loc[used_eureka,'Depth']
-moran_yerr = df.loc[used_eureka,'e_Depth']
+moran_x = df.loc[used_eureka,'Wave'].values
+moran_y = df.loc[used_eureka,'Depth'].values
+moran_yerr = df.loc[used_eureka,'e_Depth'].values
 moran_mean = np.mean(moran_y)
 
 fig, ax = plt.subplots(1,1,figsize=(4,3),tight_layout=True)

### 2024-06-04 src/scripts/setup_psg.py

@@ -6,9 +6,13 @@ import os
 
 import pypsg
 
-try:
-    key = os.environ['PSG_API_KEY']
-    pypsg.settings.save_settings(api_key=key)
+def setup_psg():
+    try:
+        key = os.environ['PSG_API_KEY']
+        pypsg.settings.save_settings(api_key=key)
+        print('API Key was successfully set.')
 
-except KeyError:
-    assert pypsg.docker.is_psg_installed()
+    except KeyError:
+        print('API Key was not set.')
+        assert pypsg.docker.is_psg_installed()
+        print('PSG is installed.')

### 2024-06-04 src/scripts/surface_map_and_lc.py

@@ -16,6 +16,7 @@ from vspec_vsm.config import MSH
 
 import pypsg
 
+import setup_psg;setup_psg.setup_psg()
 pypsg.docker.set_url_and_run()
 
 from VSPEC import ObservationModel,PhaseAnalyzer

### 2024-06-06 src/tex/ms.tex
Add Avi's additions from overleaf

@@
     We present the Variable Star Phase Curve (\href{https://github.com/VSPEC-collab/VSPEC}{ \sc VSPEC}) code,
-    a python package to simulate spectroscopic observations of rocky exoplanets in the presence of stellar variability.
+    a python package to simulate combined-light spectroscopic observations of exoplanets in the presence of stellar variability and inhomogeneity.
     {\sc VSPEC} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG GlobES) application along with a custom-built
-    stellar model to produce spectroscopic light curves of the planet-host system. We also describe associated codes. \TJ{These codes should be named here and described a little}
+    stellar model based on an existing grid of stellar photosphere models to produce spectroscopic light curves of the planet-host system.
+    VSPEC can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements,
+    and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs.
+    We also describe associated codes. \TJ{These codes should be named here and described a little}
 \end{abstract}
@@
 and the Atmospheric Remote-sensing Infrared Exoplanet Large-survey (ARIEL), spectral analysis of exoplanet atmospheres
-is increasingly sensitive to contamination due to stellar inhomogenies (spots, granulation, etc.).
+is increasingly sensitive to contamination due to stellar inhomogeneities (e.g. spots, granulation) and stochastic stellar variability (e.g. flares).
 
@@
 \citep[TLS,][see also \citet{apai2018,barclay2021,garcia2022,barclay2023}]{rackham2018}
-occurs when the occulted region of the stellar surface is not representitive of the disk-integrated spectrum
+occurs when the occulted region of the stellar surface is not representative of the disk-integrated spectrum.
 
@@
 
-We present {\sc VSPEC}: Variable Star PhasE Curve\footnote{\url{https://github.com/VSPEC-collab/VSPEC}},
+In this paper we present present {\sc VSPEC}: Variable Star PhasE Curve\footnote{\url{https://github.com/VSPEC-collab/VSPEC}},
 an open-source Python 3 package to simulate observations of exoplanet systems with variable host stars.

### 2024-06-06 src/tex/ms.tex
Write a description of VSPEC overall

@@
 \newcommand{\TJ}[1]{\textcolor{tedcommentcolor}{#1}}
+\newcommand{\vspec}[1]{\texttt{VSPEC}#1}
+% \def\vspec#{\texttt{VSPEC}}
 
-\shorttitle{\sc VSPEC}
+\shorttitle{\vspec{}}
 \shortauthors{Johnson et al.}
@@
 % Title
-\title{The {\sc VSPEC} code: Spectroscopic phase curves of 3D exoplanet models in the presence of stellar variability}
+% \title{The \vspec{} code: Spectroscopic phase curves of 3D exoplanet models in the presence of stellar variability}
+\title{\vspec{} and Friends: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.}
 
@@
 \begin{abstract}
-    We present the Variable Star Phase Curve (\href{https://github.com/VSPEC-collab/VSPEC}{ \sc VSPEC}) code,
+    We present the Variable Star Phase Curve (\href{https://github.com/VSPEC-collab/VSPEC} \vspec{}) code,
     a python package to simulate combined-light spectroscopic observations of exoplanets in the presence of stellar variability and inhomogeneity.
-    {\sc VSPEC} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG GlobES) application along with a custom-built
+    \vspec{} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG GlobES) application along with a custom-built
     stellar model based on an existing grid of stellar photosphere models to produce spectroscopic light curves of the planet-host system.
-    VSPEC can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements,
-    and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs.
-    We also describe associated codes. \TJ{These codes should be named here and described a little}
+    \vspec{} can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements,
+    and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs. We additionally present
+    a set of codes developed alongside \vspec{}, including the stellar surface model \texttt{vspec-vsm}, the stellar spectral grid interpolation code
+    Gridpolator, and a Python interface for PSG \texttt{pypsg}.
 \end{abstract}
@@
 
-In this paper we present present {\sc VSPEC}: Variable Star PhasE Curve\footnote{\url{https://github.com/VSPEC-collab/VSPEC}},
+In this paper we present \vspec{}: Variable Star PhasE Curve\footnote{\url{https://github.com/VSPEC-collab/VSPEC}},
 an open-source Python 3 package to simulate observations of exoplanet systems with variable host stars.
-{\sc VSPEC} combines NASA's Planetary Spectrum Generator \citep[PSG,][]{villanueva2018} with a grid of
-custom PHOENIX models \citep{husser2013} to produce multiwavelength phase curves of the system, accounting
-for a 3D planetary atmosphere, transits and eclipses, and multiple sources of stellar variability.
+\vspec{} itself is merely an interface that pulls together a variable star surface model \texttt{vspec-vsm}\footnote{\url{https://github.com/VSPEC-collab/vspec-vsm}},
+a stellar spectra grid interpolation code Gridpolator\footnote{\url{https://github.com/VSPEC-collab/GridPolator}}, and a Python interface for
+the Planetary Spectrum Generator \citep[PSG,][]{villanueva2018}, \texttt{pypsg}. All of these codes together allow \vspec{} to account for
+3D planetary atmospheres, time-resolved effects of both planet and star, many geometries including transit and eclipse, and realistic noise modeling.
 
-In this paper we will describe the science behind {\sc VSPEC} and demonstrate examples of its use.
+\TJ{Change this later-->}
+In this paper we will describe the science behind \vspec{} and demonstrate examples of its use.
 Section \ref{sec:star} describes the stellar model and the available sources of variability.
 Section \ref{sec:pl_model} discusses modeling the planetary atmosphere, including PSG/GlobES and the
-built-in GCM. In Section \ref{sec:examples} we provide examples of {\sc VSPEC} use cases and demonstrate
+built-in GCM. In Section \ref{sec:examples} we provide examples of \vspec{} use cases and demonstrate
 its value to the exoplanet community. Finally, in Section \ref{sec:conclusion} we will discuss the future of the code and issues it might have.
 
-{Stellar Model \label{sec:star}}
+\section{Using \vspec{}}
+\label{sec:vspec}
 
-The {\sc VSPEC} star model is designed modularly to allow
+The flow of \vspec{}'s code is broken into three parts:
+\begin{enumerate}
+    \item Read in configuration \\
+    \item Compute planetary spectra \\
+    \item Compute stellar spectra
+\end{enumerate}
+
+The main object that a user interacts with is the \texttt{ObservationModel}. This object can be initialized by passing it either a YAML file containing
+all of the simulation information, or a \texttt{InternalParameters} object encoded with the same information. Upon initialization a local working directory
+(by convention \texttt{.vspec}) is created to store simulation outputs and intermediate files.
+
+The second step is initiated by calling the \texttt{ObservationModel.build\_planet()} method. This is where all the interaction with PSG takes place.
+First, your 3D Global Circulation Model (GCM) is uploaded to PSG via the API's \texttt{set} method. Then, PSG is updated with information about the observation
+that does not change with time, such as instrument parameters and system information. Before entering the main loop \vspec{} computes the geometry of the system
+at each epoch, storing it for later. \vspec{} then iterates through each phase, making four calls to the PSG API:
+\begin{enumerate}
+    \item An \texttt{upd} request setting the geometric parameters for this epoch. \\
+    \item An \texttt{all} request to retrieve the combined light spectrum. \\
+    \item A second \texttt{upd} request to set the stellar flux to zero. \\
+    \item A second \texttt{all} request to retrieve the thermal emission of the planet.
+\end{enumerate}
+
+Results from these calls are stored locally in the \texttt{.fits} format.
+
+The last step is to replace the stellar spectra used by PSG with \vspec{}'s stellar model. First, this model is initialized based on the user-passed parameters.
+One consideration made during this step is that a user may want a shorter cadence here than is necessary to resolve temporal features in the planetary spectrum,
+so the geometry is recalculated and the planetary spectrum at each epoch is a linear interpolation of the results from the previous step. After this setup
+\vspec{} enters the main loop, where it computes the composite stellar spectrum based on the stellar model and spectral grid. For each epoch it returns this stellar
+spectrum as seen by the observer, the spectrum as seen by the planet, the reflected spectrum, the planetary thermal spectrum, the total combined spectrum, and the
+total noise spectrum. The data at each epoch is stored in a \texttt{.fits} file.
+
+The final outputs of \vspec{} should be easily read using Astropy's \texttt{QTable} object, but we have also included a class \texttt{PhaseAnalyzer} to automatically
+parse all output files and conveniently access 1D spectra or lightcurves.
+
+\section{Stellar Model}
+\label{sec:star}
+
+The \vspec{} star model is designed modularly to allow
 for both simple and complex behaviors. Currently, it
@@
 Once the surface coverage is calculated, a composite spectrum is computed from a grid of PHOENIX stellar spectra
-\citep{husser2013} As of \texttt{VSPEC 0.1}, we have spectra between 2300 K and 3900 K, with steps of
+\citep{husser2013}.
+As of \texttt{VSPEC 0.1} \TJ{fix!}, we have spectra between 2300 K and 3900 K, with steps of
 100 K. Each spectrum has $\log{g} = 5$ and solar metalicity.
@@
 
-{\sc VSPEC} also takes in parameters describing the population of spots, such as the mean of the (lognormal) peak area distribution
+\vspec{} also takes in parameters describing the population of spots, such as the mean of the (lognormal) peak area distribution
 \citep{bogdan1988}, the equilibrium surface coverage, and the distribution style.
@@
 
-Before simulating an observation, {\sc VSPEC} pre-computes a population of flares based on the frequency-energy relationship from \citet{gao2022}:
+Before simulating an observation, \vspec{} pre-computes a population of flares based on the frequency-energy relationship from \citet{gao2022}:
 \begin{equation} \label{eq:flare_freq}
@@
 \ref{subsec:pypsg}
-{\sc VSPEC} interfaces with PSG using the {\sc pypsg} package \TJ{link}, which is essentially a Python wrapper for the PSG API system.
+\vspec{} interfaces with PSG using the {\sc pypsg} package \TJ{link}, which is essentially a Python wrapper for the PSG API system.
 
@@
 \subsection{3D Planet models}
-{\sc VSPEC} allows the user to upload any desired GCM. However, for questions that focus on detectability,
+\vspec{} allows the user to upload any desired GCM. However, for questions that focus on detectability,
 the specific physics of the GCM are less important than properties like day/night temperature, surface pressure, or
-the major absorbers in the atmosphere. To simplify these studies, {\sc VSPEC} has a built-in GCM that takes in just a
+the major absorbers in the atmosphere. To simplify these studies, \vspec{} has a built-in GCM that takes in just a
 few parameters and produces a simple-yet-robust exoplanet climate. Simplicity allows the user to know everything about
@@
 
-The {\sc VSPEC} GCM centers around a 2D temperature map of the planet's surface. Based on \citet{cowan2011}, this map
+The \vspec{} GCM centers around a 2D temperature map of the planet's surface. Based on \citet{cowan2011}, this map
 is constructed completely given the incident flux from the star (parameterized by the host \teff and radius
@@
 Figure \ref{fig:thermal_inertia} shows a temperature map example for an Earth-like planet with $\epsilon = 2\pi$.
-These temperature maps are validated by the energy balance of the planet; {\sc VSPEC} computes the incident and emergent
+These temperature maps are validated by the energy balance of the planet; \vspec{} computes the incident and emergent
 flux across the planet and asserts that they agree.
@@
     \caption{
-        Example of a surface temperature map created by {\sc VSPEC}. The solar flux at Earth,
+        Example of a surface temperature map created by \vspec{}. The solar flux at Earth,
         a Bond albedo of 0.3, and $\epsilon = 2\pi$ were used in this case.
@@
 
-We will do three \texttt{VSPEC} simulations:
+We will do three \vspec{} simulations:
 \begin{enumerate}
@@
 contamination because of the slow rotation rate of the host star \citep[approximately 1/80$^{\text{th}}$ the orbital frequency,][]{cloutier2021}.
-In this example we use {\sc VSPEC} to demonstrate that this is a reasonable assumption.
+In this example we use \vspec{} to demonstrate that this is a reasonable assumption.
 
-We initialize {\sc VSPEC} to run a 41.0 hour observation with a cadence of 15 minutes. We use stellar and planetary parameters
+We initialize \vspec{} to run a 41.0 hour observation with a cadence of 15 minutes. We use stellar and planetary parameters
 from the NExScI Exoplanet Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/GJ1214b}}.
@@
 
-After running the {\sc VSPEC} simulations, we bin into 0.5 $\mu m$ wavelength channels and normalize each channel to the two eclipses.
+After running the \vspec{} simulations, we bin into 0.5 $\mu m$ wavelength channels and normalize each channel to the two eclipses.
 We assume the star behaves linearly between eclipse measurements and attribute any deviation to the planet.
@@
 mission concept: Proxima Centauri b with no stellar variability. We will instead examine the expected noise produced by this observation
-and assess the detectability of the planet in a single orbit. {\sc VSPEC} uses the noise models built into PSG, allowing native support for
+and assess the detectability of the planet in a single orbit. \vspec{} uses the noise models built into PSG, allowing native support for
 noise due to the instrument, source, and astrophysical background.
 
-We use \texttt{VSPEC}'s built-in MIRECLE instrument template, which includes a 2 m aperture and observes from 1-18 $\mu m$ with $R=50$.
+We use \vspec{}'s built-in MIRECLE instrument template, which includes a 2 m aperture and observes from 1-18 $\mu m$ with $R=50$.
 We draw stellar and bulk planetary parameters from the NExScI Exoplanet Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/proxima\%20cen}}

### 2024-07-19 src/tex/ms.tex
Complete VSPEC and pypsg sections

@@
 \usepackage[version=4]{mhchem}
+\usepackage{listings}
 
@@
 
-The flow of \vspec{}'s code is broken into three parts:
-\begin{enumerate}
-    \item Read in configuration \\
-    \item Compute planetary spectra \\
-    \item Compute stellar spectra
-\end{enumerate}
+The flow of \vspec{}'s code is broken into three parts: i) read in the configuration, ii) compute planetary spectra, and iii) compute stellar spectra.
+\subsection{Configuring \vspec{}}
+\label{subsec:config}
 
-The main object that a user interacts with is the \texttt{ObservationModel}. This object can be initialized by passing it either a YAML file containing
-all of the simulation information, or a \texttt{InternalParameters} object encoded with the same information. Upon initialization a local working directory
+\vspec{} configurations are designed to minimize human errors by providing two equivalent formats. The first, a file written in YAML, is optimized for
+human readability. For example, the \texttt{system} section, which describes the relationship between the observed planetary system and the observer, could
+be written:
+\begin{lstlisting}[label={ls:yaml}]
+system:
+    distance: 12.4 pc
+    inclination: 89.7 deg
+    phase_of_pariasteron: 0 deg
+\end{lstlisting}
+Note that the units of each of these parameters is included in a human-readable way. Internally, \vspec{} casts each of these parameters to an Astropy
+\texttt{Quantity} instance, so that the user can input any desired value and unit combination so long as the physical type is correct -- i.e. it would be
+equivalent to write \texttt{3.83e17 m} as the value for the \texttt{distance} parameter.
+
+The second input method available is to directly initialize the Python object that \vspec{} uses internally to store its model parameters.
+The top-level object, \texttt{InternalParameters}, is structured to mirror the YAML input file, with one argument per YAML section. The
+parameters in Listing \ref{ls:yaml} would be written:
+\begin{lstlisting}[language=Python]
+from astropy import units as u
+from VSPEC.params import *
+params = InternalParameters(
+    ... # Other arguments
+    system=SystemParameters(
+        distance=12.4*u.pc,
+        inclination=89.7*u.deg,
+        phase_of_pariasteron=0*u.deg
+    )
+)
+\end{lstlisting}
+This is convenient for producing configurations programically, for example for producing a grid of model phase curves.
+
+However the user decides to provide model parameters, they are read into the main \vspec{} object: the \texttt{ObservationModel}.
+This is the class that runs the model and handles files. Upon initialization a local working directory
 (by convention \texttt{.vspec}) is created to store simulation outputs and intermediate files.
 
-The second step is initiated by calling the \texttt{ObservationModel.build\_planet()} method. This is where all the interaction with PSG takes place.
-First, your 3D Global Circulation Model (GCM) is uploaded to PSG via the API's \texttt{set} method. Then, PSG is updated with information about the observation
-that does not change with time, such as instrument parameters and system information. Before entering the main loop \vspec{} computes the geometry of the system
-at each epoch, storing it for later. \vspec{} then iterates through each phase, making four calls to the PSG API:
-\begin{enumerate}
-    \item An \texttt{upd} request setting the geometric parameters for this epoch. \\
-    \item An \texttt{all} request to retrieve the combined light spectrum. \\
-    \item A second \texttt{upd} request to set the stellar flux to zero. \\
-    \item A second \texttt{all} request to retrieve the thermal emission of the planet.
-\end{enumerate}
+\subsection{Planetary Phase Curve}
+\label{subsec:phase-curve}
+\vspec{} generates a phase curve though a series of API calls to the Planetary Spectrum Generator \citep[PSG][]{villanueva2018}, using
+\texttt{pypsg} (see Section \ref{sec:pypsg}) to interface with the API. Initially, configurations are sent that give PSG information about
+the system that does not change with time -- that includes the 3D Global Circulation Model (GCM) and instrument parameters. \vspec{} then enters
+the main loop, iterating through each observation epoch and making a pair of API calls: one to get a combined light spectrum and one to get the thermal flux only.
+Results from PSG in each epoch are stored locally in the \texttt{.fits} format.
 
-Results from these calls are stored locally in the \texttt{.fits} format.
+\subsubsection{Phase Sampling}
+Making API calls to PSG can be computationally expensive, especially when trying to resolve small structures in the climate model. In order to reduce unnecessary
+calls to PSG the temporal sampling of the planetary spectrum in independent of the stellar model/output sampling. When computing the planetary flux in the output files
+the raw PSG output is interpolated and averaged over the integration using the trapazoid rule. By default one spectrum is computed at each interface separating integrations.
 
-The last step is to replace the stellar spectra used by PSG with \vspec{}'s stellar model. First, this model is initialized based on the user-passed parameters.
-One consideration made during this step is that a user may want a shorter cadence here than is necessary to resolve temporal features in the planetary spectrum,
-so the geometry is recalculated and the planetary spectrum at each epoch is a linear interpolation of the results from the previous step. After this setup
-\vspec{} enters the main loop, where it computes the composite stellar spectrum based on the stellar model and spectral grid. For each epoch it returns this stellar
-spectrum as seen by the observer, the spectrum as seen by the planet, the reflected spectrum, the planetary thermal spectrum, the total combined spectrum, and the
-total noise spectrum. The data at each epoch is stored in a \texttt{.fits} file.
+\subsection{Stellar Lightcurve}
 
+The last step is to replace the stellar spectra used by PSG with \vspec{}'s stellar model. In this step \vspec{} acts as a bridge between
+the stellar surface model \texttt{vspec-vsm} (which models spatial changes to the star with time, see Section \ref{sec:star}) and a grid of
+pre-computed stellar spectra (see Section \ref{sec:gridpolator}). The star is initialized and allowed to evolve for a user-specified amount of time
+in order for spots and faculae to approach growth-decay equilibrium (this is also called the ``burn in'' phase). The star is then evolved at the output cadence,
+at each epoch generating a set of key-value pairs that describe the surface temperature and coverage fractions of the portion of the stellar disk visible to the observer.
+These \teff\ s and coverage fractions are fed into GridPolator to produce composite stellar spectra.
+
+\subsubsection{Reflected light}
+\vspec{} also produces a composite spectrum of the
+portion of the star visible to the planet to accurately incorporate the variable star into the reflected light spectrum. To compute the total reflected light
+in the output files \vspec{} first divides the reflected light flux from PSG by the PSG stellar model to obtain the apparent albedo. This albedo is then multiplied
+by the composite spectrum (facing the planet) to obtain total reflected light flux.
+
+\subsubsection{Noise}
+Alongside the planetary radiance files PSG returns a file giving the noise profile for each epoch broken down by source. We assume that the total noise is the
+quadrature sum of its constituent parts and that the \texttt{source} column (i.e. photon noise) is the only one effected by replacing one stellar spectrum with the other.
+This means the photon noise can be calculated:
+\begin{equation}
+    N_{\rm photon} = N_{\rm photon,~PSG} \sqrt{\frac{F_{*,\rm ~VSPEC}}{F_{*,\rm ~PSG}}}
+\end{equation}
+where $F_*$ is the stellar flux. The total noise associated with the integration is then
+\begin{equation}
+    N_{\rm tot} = \sqrt{
+        N_{\rm photon}^2 + N_{\rm detector}^2 + N_{\rm telescope}^2 + N_{\rm background}
+    }
+\end{equation}
+
+\subsubsection{Transit and Eclipse}
+Users must be careful when using \vspec{} to simulate transit and eclipse measurements because of the vast timescale differences between those events
+and the typical phase variations of the planet -- a sparsely sampled phase curve that happens to include one epoch of transit will not describe that lightcurve accurately
+because the interpolator knows no better than to assume a linear interpolation scheme. It is recommended that users use a high cadence for observations of transits
+and eclipses so that each event will be sampled by more than one point.
+
+In the case of a transiting geometry \vspec{} computes the spectrum of the occulted portion of the star in order to properly simulate the
+transit light source effect \citep[TLS][]{rackham2018}. This spectrum is the flux blocked by the planet in the case that it is a bare rock (i.e. a flat
+transmission spectrum). The effect of atmospheric transmission is added by comparing the PSG-computed effective radius to a purely geometric
+calculation of a bare-rock transit depth. That is
+\begin{equation}
+    F_{\lambda} = F_{{\rm rock}, \lambda}\, F_{{\rm PSG,} \lambda} \,\left (\frac{R_*}{R_p} \right )^2
+\end{equation}
+where $F_\lambda$ is the flux of the variable star surface occulted by the planet with the atmosphere considered. $F_{{\rm rock}, \lambda}$ is the
+flux of the surface occulted by a bare-rock with the planet's radius, $F_{{\rm PSG,} \lambda}$ is the flux computed by PSG to be lost
+due to occultation and transmission, and $R_*$ and $R_p$ are the stellar and planetary radii, respectively.
+
+In the case of a total eclipse of the planet the thermal and reflected spectra are set to zero flux. However, in the case of a partial eclipse,
+\vspec{} computes the fraction of the planet that is visible to the observer and reduces both thermal and reflected spectra accordingly. In this case
+\vspec{} treats the planet homogeneously, however future work could utilize PSG/GlobES's ability to return a hypercube of spatially resolved
+spectra in order to construct planetary spectra only from regions of the planet that are not eclipsed.
+
+\subsection{Analysis Tools}
 The final outputs of \vspec{} should be easily read using Astropy's \texttt{QTable} object, but we have also included a class \texttt{PhaseAnalyzer} to automatically
-parse all output files and conveniently access 1D spectra or lightcurves.
+parse all output files and conveniently access 1D spectra or lightcurves. \TJ{Is it worth talking about this?}
+
+\section{Interfacing with the Planetary Spectrum Generator via \texttt{pypsg}}
+\label{sec:pypsg}
+
+PSG\footnote{\url{https://psg.gsfc.nasa.gov/}} \citep{villanueva2018} is a powerful radiative transfer tool that is ubiquitous in the exoplanet and solar system atmosphere fields.
+Either though its web graphical interface, its public API, or the local version available through Docker, PSG allows users to simulate observations of planets, comets, and
+moons with a variety of geometries and realistic noise models \TJ{probably someone like Geronimo should write this part}.
+
+During development of \vspec{} it became clear that a Python interface for the PSG API system would be generally useful to planetary scientists and astronomers, and
+so \vspec{}'s own interface was spun off into a stand-alone library: \texttt{pypsg}.
+
+\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django to encode a PSG configuration file as a native Python object.
+Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
+
+Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's
+\texttt{requests} package. This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key
+could be committed to a public repository.
+
+\subsection{GlobES}
+\label{subsec:globes}
+
+A powerful feature of \texttt{pypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
+GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial. \texttt{pypsg}'s \texttt{PyGCM} class provides
+a native Python representation of GlobES input, and includes built-in methods for converting popular climate models (e.g. ExoCAM, WACCM \TJ{full names and ref?}) to
+the correct format.
 

### 2024-07-19 src/tex/ms.tex
Changes from Avi/Ted editiing session on July 19 2024

@@
 
-In the era of high-sensitivity exoplanet characterization missions such as the James Webb Space Telescope (JWST)
-and the Atmospheric Remote-sensing Infrared Exoplanet Large-survey (ARIEL), spectral analysis of exoplanet atmospheres
-is increasingly sensitive to contamination due to stellar inhomogeneities (e.g. spots, granulation) and stochastic stellar variability (e.g. flares).
+In the era of high-sensitivity transiting exoplanet characterization missions such as the Hubble Space Telescope (HST), the James Webb Space Telescope (JWST) and the future Atmospheric Remote-sensing Infrared Exoplanet Large-survey (ARIEL), spectral analysis of exoplanet atmospheres is increasingly sensitive to contamination due to stellar inhomogeneities (e.g.
+spots, granulation) and stochastic stellar variability (e.g.
+flares).
+As an example, recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between atmospheric absorption by water and water-rich spots on the stellar surface, and similar effects were seen in the transit spectrum of LHS 1140b \citet{Cadieux2024}.
+This ``transit light source effect'' \citep[TLS,][see also \citet{apai2018,barclay2021,garcia2022,barclay2023}]{rackham2018} occurs when the region of the stellar surface occulted by the transiting planet is not representative of the disk-integrated spectrum.
 
-Recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between
-atmospheric absorption by water and water-rich spots on the stellar surface. This ``transit light source effect''
-\citep[TLS,][see also \citet{apai2018,barclay2021,garcia2022,barclay2023}]{rackham2018}
-occurs when the occulted region of the stellar surface is not representative of the disk-integrated spectrum.
-
-Similarly, the Mid-IR Exoplanet CLimate Explorer \citep[MIRECLE,][]{mandell2022} mission concept
-will employ the Planetary Infrared Excess \citep[PIE][]{stevenson2020} technique to extract the planetary
-contribution from combined-light observations. Uncertainties on the stellar spectrum will dominate the analysis if it is not removed appropriately.
+Stellar contamination may also cause errors in the extraction of thermal phase curve spectroscopy if the stellar spectrum changes significantly over the period of the planet's orbit; if the variations are non-linear but only linear interpolation between subsequent secondary eclipses is used to remove the stellar contribution, the results planetary flux measurements will be contaminated.
+This will be particularly important for observations of planets around more variable stars such as active M-stars or later-type evolved host stars.
+%the Mid-IR Exoplanet CLimate Explorer \citep[MIRECLE,][]{mandell2022} mission concept will employ the Planetary Infrared Excess \citep[PIE][]{stevenson2020} technique to extract the planetary contribution from combined-light observations.
+Uncertainties on the stellar spectrum will dominate the analysis if it is not removed appropriately.
 
@@
 This task requires a flexible tool that combines models of exoplanet atmospheres and stellar variability in a robust way.
-
 In this paper we present \vspec{}: Variable Star PhasE Curve\footnote{\url{https://github.com/VSPEC-collab/VSPEC}},
@@
 \end{lstlisting}
-This is convenient for producing configurations programically, for example for producing a grid of model phase curves.
+This is convenient for producing configurations programmatically, for example for producing a grid of model phase curves.
 
@@
 the system that does not change with time -- that includes the 3D Global Circulation Model (GCM) and instrument parameters. \vspec{} then enters
-the main loop, iterating through each observation epoch and making a pair of API calls: one to get a combined light spectrum and one to get the thermal flux only.
+the main loop, iterating through each observation epoch (e.g.integration or combination of integrations) and making a pair of API calls: one to get a combined light spectrum and one to get the thermal flux only.
 Results from PSG in each epoch are stored locally in the \texttt{.fits} format.
@@
 \subsubsection{Phase Sampling}
-Making API calls to PSG can be computationally expensive, especially when trying to resolve small structures in the climate model. In order to reduce unnecessary
-calls to PSG the temporal sampling of the planetary spectrum in independent of the stellar model/output sampling. When computing the planetary flux in the output files
-the raw PSG output is interpolated and averaged over the integration using the trapazoid rule. By default one spectrum is computed at each interface separating integrations.
+Making API calls to PSG can be computationally expensive, especially when trying to resolve small structures in the climate model.
+In order to reduce unnecessary calls to PSG, the temporal sampling of the planetary spectrum is independent of the stellar model/output sampling.
+When computing the planetary flux in the output files, the raw PSG output is interpolated and averaged over each time step using the trapezoid integration rule.
+By default one spectrum is computed with PSG at each interface separating time steps; this eases the calculation of edge effects between steps.
 
@@
 the stellar surface model \texttt{vspec-vsm} (which models spatial changes to the star with time, see Section \ref{sec:star}) and a grid of
-pre-computed stellar spectra (see Section \ref{sec:gridpolator}). The star is initialized and allowed to evolve for a user-specified amount of time
-in order for spots and faculae to approach growth-decay equilibrium (this is also called the ``burn in'' phase). The star is then evolved at the output cadence,
-at each epoch generating a set of key-value pairs that describe the surface temperature and coverage fractions of the portion of the stellar disk visible to the observer.
+pre-computed stellar spectra (from the GridPolator package; see Section \ref{sec:gridpolator}).
+The star is initialized and allowed to evolve for a user-specified amount of time in order for spots and faculae to approach growth-decay equilibrium (this is also called the ``burn in'' phase).
+The star is then evolved at the output cadence, at each epoch generating a set of key-value pairs that describe the surface temperature and coverage fractions of the portion of the stellar disk visible to the observer.
 These \teff\ s and coverage fractions are fed into GridPolator to produce composite stellar spectra.
@@
 \subsubsection{Reflected light}
-\vspec{} also produces a composite spectrum of the
-portion of the star visible to the planet to accurately incorporate the variable star into the reflected light spectrum. To compute the total reflected light
-in the output files \vspec{} first divides the reflected light flux from PSG by the PSG stellar model to obtain the apparent albedo. This albedo is then multiplied
+\vspec{} also produces a composite spectrum of the portion of the star visible to the planet to accurately incorporate the variable star into the reflected light spectrum.
+To compute the total reflected light in the output files, \vspec{} first divides the reflected light flux from PSG by the PSG stellar model to obtain the apparent albedo (I/F).
+This albedo is then multiplied
 by the composite spectrum (facing the planet) to obtain total reflected light flux.
@@
 \subsubsection{Noise}
-Alongside the planetary radiance files PSG returns a file giving the noise profile for each epoch broken down by source. We assume that the total noise is the
-quadrature sum of its constituent parts and that the \texttt{source} column (i.e. photon noise) is the only one effected by replacing one stellar spectrum with the other.
+Alongside the planetary radiance files, PSG returns a file giving the noise profile for each epoch broken down by source.
+We assume that the total noise is the quadrature sum of its constituent parts and that the \texttt{source} column (i.e.
+photon noise) is the only one affected by replacing one stellar spectrum with the other.
 This means the photon noise can be calculated:
@@
 Users must be careful when using \vspec{} to simulate transit and eclipse measurements because of the vast timescale differences between those events
-and the typical phase variations of the planet -- a sparsely sampled phase curve that happens to include one epoch of transit will not describe that lightcurve accurately
-because the interpolator knows no better than to assume a linear interpolation scheme. It is recommended that users use a high cadence for observations of transits
-and eclipses so that each event will be sampled by more than one point.
+and the typical phase variations of a planet -- a sparsely sampled phase curve that happens to include one epoch of transit will not adequately sample the lightcurve accurately during a transit or eclipse because the interpolator knows no better than to assume a linear interpolation scheme.
+It is recommended that users use a high cadence for observations of transits and eclipses so that each event will be sampled by more than one point.
 
-In the case of a transiting geometry \vspec{} computes the spectrum of the occulted portion of the star in order to properly simulate the
-transit light source effect \citep[TLS][]{rackham2018}. This spectrum is the flux blocked by the planet in the case that it is a bare rock (i.e. a flat
-transmission spectrum). The effect of atmospheric transmission is added by comparing the PSG-computed effective radius to a purely geometric
-calculation of a bare-rock transit depth. That is
+In the case of a transiting geometry, \vspec{} computes the spectrum of the occulted portion of the star in order to properly simulate the
+transit light source effect \citep[TLS][]{rackham2018}.
+This spectrum is the flux blocked by the planet in the case that it is a solid occulting circle, and therefore produces a flat transmission spectrum.
+The effect of atmospheric transmission is added by comparing the PSG-computed effective radius to a purely geometric calculation of such a ``bare-rock" transit depth:
 \begin{equation}
@@
 where $F_\lambda$ is the flux of the variable star surface occulted by the planet with the atmosphere considered. $F_{{\rm rock}, \lambda}$ is the
-flux of the surface occulted by a bare-rock with the planet's radius, $F_{{\rm PSG,} \lambda}$ is the flux computed by PSG to be lost
+flux of the surface occulted by a bare rock with the planet's radius, $F_{{\rm PSG,} \lambda}$ is the flux computed by PSG to be lost
 due to occultation and transmission, and $R_*$ and $R_p$ are the stellar and planetary radii, respectively.
 
-In the case of a total eclipse of the planet the thermal and reflected spectra are set to zero flux. However, in the case of a partial eclipse,
-\vspec{} computes the fraction of the planet that is visible to the observer and reduces both thermal and reflected spectra accordingly. In this case
-\vspec{} treats the planet homogeneously, however future work could utilize PSG/GlobES's ability to return a hypercube of spatially resolved
+In the case of a total eclipse of the planet, the planetary thermal and reflected contributions are set to zero flux. However, in the case of a partial eclipse,
+\vspec{} computes the fraction of the planet that is visible to the observer and reduces both thermal and reflected flux accordingly. In this case
+\vspec{} treats the planet homogeneously; however, future work could utilize PSG/GlobES's ability to return a hypercube of spatially resolved
 spectra in order to construct planetary spectra only from regions of the planet that are not eclipsed.
-
-\subsection{Analysis Tools}
-The final outputs of \vspec{} should be easily read using Astropy's \texttt{QTable} object, but we have also included a class \texttt{PhaseAnalyzer} to automatically
-parse all output files and conveniently access 1D spectra or lightcurves. \TJ{Is it worth talking about this?}
 
@@
 PSG\footnote{\url{https://psg.gsfc.nasa.gov/}} \citep{villanueva2018} is a powerful radiative transfer tool that is ubiquitous in the exoplanet and solar system atmosphere fields.
-Either though its web graphical interface, its public API, or the local version available through Docker, PSG allows users to simulate observations of planets, comets, and
-moons with a variety of geometries and realistic noise models \TJ{probably someone like Geronimo should write this part}.
+Either through its web graphical interface, its public API, or the local version available through Docker, PSG allows users to simulate observations of planets, comets, and moons with a variety of geometries and realistic noise models \TJ{probably someone like Geronimo should write this part}.
 
-During development of \vspec{} it became clear that a Python interface for the PSG API system would be generally useful to planetary scientists and astronomers, and
-so \vspec{}'s own interface was spun off into a stand-alone library: \texttt{pypsg}.
-
-\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django to encode a PSG configuration file as a native Python object.
-Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
-
-Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's
-\texttt{requests} package. This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key
+\vspec{} uses a Python interface for the PSG API system called \texttt{pypsg}, which is a general stand-alone library built for any Python-based API call to PSG.
+\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django \TJ{cite} to encode a PSG configuration file as a native Python object.Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
+Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests} \TJ{cite} package.
+This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key
 could be committed to a public repository.
 
-\subsection{GlobES}
-\label{subsec:globes}
-
 A powerful feature of \texttt{pypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
-GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial. \texttt{pypsg}'s \texttt{PyGCM} class provides
-a native Python representation of GlobES input, and includes built-in methods for converting popular climate models (e.g. ExoCAM, WACCM \TJ{full names and ref?}) to
-the correct format.
+GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial.
+\texttt{pypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models (e.g.
+ExoCAM, WACCM \TJ{full names and ref?}) to the correct format.
 

### 2024-07-19 src/scripts/moran23.py
Set API key with a script at build time

@@ -57,6 +57,7 @@ psg_kwargs = dict(
     phase_binning=1,
     nmax=0,
     lmax=0,
+    use_continuum_stellar=True,
     continuum=['Rayleigh', 'Refraction'],
 )
 psg_params = params.psgParameters(

### 2024-07-19 src/scripts/granule_lc.py
Add missing stellar signatures param

@@ -64,6 +64,7 @@ psg_params = params.psgParameters(
     gcm_binning=200,
     phase_binning=1,
     use_molecular_signatures=True,
+    use_continuum_stellar=True,
     nmax=0,
     lmax=0,
     continuum=['Rayleigh', 'Refraction', 'CIA_all'],

### 2024-07-19 src/scripts/kempton23.py

@@ -29,6 +29,7 @@ observation = params.ObservationParameters(
 psg_params = params.psgParameters(
     gcm_binning=200,
     phase_binning=1,
+    use_continuum_stellar=True,
     nmax=0,
     lmax=0,
     continuum=['Rayleigh', 'Refraction', 'CIA_all'],

### 2024-07-19 src/scripts/mirecle.py

@@ -32,6 +32,7 @@ observation = params.ObservationParameters(
 # PSG
 psg_params = params.psgParameters(
     use_molecular_signatures=True,
+    use_continuum_stellar=True,
     gcm_binning=200,
     phase_binning=1,
     nmax=0,

### 2024-07-19 src/tex/ms.tex
Add Cadieux reference

@@
 flares).
-As an example, recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between atmospheric absorption by water and water-rich spots on the stellar surface, and similar effects were seen in the transit spectrum of LHS 1140b \citet{Cadieux2024}.
+As an example, recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between atmospheric absorption by water and water-rich spots on the stellar surface, and similar effects were seen in the transit spectrum of LHS 1140b \citet{cadieux2024}.
 This ``transit light source effect'' \citep[TLS,][see also \citet{apai2018,barclay2021,garcia2022,barclay2023}]{rackham2018} occurs when the region of the stellar surface occulted by the transiting planet is not representative of the disk-integrated spectrum.

### 2024-07-19 src/scripts/kempton23.py
Fix changed variable names in scripts

@@ -176,8 +176,8 @@ data_spotted = PhaseAnalyzer(model_spotted.directories['all_model'])
 data_quiet = PhaseAnalyzer(model_quiet.directories['all_model'])
 flux_unit = u.Unit('W m-2 um-1')
 def get_star(data:PhaseAnalyzer):
-    i_eclipse1 = np.argmin(data.lightcurve('total',(0,-1))[:data.N_images//4])
-    i_eclipse2 = np.argmin(data.lightcurve('total',(0,-1))[3*data.N_images//4:]) + 3*data.N_images//4
+    i_eclipse1 = np.argmin(data.lightcurve('total',(0,-1))[:data.n_images//4])
+    i_eclipse2 = np.argmin(data.lightcurve('total',(0,-1))[3*data.n_images//4:]) + 3*data.n_images//4
     time = (data.time-data.time[0]).to_value(u.hr)
     star_spec1 = data.spectrum('total',i_eclipse1).to_value(flux_unit)
     star_spec2 = data.spectrum('total',i_eclipse2).to_value(flux_unit)

### 2024-07-19 src/scripts/surface_map_and_lc.py

@@ -35,8 +35,8 @@ ax2 = fig.add_subplot(2,1,2)
 CFG_PATH = Path(__file__).parent / 'spot_lightcurve.yaml'
 
 model = ObservationModel.from_yaml(CFG_PATH)
-model.build_star()
-model.warm_up_star(facula_warmup_time=10*u.hr)
+model._build_star()
+model._warm_up_star(facula_warmup_time=10*u.hr)
 
 
 

### 2024-07-21 src/tex/ms.tex
Revise vspec-vsm section

@@
 
-\section{Stellar Model}
+\section{Stellar Variability Model}
 \label{sec:star}
 
-The \vspec{} star model is designed modularly to allow
-for both simple and complex behaviors. Currently, it
-is represented by a rectangular grid of points on the stellar surface,
-each assigned an effective temperature. At any given time, the model computes
-the surface coverage fractions of each temperature visible to the observer, accounting for the
-spherical geometry, limb darkening, and any occultation by a transiting planet.
+The \vspec{} stellar model is designed modularly to allow for both simple and complex behaviors.
+The model is available as the Python package \texttt{vspec-vsm} \TJ{(VSPEC Variable Star Model)}.
+It is spectral-model agnostic, meaning that it relies on another package (GridPolator, see Section \ref{sec:gridpolator}) to compute the actual variable spectra.
+Instead, the surface model takes a pair of latitude and longitude points as input and returns a dictionary $\{ T_{\rm eff} : f_{T_{\rm eff}} \}$ where the keys \teff~ are the effective temperatures of each surface component and the values $f_{T_{\rm eff}}$ are the fraction of the visible stellar disk with that effective temperature.
 
-The attributes of the \texttt{Star} class describe the bulk properties of the star, including radius,
-period, and the effective temperature of quiet photosphere. Herein we refer to this temperature as the
-photosphere temperature to differentiate it from the temperature of spots, faculae, or other sources of variability.
+The main model class is the \texttt{Star} object, which describes bulk properties like radius, nominal \teff, and limb darkening parameters, in addition to acting as 
+a container for objects representing sources of variability.
+
+\subsection{Stellar Surface Grid}
+Sources of variability that exist as major structures on the stellar surface (i.e.
+spots and faculae) are encoded as deviations from the nominal effective temperature on a grid of points representing the stellar surface.
+Each time the fractions of each surface component are calculated, the grid is initialized with constant \teff.
+The code then iterates through each spot and facula, replacing portions of the grid with the effective temperatures that make up each structure.
+
+The simplest grid to build is one that is rectangular in latitude and longitude, and in which the area that each point represents decreases towards the poles.
+However, this can lead to unnecessary spatial oversampling at the poles, and may lead to poor resolution at the equator to keep the total number of points from being excessively high.
+To mitigate these problems \texttt{vspec-vsm} supports a ``grid'' generated using a Fibbonacci spiral to create a near-isotropic distribution of points.
+This is especially useful when simulating transits, giving optimal sampling at the stellar equator.
+It is also useful for resolving small structures on the surface, as the number of points needed for a given minimum resolution are greatly reduced.
 
@@
     \script{surface_map_and_lc.py}
-    \caption{{\bf Top}: The surface map of a star with spots and faculae during a transit.
-    The large circles are spots. Faculae display their 3D structure, as the hot wall is more prevalent near the limbs.
+    \caption{{\bf Top}: The surface map of a star with spots during a transit.
+    The large elipses are spots, while the transiting planet occults a portion of the disk represented by the black circle.
     {\bf Bottom}: An example lightcurve of a spotted star.}
@@
 \end{figure}
-
-\subsection{Stellar Spectra \label{subsec:spectra}}
-Once the surface coverage is calculated, a composite spectrum is computed from a grid of PHOENIX stellar spectra
-\citep{husser2013}.
-As of \texttt{VSPEC 0.1} \TJ{fix!}, we have spectra between 2300 K and 3900 K, with steps of
-100 K. Each spectrum has $\log{g} = 5$ and solar metalicity.
-The raw spectra span from 0.1 to 20 \micron~ with $5 \times 10^{-6}$ \micron~ steps, however binned
-versions are pre-computed for faster runtimes.
 
@@
 
-On the Sun, spots have two regions: the dark, central umbra and the lighter, surrounding penumbra. A detailed review of sunspot behavior,
+On the Sun, spots have two regions shown if Figure \ref{fig:surface_map}: the dark, central umbra and the lighter, surrounding penumbra. A detailed review of sunspot behavior,
 including sizes and lifetimes, is \citet{solanki2003}, however the specific solar values are not relevant to this description. However,
@@
 
-\vspec{} also takes in parameters describing the population of spots, such as the mean of the (lognormal) peak area distribution
-\citep{bogdan1988}, the equilibrium surface coverage, and the distribution style.
+In addition to modeling the variability produced by a population of spots, \texttt{vspec-vsm} contains utilities to produce those populations given some description.
+These parameters include the average spot area \citep[lognormally distributed, ][]{bogdan1988}, \teff, growth and decay rates, and the method for distributing spots on the surface.
+These \texttt{SpotGenerator} objects are responsible for evolving the spot population as the star ages.
+The number of spots created in some interval $\Delta t$ is:
+\begin{equation}
+    N(\Delta t) = \frac{4 \pi R_*^2 f_{\rm spot}}{A_{\rm mean}} \frac{\Delta t}{T}
+\end{equation}
+where $R_*$ is the stellar radius, $f_{\rm spot}$ is the fraction of the surface covered by spots in growth-decay equilibrium, $A_{\rm mean}$ is the lifetime-averaged mean spot area, and $T$ is the spot lifetime, defined as:
+\begin{equation}
+    T = \frac{A_0}{W} - \frac{1}{\tau} \ln{\left (\frac{A_{\rm min}}{A_0}\right )}
+\end{equation}
+where $A_{\rm min}$ is the area given to each spot when it is initialized.
+
+There are also two modes to distribute spots on the star's surface.
+The first distributes them isotropically.
+The second, modeled after analysis of spots on the Sun by \citet{mandal2017}, concentrates the spot latitude around $\pm 15^{\circ}$.
 
@@
 
-Their three-dimensional structure causes faculae to behave differently depending on their angle from disk-center. Close to the limb,
+Their three-dimensional structure causes faculae's observational properties to change depending on their angle from disk-center. Close to the limb,
 the hot wall is visible to the observer, and faculae appear as bright points. Near the center, however, the cool floor is exposed and
@@
 
-According to \citet{1997ApJ...484..479T}, faculae temperatures (of both the floor and wall) are dependent on the facula radius,
+According to \citet{topka1997}, faculae temperatures (of both the floor and wall) are dependent on the facula radius,
 while depth appears to be constant. They also find that the smallest faculae have no visible floor, and that even at disk center they
@@
 \subsection{Flares \label{subsec:flares}}
-Flares are an important source of stellar variability on short timescales. We use the {\sc xoflares}
-package \citep{barclay2020} to generate realistic lightcurve shapes and assume a blackbody spectrum.
-{\sc xoflares} uses a flare profile function empirically derived by \citet{davenport2016} to
-describe a flare lightcurve. Temperature, $E$, FWHM, and the peak time together completely describe a flare's multiwavelength lightcurve.
+Flares are an important source of stellar variability on short timescales.
+\texttt{vspec-vsm}'s flare lightcurve model is based on the \texttt{xoflares} package \citep{barclay2020}, which itself is based on empirically derived lightcurve shapes from \citet{davenport2016}.
+However, \texttt{xoflares} is designed to fit lightcurves in a single spectral band, so we add additional parameters to extend it to a multiwavelength lightcurve.
+We completely describe a flare using its temperature, total energy, full width at half maximum (FWHM), and the time of its peak.
+We model a flare as a hot, optically thin region above the photosphere that produces a blackbody spectrum.
+In this model, the temperature is constant, and the sharp rise and fall see in the lightcurve is caused by the region's rapidly changing area.
+This simplification allows the total flare energy $E$ to be used as a normalization factor:
+\begin{equation}
+    \int_{-\infty}^{\infty}A\,dt = \frac{E}{\sigma T^4}
+\end{equation}
+where $A$ is the time-dependent area of the flare region, $T$ is the constant flare temperature, and $\sigma$ is the Stefan-Boltzmann constant.
+This allows flare lightcurves to be produced with a fixed bolometric luminosity.
 
-Before simulating an observation, \vspec{} pre-computes a population of flares based on the frequency-energy relationship from \citet{gao2022}:
+Because the temperature is fixed, the relevant quantity for modeling the flux of a flare in a given integration is the integrated time-area $\int A\,dt$.
+These quantities are computed numerically and fed to \vspec{} along with $T$ to produce spectra with the appropriate absolute flux.
+
+Before simulating an observation, \vspec{} asks \texttt{vspec-vsm} to pre-compute a population of flares based on the frequency-energy relationship from \citet{gao2022}:
 \begin{equation} \label{eq:flare_freq}
-    \log{(f{\rm [day]})} = \beta + \alpha~\log{(E/{\rm [erg]})}
+    \log{(f/{\rm [day]})} = \beta + \alpha~\log{(E/{\rm [erg]})}
 \end{equation}
@@
     \label{fig:gran_lc}
-a\end{figure}
+\end{figure}
 
-\section{Planet Model \label{sec:pl_model}}
-Our model uses the Planetary Spectrum Generator\footnote{\href{https://psg.gsfc.nasa.gov/}{https://psg.gsfc.nasa.gov/}}
-\citep[PSG,][]{villanueva2018} to calculate the planetary spectrum. We upload a 3D Global Circulation Model (GCM) to PSG's
-Global Emission Spectra (GlobES) application to retrieve thermal and reflected spectra for various points in the planet's phase curve. \TJ{cite Vincent's paper}
+\subsection{Transits}
+In the case of a transiting geometry, \texttt{vspec-vsm} is responsible for computing the properties of the portion of the stellar disk that is occulted.
+It does this by first projecting both the transiting planet and the visible part of the surface orthographically as they would be viewed by a distant observer.
+It then selects all the points that are near enough to the transit to be occulted.
+The code then iterates through each of those points and computes the fraction of it covered by the occulting circle using a 2D numerical integral.
 
-\subsection{Interfacing through {\sc pypsg}}
-\ref{subsec:pypsg}
-\vspec{} interfaces with PSG using the {\sc pypsg} package \TJ{link}, which is essentially a Python wrapper for the PSG API system.
+\section{Stellar Spectral Grids}
+\label{sec:gridpolator}
 
-\TJ{Write much more}
+\TJ{TODO}
 
-\subsection{3D Planet models}
+\section{Simple climate models}
 \vspec{} allows the user to upload any desired GCM. However, for questions that focus on detectability,

### 2024-07-22 src/tex/ms.tex
Changes from Avi/Ted editing on July 22 2024

@@
 
-The \vspec{} stellar model is designed modularly to allow for both simple and complex behaviors.
-The model is available as the Python package \texttt{vspec-vsm} \TJ{(VSPEC Variable Star Model)}.
+The \vspec{} stellar model is designed in a modular fashion to allow for both simple and complex behaviors.
+The model is available as the Python package \texttt{vspec-vsm}.
 It is spectral-model agnostic, meaning that it relies on another package (GridPolator, see Section \ref{sec:gridpolator}) to compute the actual variable spectra.
@@
 
-The main model class is the \texttt{Star} object, which describes bulk properties like radius, nominal \teff, and limb darkening parameters, in addition to acting as 
-a container for objects representing sources of variability.
+The main model class is the \texttt{Star} object, which describes bulk properties (e.g. radius, photospheric \teff, limb darkening parameters) in addition to acting as 
+a container for objects representing sources of variability. \TJ{Additional classes are defined for additional stellar components, e.g. spots, faculae,..., which are discussed below.}
 
@@
 Sources of variability that exist as major structures on the stellar surface (i.e.
-spots and faculae) are encoded as deviations from the nominal effective temperature on a grid of points representing the stellar surface.
+spots and faculae) are encoded as deviations from the photospheric effective temperature on a grid of points representing the stellar surface.
 Each time the fractions of each surface component are calculated, the grid is initialized with constant \teff.
@@
 This is especially useful when simulating transits, giving optimal sampling at the stellar equator.
-It is also useful for resolving small structures on the surface, as the number of points needed for a given minimum resolution are greatly reduced.
+It is also useful for resolving small structures on the surface, as the number of points needed for a given minimum resolution is greatly reduced.
 
@@
     \caption{{\bf Top}: The surface map of a star with spots during a transit.
-    The large elipses are spots, while the transiting planet occults a portion of the disk represented by the black circle.
-    {\bf Bottom}: An example lightcurve of a spotted star.}
+    The large ellipses are spots, while the transiting planet occults a portion of the disk represented by the black circle.
+    {\bf Bottom}: An example lightcurve of a spotted star.
+    \TJ{Remove transit and put it in a LD figure}}
     \label{fig:surface_map}
@@
 \subsection{Spots \label{subsec:spots}}
-Our star spot model is nearly entirely based on observations of the sun. Sun spots can be resolved and are well-studied,
-whereas spots on other stars (e.g., M dwarfs) can only be observed indirectly. We therefore designed our spot model to mimic
-sun spots but with parameterized values for spot temperature and lifetime that can be matched to observations of other stellar types and ages.
-
-On the Sun, spots have two regions shown if Figure \ref{fig:surface_map}: the dark, central umbra and the lighter, surrounding penumbra. A detailed review of sunspot behavior,
-including sizes and lifetimes, is \citet{solanki2003}, however the specific solar values are not relevant to this description. However,
-the spot area as a function of time is
+Our star spot model is nearly entirely based on observations of the Sun.
+Sunspots can be resolved and are well-studied, whereas spots on other stars (especially non-solar-type stars) can only be observed indirectly.
+We therefore designed our spot model to mimic the behavior of sunspots but with parameterized values for spot temperature and lifetime that can be matched to observations of other stellar types and ages.
+On the Sun, spots have two regions shown in Figure \ref{fig:surface_map}: the dark, central umbra and the lighter, surrounding penumbra (a detailed review of sunspot behavior, including sizes and lifetimes, is described in \citet{solanki2003}).
+The spot area as a function of time is
 \begin{equation}
@@
 where $A_{\rm min}$ is the area given to each spot when it is initialized.
-
 There are also two modes to distribute spots on the star's surface.
-The first distributes them isotropically.
-The second, modeled after analysis of spots on the Sun by \citet{mandal2017}, concentrates the spot latitude around $\pm 15^{\circ}$.
+The first distributes them isotropically; the second, modeled after analysis of spots on the Sun by \citet{mandal2017}, concentrates the spot latitude around $\pm 15^{\circ}$.
 
@@
 Faculae are magnetically-generated regions of the solar surface that usually appear as bright points near the limb; we employ the ``hot wall''
-model \citep{spruit1976} where faculae are described as three-dimensional pores in the stellar surface with a hot, bright wall, and a
-cool, dark floor, as shown in Figure \ref{fig:fac_struct}.
+model \citep{spruit1976} where faculae are described as three-dimensional pores in the stellar surface with a hot, bright wall and a cool, dark floor, as shown in Figure \ref{fig:fac_struct}.
+Their three-dimensional structure causes faculaes' observational properties to change depending on their angle from disk-center.
+Close to the limb,
+the hot wall is visible to the observer, and faculae appear as bright points; near the center, however, the cool floor is exposed and faculae appear dark.
+To consider this effect in the faculae lightcurve, we compute the fraction of each facula's projected area -- the area on the disk it would occupy as seen by the observer -- that is occupied by the hot wall versus the cool floor.
+This is done via numerical integral along the radius of the spot.
 
-Their three-dimensional structure causes faculae's observational properties to change depending on their angle from disk-center. Close to the limb,
-the hot wall is visible to the observer, and faculae appear as bright points. Near the center, however, the cool floor is exposed and
-faculae appear dark. To consider this effect in the faculae lightcurve, we compute the fraction of the facula's normalized area -- the
-area on the disk it would occupy as a flat spot -- that is occupied by each the hot wall and cool floor. This is done via numerical integral
-along the radius of the spot.
-
+We calculate each portion using geometrical relationships between the spot dimensions and the viewing angle:
 \begin{equation}
@@
 \end{equation}
-for facula radius $R$, depth $Z_w$, and angle from disk-center $\alpha$. $r$ in this numerical scheme
-is defined as the distance from the center of the facula along the radial line connecting the facula center to the
-disk center. $\alpha_{\rm crit}$ is the value of alpha at which the floor is no longer visible and is defined to be $\arctan{\frac{2\sqrt{R^2-r^2}}{Z_w}}$.
+for facula radius $R$, depth $Z_w$, and angle from disk-center $\alpha$.
+$r$ in this numerical scheme is defined as the distance from the center of the facula along the radial line connecting the facula center to the disk center.
+$\alpha_{\rm crit}$ is the value of alpha at which the floor is no longer visible and is defined to be $\arctan{\frac{2\sqrt{R^2-r^2}}{Z_w}}$.
 
-According to \citet{topka1997}, faculae temperatures (of both the floor and wall) are dependent on the facula radius,
-while depth appears to be constant. They also find that the smallest faculae have no visible floor, and that even at disk center they
-appear as bright points. We parameterize the floor temperature to be
+According to studies of solar faculae \citep{topka1997}, faculae temperatures (of both the floor and wall) are dependent on the facula radius, while depth appears to be constant.
+They also find that the smallest faculae have no visible floor, and that even at disk center they appear as bright points.
+We parameterize the floor temperature to be
 \begin{equation}
@@
 \end{equation}
-
-Where $r$ is the radius of the facula, $r_{\rm min}$ is the minimum radius where the floor is visible, $\Delta  T_{\rm eff, floor, 0}$ is
-the difference between the floor and photosphere \teff at $r_{\rm min}$, and $m_{\rm floor}$ is the slope of the relationship with
-units of [temperature] [length]$^{-1}$.
-
-Similarly, the wall temperature is parameterized
-
+\TJ{Change variable names, make diagram}
+where $r$ is the radius of the facula, $r_{\rm min}$ is the minimum radius where the floor is visible, $\Delta  T_{\rm eff, floor, 0}$ is the difference between the floor and photosphere \teff at $r_{\rm min}$, and $m_{\rm floor}$ is the slope of the relationship with units of [temperature] [length]$^{-1}$.
+Similarly, the wall temperature is parameterized as
 \begin{equation}
@@
 \end{equation}
-
-Where $\Delta T_{\rm eff,wall,0}$ is the temperature of a zero-radius facula and and $m_{\rm wall}$ is the slope of the radius-temperature
-relationship with units of [temperature] [length]$^{-1}$.
+where $\Delta T_{\rm eff,wall,0}$ is the temperature of a zero-radius facula and $m_{\rm wall}$ is the slope of the radius-temperature relationship with units of [temperature] [length]$^{-1}$.
 These relationships can be defined by the user, for example setting $m=0$ for constant temperatures.
 
-Facula lifetimes are defined as the time it takes to decay by $e^{-2}$. Because faculae grow and decay exponentially
-at the same rate, each facula spends 1 lifetime with a radius greater than $e^{-1}$ of it's maximum. Each facula is born
-and dies at $e^{-2}$ of its maximum, effectively existing in the code for two lifetimes.
-
-\citet{hovis-afflerbach2022} suggest a typical facula lifetime to be on the order of 6 hours,
-with a distribution that resembles a Poisson function. However, we choose a lognormal distribution
-for both lifetime and maximum radius because it does not allow for these values to be 0. We choose to
-correlate lifetime and radius so that they are determined by the normally distributed random variable $\mu$;
-for each new facula, $\mu$ is randomly drawn and determines both the faclue lifetime and maximum radius.
+Facula lifetimes are defined as the time it takes its radius to decay by $e^{-2}$.
+Because faculae grow and decay exponentially at the same rate, each facula spends one lifetime with a radius greater than $e^{-1}$ of its maximum.
+Each facula is born and dies at a radius of $e^{-2}$ of its maximum, effectively existing in the code for two lifetimes.
+\citet{hovis-afflerbach2022} suggest a typical facula lifetime to be on the order of 6 hours, with a distribution that resembles a Poisson function; however, we choose a lognormal distribution for both lifetime and maximum radius because it does not allow for these values to be 0.
+We choose to correlate lifetime and radius so that they are determined by a normalized random distribution - for each new facula, a value is randomly drawn and determines both the facula lifetime and maximum radius.
 
@@
     \caption{
-        ``Hot Wall'' model of faculae. Faculae structure causes their contrast to be dependent on their distance
-        from the center of the disk. {\bf Top}: Depiction of a facula on the limb of a star. The hot wall is exposed
-        to the observer causing the pore to appear bright. At disk center, the cool floor is most visible. {\bf Bottom}:
-        The effects of depression depth and viewing angle on facula brightness. The 3D structure of faculae is most apparent
-        when radius $\sim$ depth. A toy flux model was used to demonstrate the shape of these curves, but their magnitudes
-        in practice depend on stellar spectral models.
+        ``Hot Wall'' model of faculae.
+Faculae structure causes their contrast to be dependent on their distance from the center of the disk.
+{\bf Top}: Depiction of a facula on the limb of a star.
+The hot wall is exposed to the observer causing the pore to appear bright.
+At disk center, the cool floor is most visible.
+{\bf Bottom}: The effects of depression depth and viewing angle on facula brightness.
+The 3D structure of faculae is most apparent when radius $\sim$ depth.
+A toy flux model was used to demonstrate the shape of these curves, but in practice their magnitudes depend on stellar spectral models.
         }
@@
 We model a flare as a hot, optically thin region above the photosphere that produces a blackbody spectrum.
-In this model, the temperature is constant, and the sharp rise and fall see in the lightcurve is caused by the region's rapidly changing area.
+In this model, the temperature is constant, and the sharp rise and fall seen in the lightcurve is caused by the region's rapidly changing area.
 This simplification allows the total flare energy $E$ to be used as a normalization factor:
@@
 This allows flare lightcurves to be produced with a fixed bolometric luminosity.
-
 Because the temperature is fixed, the relevant quantity for modeling the flux of a flare in a given integration is the integrated time-area $\int A\,dt$.
@@
 
-Before simulating an observation, \vspec{} asks \texttt{vspec-vsm} to pre-compute a population of flares based on the frequency-energy relationship from \citet{gao2022}:
+Before simulating an observation, \vspec{} asks \texttt{vspec-vsm} to pre-compute a population of flares based on a power-law frequency-energy relationship similar to the results of \TJ{what data?} \citep{gao2022}:
 \begin{equation} \label{eq:flare_freq}
@@
 \end{equation}
-
-Where $f$ is the frequency of flares with energies $\ge E$.
+where $f$ is the frequency of flares with energies $\ge E$.
 We compute the number of expected flares over a time duration and determine the number to create $N$ by a random Poisson draw.
@@
 \end{equation}
-
 where $E_{\text{min}}$ is the minimum considered flare energy and $X$ is a random number on the interval $[0,1)$.
-We set the default values $\alpha=-0.829$, $\beta=26.87$ from \citet{gao2022}. Figure \ref{fig:flare_freq}
-shows our simulation results compared to their powerlaw.
+We set the default values $\alpha=-0.829$, $\beta=26.87$ from \citet{gao2022}, but these can be adjusted by the user.
+Figure \ref{fig:flare_freq} shows our simulation results compared to their power-law values.
 
@@
         {\bf Top}: Generated flare frequencies compared to those expected from \citet{gao2022}, generated using a 10,000 day simulation.
-        The $y$-axis shows the frequency of flares with energies {\em greater than or equal to} the value of the $x$-axis. Error bars
-        based on the square root of the number of observed flares as this is a Poisson process.
-        {\bf Bottom}: Lightcurve of a star flaring with the same powerlaw slope as the top panel, but the intercept ($\beta$) has been increased by 0.3
-        for visual effect. The mean flare temperature is $9000$ K and the mean FWHM is $3$ hours.
+The $y$-axis shows the frequency of flares with energies greater than or equal to the value of the $x$-axis.
+Error bars based on the square root of the number of observed flares as this is a Poisson process.
+{\bf Bottom}: Lightcurve of a star flaring with the same power-law slope as the top panel, but the intercept ($\beta$) has been increased by 0.3 for visual effect.
+The mean flare temperature is $9000$ K and the mean FWHM is $3$ hours.
         }
@@
 The result is a stellar surface that is not constant in temperature and that changes on very short timescales.
-Hydrodynamic simulations of stellar atmospheres \citep[e.g.][]{magic2014} describe hot granules of rising gas surrounded by cooler,
-sinking regions with a \teff a few percent lower than the nominal value.
+Hydrodynamic simulations of stellar atmospheres \citep[e.g.][]{magic2014} describe hot granules of rising gas surrounded by cooler, sinking regions with a temperature a few percent lower than the nominal value.
 
 We model granulation as a global process, and its effects are computed after the effects of spots and faculae.
-Of the remaining ``quiet'' photosphere (i.e. the regions not covered in spots of faculae), a fraction is computed to be part of the cool
-inter-granule surface. The surface coverage of the cool region at any given time is computed by a Gaussian process (GP) using
-the {\sc tinygp} package \citep{foreman-mackey2024} following the methodology of \citet{gordon2020}.
-The GP uses a custom kernel function based on a power spectrum \citep{anderson1990,kallinger2014} to produce random changes in the coverage.
+Of the remaining ``quiet'' photosphere (i.e.
+the regions not covered in spots of faculae), a fraction is computed to be part of the cool
+inter-granule surface; the surface coverage of the cool region at any given time is computed by a Gaussian process (GP) using the {\sc tinygp} package \citep{foreman-mackey2024} following the methodology of \citet{gordon2020}.
+The GP uses a custom kernel function based on a power spectrum \citep{anderson1990,kallinger2014} to produce random changes in the granulation coverage.
 
@@
         Effect of granulation on the lightcurve of a star. This simulation assumes a mean
-        10\% coverage by inter-granule region whose temperature is 300 K lower than the surrounding photosphere.
-        It also assumes that the coverage value varies with a magnitude of 1\% on a timescale of 6 hours.
+        10\% coverage by an inter-granule region whose temperature is 300K lower than the surrounding photosphere.
+It also assumes that the coverage value varies with a magnitude of 1\% on a timescale of 6 hours \TJ{find some constraints in Gordon paper}.
         }
@@
 \subsection{Transits}
+\TJ{Change section to talk about limb darkening and other geometric considerations}
 In the case of a transiting geometry, \texttt{vspec-vsm} is responsible for computing the properties of the portion of the stellar disk that is occulted.
@@
 The code then iterates through each of those points and computes the fraction of it covered by the occulting circle using a 2D numerical integral.
+\TJ{Add a figure to show limb darkening lightcurve.
+Do a surface map like Fig 1 with a transit, then zoom in to show pixel coverage, then a lightcurve}
 
@@
 
-\section{Simple climate models}
+\section{Simplified Climate Model}
 \vspec{} allows the user to upload any desired GCM. However, for questions that focus on detectability,

### 2024-07-23 src/tex/ms.tex
Add Gridpolator section and update scripts

@@
 
-\TJ{TODO}
+\vspec{} pulls its stellar spectral models from a library of pre-computed stellar models.
+Detailed stellar spectral models are computationally expensive to generate, so it is more efficient to interpolate over a descritized grid.
+This interpolation is handled by the Python package GridPolator, whose main class \texttt{GridSpectra} acts as a wrapper for the SciPy or JAX \TJ{(refs?)} \texttt{RegularGridInterpolator} with a specialized API.
+This setup allows spectral grids of arbitrary dimension, and treats the wavelength axis independently of the other parameters so that each wavelength channel is independent.
+When a \texttt{Gridspectra} instance's \texttt{evaluate()} method is called it creates a \texttt{RegularGridInterpolator} instance for each wavelength channel, evaluates that interpolator, and then resamples the returned spectrum at the requested wavelength points.
+In the case that the JAX implementation is used, the portion of \texttt{evaluate()} that does the interpolation is just-in-time (JIT) compiled to dramatically improve performance on subsequent calls.
+
+GridPolator grids can be built arbitrarily by the user, but the package comes with two built-in grids.
+The first\footnote{The raw spectra are hosted on Zenodo: \url{https://zenodo.org/records/10429325}} was created by Peter Hauschildt \TJ{(Private communication?
+He has said he does not feel authorship is warranted.)} using \texttt{PHOENIX} stellar models \citep{allard1994,hauschildt1999,husser2013}.
+Each spectrum spans from 0.1 to 20 microns in steps of $5\times10^{-6}$ microns.
+The stellar atmospheres used to create the spectra have solar metallicity and a $\log{g}$ of 5.0, with effective temperatures ranging from 2300K to 3900K in steps of 100K.
+This is the only grid currently supported by \vspec{}, with support for more grids planned in future releases.
+The second grid is hosted by the Space Telescope Science Institute\footnote{\url{https://archive.stsci.edu/hlsps/reference-atlases/cdbs/grid/phoenix/}}.
+This grid covers a much larger stellar parameter space, with metallicities ranging from $-4$ to $0.5$, \teff from 2000K to 70\,000K, and $\log{g}$ from 0 to 5.5.
+These models are also produced by the \texttt{PHOENIX} code.
+Except for the solar metallicity models, the spectra were computed in 2011 by \citet{allard2003, allard2007} and the solar metallicity models were updated by a new version of \texttt{PHOENIX} in 2021 \citep{allard2012}.
+\TJ{Look up wavelength ranges}
+
+When either of the two above grids are initialized, GridPolator looks for the relevant data files on the user's local system.
+If they are not present, then the files are downloaded and stored in \texttt{\~/.gridpolator/grids/}
+
+Within this suite of codes, GridPolator has the added responsibility of binning spectra and producing wavelength axes that match those output by PSG.
+Wavelength sampling is done using constant resolving power $R=\frac{\lambda}{\Delta \lambda}$.
+The binning algorithm takes a pair of high resolution flux and wavelength arrays and resamples the flux to a lower resolution wavelength axis.
+This algorithm assumes that each point on the new wavelength axis is the central value of the wavelength channel \TJ{(I have never liked this, should I rewrite it before publication?)}.
 

### 2024-07-23 src/scripts/granule_lc.py

@@ -25,9 +25,14 @@ pypsg.docker.set_url_and_run()
 
 header = params.Header(
     data_path=Path('.vspec/granulation_lightcurve'),
-    teff_min=3000*u.K,
-    teff_max=3400*u.K,
-    seed=SEED,verbose=0
+    seed=SEED,verbose=0,
+    spec_grid=params.VSPECGridParameters(
+        max_teff=3400*u.K,
+        min_teff=3000*u.K,
+        impl_bin='rust',
+        impl_interp='scipy',
+        fail_on_missing=False
+    )
 )
 star = params.StarParameters(
     psg_star_template='M',
@@ -48,7 +53,6 @@ star = params.StarParameters(
         dteff=300*u.K
     ),
     grid_params=(500, 1000),
-    spectral_grid='default'
 )
 planet = params.PlanetParameters.std(init_phase=180*u.deg,init_substellar_lon=0*u.deg)
 system = params.SystemParameters(

### 2024-07-23 src/scripts/kempton23.py

@@ -112,7 +112,6 @@ star_kwargs = dict(
     flares=params.FlareParameters.none(),
     granulation=params.GranulationParameters.none(),
     grid_params=(500,1000),
-    spectral_grid='default'
 )
 
 quiet_star = params.StarParameters(
@@ -138,9 +137,12 @@ spotted_star = params.StarParameters(
 
 # Set parameters for simulation
 header_kwargs = dict(
-    teff_min=2300*u.K,teff_max=3400*u.K,
     seed = SEED,
-    verbose = 0
+    verbose = 0,
+    spec_grid=params.VSPECGridParameters(
+        max_teff=3400*u.K,
+        min_teff=2300*u.K,
+    )
 )
 internal_params_kwargs = dict(
     planet=planet_params,

### 2024-07-23 src/scripts/mirecle.py

@@ -118,7 +118,6 @@ quiet_star = params.StarParameters(
     flares=params.FlareParameters.none(),
     granulation=params.GranulationParameters.none(),
     grid_params=(500, 1000),
-    spectral_grid='default'
 )
 
 # Set parameters for simulation
@@ -126,8 +125,11 @@ quiet_star = params.StarParameters(
 internal_params = params.InternalParameters(
     header=params.Header(
         data_path=Path('.vspec/proxcenb'),
-        teff_min=2300*u.K,teff_max=3400*u.K,
-        seed = SEED),
+        seed = SEED,
+        spec_grid=params.VSPECGridParameters(
+            max_teff=3400*u.K,
+            min_teff=2300*u.K,
+        )),
     star = quiet_star,
     psg=psg_params,
     planet=planet_params,

### 2024-07-23 src/scripts/moran23.py

@@ -173,7 +173,6 @@ star_kwargs = dict(
     flares=params.FlareParameters.none(),
     granulation=params.GranulationParameters.none(),
     grid_params=100000,
-    spectral_grid='default'
 )
 quiet_star = params.StarParameters(
     spots=params.SpotParameters.none(),
@@ -198,8 +197,11 @@ spotted_star = params.StarParameters(
 
 # Set parameters for simulation
 header_kwargs = dict(
-    teff_min=2300*u.K,teff_max=3400*u.K,
-    seed = SEED
+    seed = SEED,
+    spec_grid=params.VSPECGridParameters(
+        max_teff=3400*u.K,
+        min_teff=2300*u.K,
+    )
 )
 internal_params_kwargs = dict(
     system=system_params,

### 2024-07-24 src/tex/ms.tex
Rewrite GCM section

@@
 \section{Simplified Climate Model}
-\vspec{} allows the user to upload any desired GCM. However, for questions that focus on detectability,
-the specific physics of the GCM are less important than properties like day/night temperature, surface pressure, or
-the major absorbers in the atmosphere. To simplify these studies, \vspec{} has a built-in GCM that takes in just a
-few parameters and produces a simple-yet-robust exoplanet climate. Simplicity allows the user to know everything about
-the model and reduce surprises, but its basis in thermodynamics and robust energy balance make this model fit for publication-level science \TJ{is this really true? probably not}.
+\vspec{} allows users to run models with any climate model they desire (so long as it can be converted into the \texttt{pypsg.PyGCM} format) and even comes with utilities to work with PSG/GlobES binary files, ExoCAM \citep{wolf2022}, the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013}, and ExoPlaSim \citep{paradise2022} without any custom code from the user.
+
+However, as a full global circulation model can take days to months to converge, there are many cases where a ``quick and dirty'' 3D climate model may be sufficient.
+Built into \vspec{} is such a ``faux'' climate model -- one that does not contain any of the circulation physics of the above sophisticated codes, but instead constructs a 3D planetary atmosphere in a fraction of a second from just a few parameters.
+Such a model can be advantageous, even, because it is extremely predictable; a feature in the phase curve can be traced back to the model parameters that cause it.
 
 The \vspec{} GCM centers around a 2D temperature map of the planet's surface. Based on \citet{cowan2011}, this map
-is constructed completely given the incident flux from the star (parameterized by the host \teff and radius
+is constructed completely given the incident flux from the star (parameterized by the host \teff~and radius
 and the planet's semimajor axis and Bond albedo) and the unitless thermal redistribution efficiency $\epsilon$.
@@
 
-Now that we have the surface temperature at each GCM coordinate we can build the atmosphere on top of it.
-We choose a surface pressure for our model and generate a pressure profile as a function of layer. It is
-important to note that the choice of the pressure profile -- provided it is scaled to the surface pressure and sampled in a reasonable manner -- is
-arbitrary and will not affect the spectrum of the planet. This is because PSG calculates the altitude and thickness of each layer based on the
-pressure and surface gravity; the pressure profile is essentially a proxy for the sampling of atmosphere layers as a function of altitude.
+To compute the surface temperature we first compute the temperature at the equator as a function of longitude. This is done by integrating Equation (10) of \citet{cowan2011}
+\begin{equation}
+    \frac{d \tilde{T}}{d \Phi} = \frac{1}{\epsilon}(\max{(\cos{\Phi}, 0)} - \tilde{T}^4)
+\end{equation}
+Where $\tilde{T}$ is the dimensionless temperature defined as $T/T_0$, or the temperature normalized by a local fiducial value.
+$\Phi$ is the angle from the substellar point, so we integrate if from $-\pi$ to $\pi$.
+This integral is done numerically in Figure \ref{fig:cowan_curves}, the optimal scheme depending on the value of $\epsilon$.
+For $0 < \epsilon < 1$ we adopt a reflecting initial value problem (IVP) scheme, where we set $\tilde{T}=1$ at $\Phi=\pi$, integrate to $-\pi$, then use the last value as the initial value for an integration from $-\pi$ to $\pi$.
+Note that for extremely small values of $\epsilon$ (e.g.
+$10^{-6}$) the integration is slow, and will fail if $\epsilon=0$; we recommended a lower limit of $\epsilon=10^{-4}$.
+When $1\le \epsilon < 10$ we solve the integral using a boundary value problem (BVP) scheme.
+For each of these two methods we employ the SciPy implementation (\texttt{solve\_ivp()} and \texttt{solve\_bvp()}, respectively).
+When $\epsilon \ge 10$ it is more efficient to use the analytic approximation given by \citet{cowan2011} Equation (A2).
 
-Next we create an adiabatic temperature profile at each GCM coordinate. An adiabat assumes that
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/cowan_fig1.pdf}
+    \caption{
+        Reproduction of \citet{cowan2011} Figure 1 using the methods described in the text.
+These curves show the dimensionless temperature at the equator of a planet as a function of angle from substellar point $\Phi$ for thermal inertia $\epsilon$.
+Note that the $\epsilon=0$ curve in the original figure is replaced with $\epsilon=10^{-4}$ (see text), and we do not include additional curves for mid-latitudes as in the original figure.
+As $\epsilon$ increases heat is moved from the dayside onto the nightside, and the temperature maximum shifts away from the substellar point.
+    }
+    \script{cowan_fig.py}
+    \label{fig:cowan_curves}
+\end{figure}
+
+\subsection{Latitudinal Mixing}
+The treatment by \citet{cowan2011} includes no mixing in the latitudinal direction, and the temperature tends to 0 as you approach the poles.
+We add a parameter $\alpha$ that is defined by the ratio between the polar temperature and the equitorial temperature at any given longitude.
+Let $f(\lambda;\alpha)$ be the temperature as a function of latitude $\lambda$ at an arbitrary longitude.
+If the surface is in equilibrium, the energy balance requires the outgoing energy flux to be independent of $\alpha$:
 \begin{equation}
-    T^\gamma P^{1 - \gamma} = \text{const.}
+    \label{eq:e_out}
+    E_{\rm out} = \sigma \int_0^{\pi/2} f(\lambda; \alpha)^4 \cos{\lambda}\,d\lambda
 \end{equation}
-where $\gamma$ is the adiabatic index. Therefore the temperature profile can be computed
+Where $E_{\rm out}$ is the outbound energy flux of an infintesimal sliver of differential longitude.
+In the case of $\alpha=0$ (i.e. no mixing) we can treat each latitude value independently, so the equilibrium temperature is
 \begin{equation}
-    T = T_{\rm surf} \left(\frac{P}{P_{\rm surf}} \right)^{1 - \frac{1}{\gamma} }
+    f(\lambda; 0) = f(0;0) \cos^{1/4}{\lambda}
+\end{equation}
+
+Where $f(0;0)$ is the previously computed temperature at the equator (see Figure \ref{fig:cowan_curves}). Plugging this in to Equation (\ref{eq:e_out}) we see
+\begin{equation}
+    \label{eq:nomix}
+    \begin{array}{cl}
+        E_{\rm out} & = \sigma \int_0^{\pi/2} (f(0;0) \cos^{1/4}{\lambda})^4\cos{\lambda}\,d\lambda \\
+        & = \sigma f^4(0;0) \int_0^{\pi/2} \cos^{2}{\lambda}\,d\lambda \\
+        & = \frac{\pi}{4}\sigma f^4(0;0)
+    \end{array}
+\end{equation}
+
+If instead we let $f(\lambda;1)$ be a constant function where the equator and pole have equal temperatures, then
+\begin{equation}
+    \label{eq:allmix}
+    \begin{array}{cl}
+        E_{\rm out} & = \sigma \int_0^{\pi/2} f(\lambda;1)^4 \cos{\lambda}\,d\lambda \\
+        & = \sigma f^4(0;1) \int_0^{\pi/2} \cos{\lambda}\,d\lambda \\
+        & = \sigma f^4(0;1)
+    \end{array}
+\end{equation}
+
+We calculate $f(0,1)$ by setting Equations (\ref{eq:nomix} \& \ref{eq:allmix}) equal to find
+\begin{equation}
+    f(0;1) = \left(\frac{\pi}{4}\right)^{1/4} f(0;0) \approx 0.94 \,f(0;0)
+\end{equation}
+
+We can now write a general parameterization for the temperature as a function of latitude
+\begin{equation}
+    \label{eq:temp_gen}
+    f(\lambda;\alpha) = f(0;0)\,\left(\alpha \frac{\pi}{4} + (1-\alpha) \cos{\lambda}\right)^{1/4}
+\end{equation}
+
+\subsection{Pressure and Temperature}
+Combining Equation (\ref{eq:temp_gen}) with the numerically-computed equatorial temperatures yields a 2D temperature map that covers the surface of the planet.
+We construct a 3D atmosphere on top of this surface assuming the atmosphere is adiabatic with index $\gamma$.
+The user supplies both the adiabatic index and the surface pressure $P_{\rm surf}$.
+The temperature at some altitude $z$ is given by
+\begin{equation}
+    T_z = T_{\rm surf}\,\left( \frac{P_z}{P_{\rm surf}} \right)^{1-\frac{1}{\gamma}}
 \end{equation}
@@
 Finally, we populate our atmosphere with molecules. The user specifies the species and abundance and atmosphere is filled accordingly with a constant molecular suite.
-
-The caveats of this approach lie in the simplified physics of the atmosphere structure. The heat-redistribution model does not
-capture the effects of latitudinal mixing; an adiabatic temperature profile will never exhibit the temperature inversion
-that we know form in complex atmospheres like Earth's. Similarly, setting a constant abundance does not allow for water vapor to
-vary with temperature as it does on Earth. We must also be careful not to add so much of an absorbing species that the energy balance of
-the atmosphere would change; instead we treat them as trace species in an otherwise transparent atmosphere. Our simplified approach
-allows us to easily examine a large parameter space of atmospheres whose behaviors are easy to predict.
 

### 2024-07-24 src/scripts/cowan_fig.py

@@ -1,0 +1,35 @@
+"""
+recreate Figure 1 from Cowan & Agol (2011)
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+from VSPEC.gcm import heat_transfer as ht
+
+import paths
+
+OUTFIG = paths.figures / 'cowan_fig1.pdf'
+
+eps = [1e-4,0.2*np.pi,2*np.pi]
+label = ['10^{-4}','2\\pi/10','2\\pi']
+modes = ['ivp_reflect','ivp_reflect','bvp']
+colors = ['b','k','r']
+
+n_points = 100
+
+fig,ax = plt.subplots(1,1,figsize=(4,3))
+
+for e,l,m,c in zip(eps,label,modes,colors):
+    lons, tsurf = ht.get_equator_curve(e,n_points,m)
+    index = np.argwhere(lons>=-0.5*np.pi)[0][0]
+    lons = np.concatenate([lons[index:],lons[:index]+2*np.pi])
+    tsurf = np.concatenate([tsurf[index:],tsurf[:index]])
+    ax.plot(lons,tsurf,color=c,label=f'$\\epsilon = {l}$')
+
+ax.set_xlabel('$\\Phi$')
+ax.set_ylabel('$T/T_0$')
+ax.set_xticks([-0.5*np.pi,0,0.5*np.pi,np.pi, 1.5*np.pi,])
+ax.set_xticklabels(['$-\\pi/2$','$0$','$\\pi/2$','$\\pi$','$3\\pi/2$'])
+_=ax.legend()
+fig.tight_layout()
+fig.savefig(OUTFIG)

### 2024-07-24 src/tex/ms.tex
Add some references

@@
 \vspec{} uses a Python interface for the PSG API system called \texttt{pypsg}, which is a general stand-alone library built for any Python-based API call to PSG.
-\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django \TJ{cite} to encode a PSG configuration file as a native Python object.Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
-Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests} \TJ{cite} package.
+\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object.Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
+Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package.
 This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key
@@
 GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial.
-\texttt{pypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models (e.g.
-ExoCAM, WACCM \TJ{full names and ref?}) to the correct format.
+\texttt{pypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.
 
@@
 \section{Simplified Climate Model}
-\vspec{} allows users to run models with any climate model they desire (so long as it can be converted into the \texttt{pypsg.PyGCM} format) and even comes with utilities to work with PSG/GlobES binary files, ExoCAM \citep{wolf2022}, the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013}, and ExoPlaSim \citep{paradise2022} without any custom code from the user.
+\vspec{} allows users to run models with any climate model they desire (so long as it can be converted into the \texttt{pypsg.PyGCM} format) and even comes with utilities to work with PSG/GlobES binary files, ExoCAM \citep{wolf2022}, WACCM \citep{marsh2013}, and ExoPlaSim \citep{paradise2022} without any custom code from the user.
 

### 2024-07-24 src/tex/ms.tex
Update language about objects representing spots, etc

@@
 The main model class is the \texttt{Star} object, which describes bulk properties (e.g. radius, photospheric \teff, limb darkening parameters) in addition to acting as 
-a container for objects representing sources of variability. \TJ{Additional classes are defined for additional stellar components, e.g. spots, faculae,..., which are discussed below.}
+a container for objects representing sources of variability such as \texttt{StarSpot}, \texttt{Facula}, \texttt{StelarFlare}, and \text{Granulation}, which are described below.
 

### 2024-07-24 src/tex/ms.tex
Import changes from Ted/Avi on July 23

@@
 
-\vspec{} pulls its stellar spectral models from a library of pre-computed stellar models.
-Detailed stellar spectral models are computationally expensive to generate, so it is more efficient to interpolate over a descritized grid.
-This interpolation is handled by the Python package GridPolator, whose main class \texttt{GridSpectra} acts as a wrapper for the SciPy or JAX \TJ{(refs?)} \texttt{RegularGridInterpolator} with a specialized API.
-This setup allows spectral grids of arbitrary dimension, and treats the wavelength axis independently of the other parameters so that each wavelength channel is independent.
-When a \texttt{Gridspectra} instance's \texttt{evaluate()} method is called it creates a \texttt{RegularGridInterpolator} instance for each wavelength channel, evaluates that interpolator, and then resamples the returned spectrum at the requested wavelength points.
+\vspec{} utilizes a library of pre-computed stellar models for producing its multi-component composite stellar spectrum; detailed stellar spectral models are computationally expensive to generate, so it is more efficient to interpolate over a discretized grid.
+This interpolation is handled by the VSPEC-related Python package GridPolator, whose main class \texttt{GridSpectra} acts as a wrapper for the SciPy or JAX \TJ{(refs?)} \texttt{RegularGridInterpolator} with a specialized API.
+This setup accomodates spectral grids of arbitrary dimension, and treats the wavelength axis independently of the other parameters so that each wavelength channel is independent.
+When a \texttt{Gridspectra} instance's \texttt{evaluate()} method is called, it creates a \texttt{RegularGridInterpolator} instance for each wavelength channel, evaluates that interpolator, and then resamples the returned spectrum at the requested wavelength points.
 In the case that the JAX implementation is used, the portion of \texttt{evaluate()} that does the interpolation is just-in-time (JIT) compiled to dramatically improve performance on subsequent calls.
+\TJ{See Jax/scipy figure}
 
-GridPolator grids can be built arbitrarily by the user, but the package comes with two built-in grids.
-The first\footnote{The raw spectra are hosted on Zenodo: \url{https://zenodo.org/records/10429325}} was created by Peter Hauschildt \TJ{(Private communication?
-He has said he does not feel authorship is warranted.)} using \texttt{PHOENIX} stellar models \citep{allard1994,hauschildt1999,husser2013}.
+GridPolator grids can be built arbitrarily by the user, but the package comes with a built-in custom grid of \texttt{PHOENIX} stellar models \citep{allard1994,hauschildt1999,husser2013} created specifically for \vspec{}.
 Each spectrum spans from 0.1 to 20 microns in steps of $5\times10^{-6}$ microns.
 The stellar atmospheres used to create the spectra have solar metallicity and a $\log{g}$ of 5.0, with effective temperatures ranging from 2300K to 3900K in steps of 100K.
-This is the only grid currently supported by \vspec{}, with support for more grids planned in future releases.
-The second grid is hosted by the Space Telescope Science Institute\footnote{\url{https://archive.stsci.edu/hlsps/reference-atlases/cdbs/grid/phoenix/}}.
+Currently this is the only grid currently supported by \vspec{}, with support for additional grids planned in future releases.
+% The second grid is hosted by the Space Telescope Science Institute\footnote{\url{https://archive.stsci.edu/hlsps/reference-atlases/cdbs/grid/phoenix/}}.
 This grid covers a much larger stellar parameter space, with metallicities ranging from $-4$ to $0.5$, \teff from 2000K to 70\,000K, and $\log{g}$ from 0 to 5.5.
@@
 \TJ{Look up wavelength ranges}
-
-When either of the two above grids are initialized, GridPolator looks for the relevant data files on the user's local system.
-If they are not present, then the files are downloaded and stored in \texttt{\~/.gridpolator/grids/}
+When the grid is initialized, GridPolator looks for the relevant data files on the user's local system.
+If they are not present, then the files are downloaded and stored in \texttt{\textasciitilde /.gridpolator/grids/}.
 
@@
 Wavelength sampling is done using constant resolving power $R=\frac{\lambda}{\Delta \lambda}$.
-The binning algorithm takes a pair of high resolution flux and wavelength arrays and resamples the flux to a lower resolution wavelength axis.
-This algorithm assumes that each point on the new wavelength axis is the central value of the wavelength channel \TJ{(I have never liked this, should I rewrite it before publication?)}.
+The binning algorithm takes a pair of high-resolution flux and wavelength arrays and resamples the flux to a lower-resolution wavelength axis.
+This algorithm currently assumes that each point on the new wavelength axis is the central value of the wavelength channel, but future versions will treat each spectral channel \TJ{Ask Geronimo about the particulars of PSG's binning/wavelength axis creation}.
+\TJ{(I have never liked this, should I rewrite it before publication?)}.
 

### 2024-07-29 src/tex/ms.tex
Add limb darkening section and figure

@@
     \script{surface_map_and_lc.py}
-    \caption{{\bf Top}: The surface map of a star with spots during a transit.
-    The large ellipses are spots, while the transiting planet occults a portion of the disk represented by the black circle.
-    {\bf Bottom}: An example lightcurve of a spotted star.
-    \TJ{Remove transit and put it in a LD figure}}
+    \caption{{\bf Top}: The surface map of a star with spots. The star has a photospheric temperature of 3300K, a penumbra temperature of 2700K, and an umbra temperature of 2500K.
+    {\bf Bottom}: An example lightcurve of a spotted star. Flux variations are due to varying spot coverage as a function of sub-observer point. Two rotational periods are shown.
+    }
     \label{fig:surface_map}
@@
 
-\subsection{Transits}
-\TJ{Change section to talk about limb darkening and other geometric considerations}
+\subsection{Transits \& Limb Darkening}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/ld_transit.pdf}
+    \caption{
+        {\bf Right}: Transit lightcurves at 1 ${\rm \mu m}$ for a Lambertian (i.e.
+no limb darkening) star and one with significant limb darkening \citep[$u_1=0.5,\,u_2=0.15$,][]{espinoza2015}.
+In each case a 1 $R_e$ planet transits a 1 $R_\odot$ ($T_{\rm eff} = 3000$ K) star with an observation cadence of 1 minute.
+Scattered points show the observation with simulated noise, considering JWST/NIRSpec G395H binned to a resolving power of 270 and a target distance of 5 pc.
+        {\bf Right}: Stellar surface maps including translucent masks to show limb darkening and transit.
+The Lambertian star (top) does not exhibit any limb darkening, and its disk has constant surface brightness.
+However, the limb-darkened star (bottom) is much brighter at disk-center than it is near the limb.
+The dark circle on the leftmost limb of each star is the region occulted by the planet when it is $0.7^\circ$ from mid-transit.
+    }
+    \script{ld_transit.py}
+    \label{fig:ld_transit}
+\end{figure*}
+
 In the case of a transiting geometry, \texttt{vspec-vsm} is responsible for computing the properties of the portion of the stellar disk that is occulted.
@@
 The code then iterates through each of those points and computes the fraction of it covered by the occulting circle using a 2D numerical integral.
-\TJ{Add a figure to show limb darkening lightcurve.
-Do a surface map like Fig 1 with a transit, then zoom in to show pixel coverage, then a lightcurve}
+The weight of each point is determined by the geometric projected area as well as a quadratic limb darkening law \citep[see][]{espinoza2015}.
+This treatment allows the user to specify two parameters $u_1$ and $u_2$ that describe the flux variation as a function of angle from disk-center.
+A caveat of this method is that there is no dependence on wavelength.
+A better treatment would vary the temperature of the stellar model as a function of $\mu$; we simply apply a gray scaling factor to the flux.
+Figure \ref{fig:ld_transit} compares transit lightcurves for two different sets of limb darkening parameters.
 

### 2024-07-29 src/scripts/ld_transit.py

@@ -1,0 +1,86 @@
+"""
+Plot the lightcurve of a transit with limb darkening.
+"""
+
+import numpy as np
+from astropy import units as u
+import matplotlib.pyplot as plt
+import cartopy.crs as ccrs
+import VSPEC
+import pypsg
+
+import paths
+
+LAMBERT_CONFIG = paths.scripts / 'ld_lc_lamb.yaml'
+SOLAR_CONFIG = paths.scripts / 'ld_lc_solar.yaml'
+OUTFILE = paths.figures / 'ld_transit.pdf'
+
+pypsg.docker.set_url_and_run()
+
+lam_model = VSPEC.ObservationModel.from_yaml(LAMBERT_CONFIG)
+sol_model = VSPEC.ObservationModel.from_yaml(SOLAR_CONFIG)
+
+# lam_model._build_star()
+# sol_model._build_star()
+
+lam_model.build_planet()
+lam_model.build_spectra()
+sol_model.build_planet()
+sol_model.build_spectra()
+
+lam_data = VSPEC.PhaseAnalyzer.from_model(lam_model)
+sol_data = VSPEC.PhaseAnalyzer.from_model(sol_model)
+
+fig = plt.figure(figsize=(8, 6))
+gs = fig.add_gridspec(5, 2)
+fig.subplots_adjust(hspace=0,wspace=0.5)
+ax1: plt.Axes = fig.add_subplot(gs[1:4,1])
+
+wl_index = np.argmin(np.abs(lam_data.wavelength - 1*u.um))
+
+ax1.plot(lam_data.time.to_value(u.hr), lam_data.lightcurve('total',wl_index,normalize=0), 'r', label='Lambertian')
+ax1.scatter(lam_data.time.to_value(u.hr), lam_data.lightcurve('total',wl_index,normalize=0,noise=True), c='r',marker='.',s=1,alpha=1)
+ax1.plot(sol_data.time.to_value(u.hr), sol_data.lightcurve('total',wl_index,normalize=0), 'k', label='$u_1=0.5,~u_2=0.15$')
+ax1.scatter(sol_data.time.to_value(u.hr), sol_data.lightcurve('total',wl_index,normalize=0,noise=True), c='k', marker='.',s=1,alpha=1)
+
+
[…]

### 2024-07-29 src/scripts/surface_map_and_lc.py

@@ -47,7 +47,7 @@ pl_orbit = 0.05*u.AU
 inclination = 89.8*u.deg
 phase = 180.4*u.deg
 
-model.star.plot_surface(lat0, lon0, ax1, pl_orbit, pl_radius, phase, inclination)
+model.star.plot_surface(lat0, lon0, ax1)
 
 # LC
 

### 2024-07-29 src/tex/ms.tex
Give source of Gao2022 data

@@
 
-Before simulating an observation, \vspec{} asks \texttt{vspec-vsm} to pre-compute a population of flares based on a power-law frequency-energy relationship similar to the results of \TJ{what data?} \citep{gao2022}:
+Before simulating an observation, \vspec{} asks \texttt{vspec-vsm} to pre-compute a population of flares based on a power-law frequency-energy relationship based on Kepler and TESS lightcurves \citep{gao2022}:
 \begin{equation} \label{eq:flare_freq}

### 2024-07-29 src/tex/ms.tex
Change the language about granulation parameters. I could not find any specific parameters in the Gordon paper, so I changed the wording to make it clear that these parameters were chosen arbitrarily.

@@
     \caption{
-        Effect of granulation on the lightcurve of a star. This simulation assumes a mean
-        10\% coverage by an inter-granule region whose temperature is 300K lower than the surrounding photosphere.
-It also assumes that the coverage value varies with a magnitude of 1\% on a timescale of 6 hours \TJ{find some constraints in Gordon paper}.
+        Effect of granulation on the lightcurve of a star.
+The parameters chosen in this simulation mean that, on average, 10\% of the stellar surface is covered by the cool, inter-granule region with a temperature 300K lower than the surrounding photosphere, and that this coverage varies by about 1\% on a timescale of 6 hours.
         }

### 2024-07-29 src/tex/ms.tex
Make JAX v SciPy figure and add refs

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/jax_v_scipy.pdf}
+    \caption{
+        Speed comparison between the JAX and SciPy implementations of the GridPolator interpolation backend.
+Both interpolators use the builtin \vspec{} grid, with \teff ranging from 2800 to 3300K, and wavelengths from 5-12 ${\rm \mu m}$ with $R=100$.
+SciPy is most efficient for a small number of evaluations, with a constant cost per call.
+For a large number of evaluations, however, the speed gained via JAX's JIT compilation offsets the cost of performing that compilation.
+    }
+    \script{jax_v_scipy.pdf}
+    \label{fig:jax_v_scipy}
+\end{figure}
+
 \vspec{} utilizes a library of pre-computed stellar models for producing its multi-component composite stellar spectrum; detailed stellar spectral models are computationally expensive to generate, so it is more efficient to interpolate over a discretized grid.
-This interpolation is handled by the VSPEC-related Python package GridPolator, whose main class \texttt{GridSpectra} acts as a wrapper for the SciPy or JAX \TJ{(refs?)} \texttt{RegularGridInterpolator} with a specialized API.
+This interpolation is handled by the VSPEC-related Python package GridPolator, whose main class \texttt{GridSpectra} acts as a wrapper for the SciPy \citep{virtanen2020} or JAX \citep{jamesbradbury2018} \texttt{RegularGridInterpolator} with a specialized API.
 This setup accomodates spectral grids of arbitrary dimension, and treats the wavelength axis independently of the other parameters so that each wavelength channel is independent.
@@
 In the case that the JAX implementation is used, the portion of \texttt{evaluate()} that does the interpolation is just-in-time (JIT) compiled to dramatically improve performance on subsequent calls.
-\TJ{See Jax/scipy figure}
+Figure \ref{fig:jax_v_scipy} compares the computational cost of JAX with that of SciPy.
 

### 2024-07-29 src/scripts/jax_v_scipy.py

@@ -1,0 +1,76 @@
+"""
+Compare jax and scipy speed.
+"""
+
+import jax.numpy as jnp
+import numpy as np
+from time import time
+import matplotlib.pyplot as plt
+from astropy import units as u
+
+from GridPolator import GridSpectra
+
+import paths
+
+OUTFIE = paths.figures / 'jax_v_scipy.pdf'
+
+w1 = 5 * u.um
+w2 = 12 * u.um
+resolving_power = 100
+teffs = [2800,2900,3000,3100,3200,3300]
+impl_bin = 'rust'
+
+wl_jnp = jnp.linspace(5.0, 11.2, 100)
+wl_np = np.linspace(5.0, 11.2, 100)
+params_jnp = (jnp.array([2900.]),)
+params_np = (np.array([2900.]),)
+
+
+fig, ax  = plt.subplots(1,1,figsize=(4,3))
+
+N=3000
+
+g_jax = GridSpectra.from_vspec(
+    w1=w1,
+    w2=w2,
+    resolving_power=resolving_power,
+    teffs=teffs,
+    impl_bin=impl_bin,
+    impl_interp='jax',
+    fail_on_missing=False
+)
+g_scipy = GridSpectra.from_vspec(
+    w1=w1,
+    w2=w2,
+    resolving_power=resolving_power,
+    teffs=teffs,
+    impl_bin=impl_bin,
+    impl_interp='scipy',
+    fail_on_missing=False
+)
+
+dt_jax = np.zeros(N)
+
+for i in range(N):
+    start = time()
+    flux_jnp = g_jax.evaluate(params_jnp, wl_jnp)
+    end = time()
+    dt_jax[i] = end - start
+
+dt_scipy = np.zeros(N)
+
+for i in range(N):
+    start = time()
+    flux_np = g_scipy.evaluate(params_np, wl_np)
+    end = time()
+    dt_scipy[i] = end - start
+
+x = np.arange(N)
+
+ax.plot(x, np.cumsum(dt_jax), label='JAX',c='#B96EBD')
+ax.plot(x, np.cumsum(dt_scipy), label='Scipy',c='#0054A6')
+ax.set_xlabel('Iteration')
+ax.set_ylabel('Time (s)')
+fig.tight_layout()
+_=ax.legend()
+fig.savefig(OUTFIE)

### 2024-07-30 src/tex/ms.tex
Rewrite facula area section

@@
 To consider this effect in the faculae lightcurve, we compute the fraction of each facula's projected area -- the area on the disk it would occupy as seen by the observer -- that is occupied by the hot wall versus the cool floor.
-This is done via numerical integral along the radius of the spot.
+Mathematically, this is no different from finding the area of the intersection of two overlapping ellipses:
 
-We calculate each portion using geometrical relationships between the spot dimensions and the viewing angle:
+As shown in Figure \ref{fig:fac_struct}, peering through the top of the pore the observer can see a portion of the hot wall and a portion of the cool floor.
+Consider an ellipse centered at the origin with semimajor axis $R$ and semiminor axis $\mu R$, oriented so that the semimajor axis points along the $y$-axis.
+Now consider a second, identical ellipse, displaced along the $x$-axis by some amount $0<d<\mu R$.
+The intersection of these two ellipses is the portion of the cool floor that is visible, and the area of the second ellipse outside the intersection is the portion of the hot wall that the observer sees.
+We are not interested in the absolute areas of each of these regions, only how they relate to the total ellipse area $\pi \mu R^2$, and so we can generalize to the case that $\mu=1$ (i.e.
+circles).
+This problem is solved for arbitrary circles by \citet{weisstein2004}.
+The details of their derivation are not relevant to this text, but we use their result to create a function $f(d,R)$ that computes the fraction of a circle with radius $R$ and center $(d,0)$ that is inside the unit circle.
+For a facula with radius $r$, wall height $h$, and angle from disk center $theta$, we find that the effective area of the cool floor is:
 \begin{equation}
-    f_{wall} = \frac{\int _{-R}^{R} Z_{\rm eff} dr}{\int _{-R}^{R} Z_{\rm eff} dr + \int _{-R}^{R} R_{\rm eff} dr}
+    \frac{A_{\rm floor}}{A_{\rm total}} = f\left(\frac{h}{r}\tan{\theta}, 1 \right)
 \end{equation}
-where $Z_{\rm eff}$ is the apparent height of the wall
+and for the hot wall it is:
 \begin{equation}
-    Z_{\rm eff} = \left\{
-    \begin{array}{lr}
-        Z_w \sin{\alpha}, & \text{if } \alpha \leq \alpha_{\rm crit} \\
-        2\sqrt{R^2 - r^2}\cos{\alpha}, & \text{if } \alpha > \alpha_{\rm crit}
-    \end{array}
-    \right\}
+    \frac{A_{\rm wall}}{A_{\rm total}} = 1-  f\left(\frac{h}{r}\tan{\theta}, 1 \right)
 \end{equation}
-and $R_{\rm eff}$ is the apparent width of the floor
-\begin{equation}
-    R_{\rm eff} = \left\{
-    \begin{array}{lr}
-        2\sqrt{R^2 - r^2} - Z_w\sin{\alpha}, & \text{if } \alpha \leq \alpha_{\rm crit} \\
-        0, & \text{if } \alpha > \alpha_{\rm crit}
-    \end{array}
-    \right\}
-\end{equation}
-for facula radius $R$, depth $Z_w$, and angle from disk-center $\alpha$.
-$r$ in this numerical scheme is defined as the distance from the center of the facula along the radial line connecting the facula center to the disk center.
-$\alpha_{\rm crit}$ is the value of alpha at which the floor is no longer visible and is defined to be $\arctan{\frac{2\sqrt{R^2-r^2}}{Z_w}}$.
 
@@
 \end{equation}
-\TJ{Change variable names, make diagram}
 where $r$ is the radius of the facula, $r_{\rm min}$ is the minimum radius where the floor is visible, $\Delta  T_{\rm eff, floor, 0}$ is the difference between the floor and photosphere \teff at $r_{\rm min}$, and $m_{\rm floor}$ is the slope of the relationship with units of [temperature] [length]$^{-1}$.
@@
     \centering
-    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_model.png}}
+    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_diagram.pdf}}
     \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_depth.pdf}}

### 2024-07-30 src/scripts/fac_depth.py

@@ -1,0 +1,47 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+import paths
+
+from vspec_vsm.helpers import calc_circ_fraction_inside_unit_circle
+
+OUTFILE = paths.figures / 'fac_depth.png'
+
+h=2
+alpha = np.pi/4 * 0.1
+
+
+x = np.linspace(-2,2,1000)
+y = np.linspace(-2,2,1000)
+
+X,Y = np.meshgrid(x,y)
+
+R = np.sqrt((X/np.cos(alpha))**2 + Y**2)
+THETA = np.arctan2(Y,X)
+
+INSIDE = R<1
+
+DIST_FROM_RIGHT = (np.cos(alpha)*np.sqrt(np.abs(1-Y**2))-X)
+
+Heff = np.sin(alpha) * h
+WALL = (DIST_FROM_RIGHT<Heff) & INSIDE
+
+Z = np.ones_like(X) * np.nan
+Z = np.where(INSIDE,0,Z)
+Z = np.where(WALL,1,Z)
+
+wall_frac = np.sum(WALL)/np.sum(INSIDE)
+circ_frac = (1 - calc_circ_fraction_inside_unit_circle(h*np.tan(alpha),0,1))
+
+
+fig,ax = plt.subplots(1,1,figsize=(4,3))
+ax.set_aspect('equal')
+
+im = ax.pcolormesh(X/np.cos(alpha),Y,Z,vmin=0,vmax=1)
+
+ax.text(0.3,1.1,f'$f_{{\\rm wall}} = {wall_frac:.4f}$')
+ax.text(0.3,1.4,f'$f_{{\\rm circ}} = {circ_frac:.4f}$')
+
+fig.colorbar(im)
+fig.savefig(OUTFILE,dpi=300)
+

### 2024-07-30 src/scripts/facula_depth.py

@@ -80,7 +80,7 @@ fig.savefig(outfig)
 
 # Copy the other figure now
 
-static_fig_name = 'facula_model.png'
+static_fig_name = 'facula_diagram.pdf'
 static_fig_path = paths.static / static_fig_name
 target_path = paths.figures / static_fig_name
 

### 2024-07-30 src/scripts/ld_transit.py

@@ -20,13 +20,13 @@ pypsg.docker.set_url_and_run()
 lam_model = VSPEC.ObservationModel.from_yaml(LAMBERT_CONFIG)
 sol_model = VSPEC.ObservationModel.from_yaml(SOLAR_CONFIG)
 
-# lam_model._build_star()
-# sol_model._build_star()
+lam_model._build_star()
+sol_model._build_star()
 
-lam_model.build_planet()
-lam_model.build_spectra()
-sol_model.build_planet()
-sol_model.build_spectra()
+# lam_model.build_planet()
+# lam_model.build_spectra()
+# sol_model.build_planet()
+# sol_model.build_spectra()
 
 lam_data = VSPEC.PhaseAnalyzer.from_model(lam_model)
 sol_data = VSPEC.PhaseAnalyzer.from_model(sol_model)
@@ -39,9 +39,13 @@ ax1: plt.Axes = fig.add_subplot(gs[1:4,1])
 wl_index = np.argmin(np.abs(lam_data.wavelength - 1*u.um))
 
 ax1.plot(lam_data.time.to_value(u.hr), lam_data.lightcurve('total',wl_index,normalize=0), 'r', label='Lambertian')
-ax1.scatter(lam_data.time.to_value(u.hr), lam_data.lightcurve('total',wl_index,normalize=0,noise=True), c='r',marker='.',s=1,alpha=1)
+obs = lam_data.lightcurve('total',wl_index)
+obs = obs/obs[0] + np.random.normal(scale=lam_data.lightcurve('noise',wl_index)/obs[0])
+ax1.scatter(lam_data.time.to_value(u.hr), obs, c='r',marker='.',s=1,alpha=1)
 ax1.plot(sol_data.time.to_value(u.hr), sol_data.lightcurve('total',wl_index,normalize=0), 'k', label='$u_1=0.5,~u_2=0.15$')
-ax1.scatter(sol_data.time.to_value(u.hr), sol_data.lightcurve('total',wl_index,normalize=0,noise=True), c='k', marker='.',s=1,alpha=1)
+obs = sol_data.lightcurve('total',wl_index)
+obs = obs/obs[0] + np.random.normal(scale=sol_data.lightcurve('noise',wl_index)/obs[0])
+ax1.scatter(sol_data.time.to_value(u.hr), obs, c='k', marker='.',s=1,alpha=1)
 
 
[…]

### 2024-07-30 src/tex/ms.tex
A few small edits

@@
 no limb darkening) star and one with significant limb darkening \citep[$u_1=0.5,\,u_2=0.15$,][]{espinoza2015}.
-In each case a 1 $R_e$ planet transits a 1 $R_\odot$ ($T_{\rm eff} = 3000$ K) star with an observation cadence of 1 minute.
-Scattered points show the observation with simulated noise, considering JWST/NIRSpec G395H binned to a resolving power of 270 and a target distance of 5 pc.
+In each case a 1 $R_e$ planet transits a 0.12 $R_\odot$ ($T_{\rm eff} = 3000$ K) star with an observation cadence of 1 minute.
+Scattered points show the observation with simulated noise, considering a single wavelength channel from JWST/NIRSpec G395H binned to a resolving power of 270 and a target distance of 5 pc.
         {\bf Right}: Stellar surface maps including translucent masks to show limb darkening and transit.
@@
     }
-    \script{jax_v_scipy.pdf}
+    \script{jax_v_scipy.py}
     \label{fig:jax_v_scipy}

### 2024-07-30 src/tex/ms.tex
Commit changes from Ted/Avi last friday

@@
     inclination: 89.7 deg
-    phase_of_pariasteron: 0 deg
+    phase_of_periasteron: 0 deg
 \end{lstlisting}
@@
 \section{Simplified Climate Model}
-\vspec{} allows users to run models with any climate model they desire (so long as it can be converted into the \texttt{pypsg.PyGCM} format) and even comes with utilities to work with PSG/GlobES binary files, ExoCAM \citep{wolf2022}, WACCM \citep{marsh2013}, and ExoPlaSim \citep{paradise2022} without any custom code from the user.
+\vspec{} allows users to run simulations with any climate model they desire (so long as it can be converted into the \texttt{pypsg.PyGCM} format) and even comes with utilities to work with PSG/GlobES binary files, ExoCAM \citep{wolf2022}, WACCM \citep{marsh2013}, and ExoPlaSim \citep{paradise2022} without any custom code from the user.
 
@@
 Built into \vspec{} is such a ``faux'' climate model -- one that does not contain any of the circulation physics of the above sophisticated codes, but instead constructs a 3D planetary atmosphere in a fraction of a second from just a few parameters.
-Such a model can be advantageous, even, because it is extremely predictable; a feature in the phase curve can be traced back to the model parameters that cause it.
+Such a model can also be advantageous because it is extremely predictable; a feature in the phase curve can easily be traced back to the model parameters that cause it.
 
 The \vspec{} GCM centers around a 2D temperature map of the planet's surface. Based on \citet{cowan2011}, this map
-is constructed completely given the incident flux from the star (parameterized by the host \teff~and radius
+is determined exclusively by the incident flux from the star (parameterized by the host \teff~and radius
 and the planet's semimajor axis and Bond albedo) and the unitless thermal redistribution efficiency $\epsilon$.
-This efficiency is 0 for a planet that reradiates instantly and $\gg 1$ for an atmosphere that efficiently redistributes heat to its night side.
+This efficiency is 0 for a planet that re-radiates instantly and $\gg 1$ for an atmosphere that efficiently redistributes heat to its night side.
 Figure \ref{fig:thermal_inertia} shows a temperature map example for an Earth-like planet with $\epsilon = 2\pi$.
-These temperature maps are validated by the energy balance of the planet; \vspec{} computes the incident and emergent
-flux across the planet and asserts that they agree.
+These temperature maps are validated by the energy balance of the planet; \vspec{} computes the incident and emergent flux across the planet and issues a warning if they are not in agreement.
 
@@
     \caption{
-        Example of a surface temperature map created by \vspec{}. The solar flux at Earth,
-        a Bond albedo of 0.3, and $\epsilon = 2\pi$ were used in this case.
+        Example of a surface temperature map created by the \vspec{} GCM, using values for the solar flux equivalent to Earth, a Bond albedo of 0.3, and $\epsilon = 2\pi$.
 Notice that the hottest point is offset from the sub-stellar point due to thermal inertia.
@@
 Where $\tilde{T}$ is the dimensionless temperature defined as $T/T_0$, or the temperature normalized by a local fiducial value.
-$\Phi$ is the angle from the substellar point, so we integrate if from $-\pi$ to $\pi$.
-This integral is done numerically in Figure \ref{fig:cowan_curves}, the optimal scheme depending on the value of $\epsilon$.
+$\Phi$ is the angle from the substellar point, so we numerically integrate it from $-\pi$ to $\pi$ using one of several different schemes depending on the value of $\epsilon$.
 For $0 < \epsilon < 1$ we adopt a reflecting initial value problem (IVP) scheme, where we set $\tilde{T}=1$ at $\Phi=\pi$, integrate to $-\pi$, then use the last value as the initial value for an integration from $-\pi$ to $\pi$.
 Note that for extremely small values of $\epsilon$ (e.g.
-$10^{-6}$) the integration is slow, and will fail if $\epsilon=0$; we recommended a lower limit of $\epsilon=10^{-4}$.
+$10^{-6}$), the integration is slow, and will fail if $\epsilon=0$; we recommended a lower limit of $\epsilon=10^{-4}$.
 When $1\le \epsilon < 10$ we solve the integral using a boundary value problem (BVP) scheme.
-For each of these two methods we employ the SciPy implementation (\texttt{solve\_ivp()} and \texttt{solve\_bvp()}, respectively).
-When $\epsilon \ge 10$ it is more efficient to use the analytic approximation given by \citet{cowan2011} Equation (A2).
+For each of these two methods we employ the SciPy \TJ{(cite scipy and the specific algorithms)} implementation (\texttt{solve\_ivp()} and \texttt{solve\_bvp()}, respectively).
+When $\epsilon \ge 10$ it is more efficient to use the analytic approximation given by Equation (A2) in \citet{cowan2011}.
+In Figure \ref{fig:cowan_curves} we show results for several different values of $\epsilon$.
 
@@
 These curves show the dimensionless temperature at the equator of a planet as a function of angle from substellar point $\Phi$ for thermal inertia $\epsilon$.
-Note that the $\epsilon=0$ curve in the original figure is replaced with $\epsilon=10^{-4}$ (see text), and we do not include additional curves for mid-latitudes as in the original figure.
-As $\epsilon$ increases heat is moved from the dayside onto the nightside, and the temperature maximum shifts away from the substellar point.
+Note that the $\epsilon=0$ curve in the original figure is replaced with $\epsilon=10^{-4}$ (see explanation in the text), and we do not include additional curves for mid-latitudes as in the original figure.
+As $\epsilon$ increases, heat is moved from the dayside onto the nightside and the temperature maximum shifts away from the substellar point.
     }
@@
 
-\subsection{Latitudinal Mixing}
-The treatment by \citet{cowan2011} includes no mixing in the latitudinal direction, and the temperature tends to 0 as you approach the poles.
-We add a parameter $\alpha$ that is defined by the ratio between the polar temperature and the equitorial temperature at any given longitude.
-Let $f(\lambda;\alpha)$ be the temperature as a function of latitude $\lambda$ at an arbitrary longitude.
-If the surface is in equilibrium, the energy balance requires the outgoing energy flux to be independent of $\alpha$:
+\subsection{Extending to 3 Dimensions}
+\TJ{lat mixing figure}
+
+The treatment by \citet{cowan2011} includes no mixing in the latitudinal direction, nor a vertical temperature profile; to allow for variations in pressure and inclination, we augment their treatment with several simple approximations for latitudinal and vertical temperature variations.
+
+We add a parameter $\alpha$ that is defined by the ratio between the polar temperature and the equatorial temperature at any given longitude.
+We define $f(\lambda;\alpha)$ to be the temperature as a function of latitude $\lambda$ at an arbitrary longitude; therefore if the surface is in equilibrium, the energy balance requires the outgoing energy flux to be independent of $\alpha$:
 \begin{equation}
@@
 \end{equation}
-Where $E_{\rm out}$ is the outbound energy flux of an infintesimal sliver of differential longitude.
-In the case of $\alpha=0$ (i.e. no mixing) we can treat each latitude value independently, so the equilibrium temperature is
+where $E_{\rm out}$ is the outbound energy flux of an infinitesimal sliver of differential longitude.
+In the case of $\alpha=0$ (i.e.
+no mixing) we can treat each latitude value independently, so the equilibrium temperature is
 \begin{equation}
@@
 \end{equation}
-
-Where $f(0;0)$ is the previously computed temperature at the equator (see Figure \ref{fig:cowan_curves}). Plugging this in to Equation (\ref{eq:e_out}) we see
+where $f(0;0)$ is the previously computed temperature at the equator (see Figure \ref{fig:cowan_curves}). Plugging this in to Equation (\ref{eq:e_out}), we see
 \begin{equation}
@@
 
-\subsection{Pressure and Temperature}
+%\subsection{Pressure and Temperature}
 Combining Equation (\ref{eq:temp_gen}) with the numerically-computed equatorial temperatures yields a 2D temperature map that covers the surface of the planet.

### 2024-07-30 src/tex/ms.tex
lat mixing panel in GCM section

@@
 Scattered points show the observation with simulated noise, considering a single wavelength channel from JWST/NIRSpec G395H binned to a resolving power of 270 and a target distance of 5 pc.
-        {\bf Right}: Stellar surface maps including translucent masks to show limb darkening and transit.
+        {\bf Left}: Stellar surface maps including translucent masks to show limb darkening and transit.
 The Lambertian star (top) does not exhibit any limb darkening, and its disk has constant surface brightness.
@@
     \caption{
-        Reproduction of \citet{cowan2011} Figure 1 using the methods described in the text.
+        {\bf Top}: Reproduction of \citet{cowan2011} Figure 1 using the methods described in the text.
 These curves show the dimensionless temperature at the equator of a planet as a function of angle from substellar point $\Phi$ for thermal inertia $\epsilon$.
@@
 As $\epsilon$ increases, heat is moved from the dayside onto the nightside and the temperature maximum shifts away from the substellar point.
+        {\bf Bottom}: Relative surface temperature as a function of latitude $\lambda$ for varying latitudinal mixing parameter $\alpha$.
+A value of 1 means the equatorial temperature as computed in the upper panel.
+Without mixing ($\alpha=0$) heat is concentrated at the Equator.
+However, as $\alpha$ increases, the equator cools and heat is redistributed towards the poles.
     }
@@
 \subsection{Extending to 3 Dimensions}
-\TJ{lat mixing figure}
 

### 2024-07-30 src/scripts/cowan_fig.py

@@ -17,7 +17,8 @@ colors = ['b','k','r']
 
 n_points = 100
 
-fig,ax = plt.subplots(1,1,figsize=(4,3))
+fig,axes = plt.subplots(2,1,figsize=(4,6))
+ax = axes[0]
 
 for e,l,m,c in zip(eps,label,modes,colors):
     lons, tsurf = ht.get_equator_curve(e,n_points,m)
@@ -31,5 +32,26 @@ ax.set_ylabel('$T/T_0$')
 ax.set_xticks([-0.5*np.pi,0,0.5*np.pi,np.pi, 1.5*np.pi,])
 ax.set_xticklabels(['$-\\pi/2$','$0$','$\\pi/2$','$\\pi$','$3\\pi/2$'])
 _=ax.legend()
+
+ax:plt.Axes = axes[1]
+alphas = [0.0,0.5,1.0]
+labels = ['0','\\frac{1}{2}','1']
+colors = ['xkcd:azure', 'xkcd:lavender', 'xkcd:purple']
+
+def get_curve(alpha:float, lat:np.ndarray) -> np.ndarray:
+    return (np.pi/4 * alpha + (1-alpha)*np.cos(lat))**(0.25)
+
+n_points = 1000
+for a,l,c in zip(alphas,labels,colors):
+    x = np.linspace(0,np.pi/2,n_points)
+    tsurf = get_curve(a,x)
+    ax.plot(x,tsurf,label=f'$\\alpha = {l}$',color=c)
+
+ax.set_xlabel('$\\lambda$')
+ax.set_ylabel('$T/T_{\\rm eq}$')
+ax.set_xticks([0,0.25*np.pi,0.5*np.pi])
+ax.set_xticklabels(['$0$','$\\pi/4$','$\\pi/2$'])
+_=ax.legend()
+
 fig.tight_layout()
 fig.savefig(OUTFIG)

### 2024-07-30 src/tex/ms.tex
Remove MIRECLE example

@@
 
-\subsection{MIRECLE Phase Curve}
-In this example we will look at the best-case target for the Mid-IR Exoplanet CLimate Explorer \citep[MIRECLE,][]{mandell2022}
-mission concept: Proxima Centauri b with no stellar variability. We will instead examine the expected noise produced by this observation
-and assess the detectability of the planet in a single orbit. \vspec{} uses the noise models built into PSG, allowing native support for
-noise due to the instrument, source, and astrophysical background.
+% \subsection{MIRECLE Phase Curve}
+% In this example we will look at the best-case target for the Mid-IR Exoplanet CLimate Explorer \citep[MIRECLE,][]{mandell2022}
+% mission concept: Proxima Centauri b with no stellar variability. We will instead examine the expected noise produced by this observation
+% and assess the detectability of the planet in a single orbit. \vspec{} uses the noise models built into PSG, allowing native support for
+% noise due to the instrument, source, and astrophysical background.
 
-We use \vspec{}'s built-in MIRECLE instrument template, which includes a 2 m aperture and observes from 1-18 $\mu m$ with $R=50$.
-We draw stellar and bulk planetary parameters from the NExScI Exoplanet Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/proxima\%20cen}}
-and assume $i=85^\circ$ and Earth density. The GCM we use has a Bond albedo of 0.3, $\epsilon=1.5$, $P_{\rm surf} = 1 \text{bar}$ and a 100 ppm \ce{CO2} atmosphere with a \ce{N2} background.
-We ran the simulation for one orbit with an 24 hour cadence. Because this is a best-case scenario, we assumed perfect knowledge of the stellar spectrum.
-We are left with the planetary spectrum and photon noise from the star.
+% We use \vspec{}'s built-in MIRECLE instrument template, which includes a 2 m aperture and observes from 1-18 $\mu m$ with $R=50$.
+% We draw stellar and bulk planetary parameters from the NExScI Exoplanet Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/proxima\%20cen}}
+% and assume $i=85^\circ$ and Earth density. The GCM we use has a Bond albedo of 0.3, $\epsilon=1.5$, $P_{\rm surf} = 1 \text{bar}$ and a 100 ppm \ce{CO2} atmosphere with a \ce{N2} background.
+% We ran the simulation for one orbit with an 24 hour cadence. Because this is a best-case scenario, we assumed perfect knowledge of the stellar spectrum.
+% We are left with the planetary spectrum and photon noise from the star.
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{figures/mirecle.pdf}
-    \caption{
-        Simulated spectra of Proxima Centauri b observed by MIRECLE. Noise and error bars based on 24 hour integrations with a 2 m aperture MIRECLE
-        space telescope.
-        }
-    \script{mirecle.py}
-    \label{fig:mirecle}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=0.5\textwidth]{figures/mirecle.pdf}
+%     \caption{
+%         Simulated spectra of Proxima Centauri b observed by MIRECLE. Noise and error bars based on 24 hour integrations with a 2 m aperture MIRECLE
+%         space telescope.
+%         }
+%     \script{mirecle.py}
+%     \label{fig:mirecle}
+% \end{figure}
 

### 2024-07-30 src/tex/ms.tex
Cite scipy ODE solvers

@@
 When $1\le \epsilon < 10$ we solve the integral using a boundary value problem (BVP) scheme.
-For each of these two methods we employ the SciPy \TJ{(cite scipy and the specific algorithms)} implementation (\texttt{solve\_ivp()} and \texttt{solve\_bvp()}, respectively).
+For each of these two methods we employ the SciPy \citep{virtanen2020} implementation: \texttt{solve\_ivp()}, employing a 5th order Runga-Kutta scheme \citep{dormand1980} and \texttt{solve\_bvp()}, based on an earlier MATLAB implementation \citep{kierzenka2001}, respectively.
 When $\epsilon \ge 10$ it is more efficient to use the analytic approximation given by Equation (A2) in \citet{cowan2011}.

### 2024-07-30 src/tex/ms.tex
Describe binning algorithm better

@@
 The binning algorithm takes a pair of high-resolution flux and wavelength arrays and resamples the flux to a lower-resolution wavelength axis.
-This algorithm currently assumes that each point on the new wavelength axis is the central value of the wavelength channel, but future versions will treat each spectral channel \TJ{Ask Geronimo about the particulars of PSG's binning/wavelength axis creation}.
-\TJ{(I have never liked this, should I rewrite it before publication?)}.
+PSG specifies the central wavelength of each channel, and so this algorithm defines bin edges to be $\lambda_{\rm cen} \pm \frac{\lambda_{\rm cen}}{R}$.
 

### 2024-07-30 src/tex/ms.tex
Add my own affiliations
Change some wording in examples

@@
 \author{Ted Johnson}
+\affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
+\affiliation{Center for Research and Exploration in Space Science and Technology, NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
 \affiliation{NASA Goddard Space Flight Center \\
@@
 \author{Avi M. Mandell}
+\affiliation{NASA Goddard Space Flight Center \\
+8800 Greenbelt Rd \\
+Greenbelt, MD 20771, USA}
+
+\author{Ashraf Dhahbi}
+\affiliation{NASA Goddard Space Flight Center \\
+8800 Greenbelt Rd \\
+Greenbelt, MD 20771, USA}
+
+\author{Tobi Hammond}
 \affiliation{NASA Goddard Space Flight Center \\
@@
 In the case that the stellar surface is not homogeneous, the transit signal is contaminated by the star.
-\citet{moran2023} first observed this effect using JWST from super-Earth GJ 486b.
+We will reproduce results by \citet{moran2023}, who observed this effect with JWST/NIRSpec during a transit of super-Earth GJ 486b.
 
@@
 
-We will do three \vspec{} simulations:
+We will do two \vspec{} simulations:
 \begin{enumerate}
-    \item Bare rock, no spots.
-    \item 1 bar \ce{H2O} atmosphere, no spots.
-    \item Bare rock, 2.5\% coverage by 2700 K spots
+    \item 1 bar \ce{H2O} isothermal atmosphere, no spots. \label{case:1barh2o}
+    \item No atmosphere, 2.5\% coverage by 2700 K spots \label{case:no_atmo}
 \end{enumerate}
-
-We expect in case 1 to observe a flat spectrum, and in case 2 to observe additional
-absorption from the \ce{H2O} atmosphere. Note also that we disabled \ce{H2O}-\ce{H2O} collision-induced absorption (CIA) for
+In Case \ref{case:1barh2o} we expect to see an absorption spectrum characteristic of \ce{H2O}.
+Case \ref{case:no_atmo} is more complicated, though in the absence of spots we would expect a flat spectrum.
+Note also that we disabled \ce{H2O}-\ce{H2O} collision-induced absorption (CIA) for
 these simulations because the effect was not considered in the original analysis of GJ 486b.
@@
 light lost to absorbers in a planetary atmosphere. Figure \ref{fig:moran_transit} shows us that, in this case, we observe false absorption;
-this is due to an increase in the spotted fraction of the surface relative to the quiet photosphere. However, if a spot was blocked by
+this is due to a temporary increase in the spotted fraction of the surface relative to the quiet photosphere. However, if a spot was blocked by
 the planet we might see similar features inverted to mimic emission.
@@
 contamination because of the slow rotation rate of the host star \citep[approximately 1/80$^{\text{th}}$ the orbital frequency,][]{cloutier2021}.
-In this example we use \vspec{} to demonstrate that this is a reasonable assumption.
+In this example we simulate this system using \vspec{} and find that this is a reasonable assumption.
 

### 2024-07-30 src/tex/ms.tex
Update section listing and remove old comments

@@
     a set of codes developed alongside \vspec{}, including the stellar surface model \texttt{vspec-vsm}, the stellar spectral grid interpolation code
-    Gridpolator, and a Python interface for PSG \texttt{pypsg}.
+    GridPolator, and a Python interface for PSG \texttt{pypsg}.
 \end{abstract}
@@
 a stellar spectra grid interpolation code Gridpolator\footnote{\url{https://github.com/VSPEC-collab/GridPolator}}, and a Python interface for
-the Planetary Spectrum Generator \citep[PSG,][]{villanueva2018}, \texttt{pypsg}. All of these codes together allow \vspec{} to account for
+the Planetary Spectrum Generator \citep[PSG,][]{villanueva2018}, \texttt{pypsg}\footnote{\url{https://github.com/tedjohnson12/pypsg}}. All of these codes together allow \vspec{} to account for
 3D planetary atmospheres, time-resolved effects of both planet and star, many geometries including transit and eclipse, and realistic noise modeling.
 
-\TJ{Change this later-->}
-In this paper we will describe the science behind \vspec{} and demonstrate examples of its use.
-Section \ref{sec:star} describes the stellar model and the available sources of variability.
-Section \ref{sec:pl_model} discusses modeling the planetary atmosphere, including PSG/GlobES and the
-built-in GCM. In Section \ref{sec:examples} we provide examples of \vspec{} use cases and demonstrate
-its value to the exoplanet community. Finally, in Section \ref{sec:conclusion} we will discuss the future of the code and issues it might have.
+In this paper we will describe the science behind all of these codes and demonstrate examples of their use.
+Section \ref{sec:vspec} describes \vspec{} and goes into some detail about setup, use, and the code's internals.
+In Section \ref{sec:pypsg} we describe \texttt{pypsg} as an interface for the Planetary Spectrum Generator.
+Section \ref{sec:star} describes \vspec{}'s stellar model, \texttt{vspec-vsm}.
+We discuss the spectral model interpolation code GridPolator in Section \ref{sec:gridpolator}.
+Section \ref{sec:vspec-gcm} is dedicated to the simple climate model that \vspec{} uses to produce realistic phase curves in the absence of a sophisticated GCM.
+Section \ref{sec:examples} demonstrates a few simple \vspec{} use cases.
+Finally, in Section \ref{sec:conclusion} we will discuss the future of the code and issues it might have.
 
@@
 PSG\footnote{\url{https://psg.gsfc.nasa.gov/}} \citep{villanueva2018} is a powerful radiative transfer tool that is ubiquitous in the exoplanet and solar system atmosphere fields.
-Either through its web graphical interface, its public API, or the local version available through Docker, PSG allows users to simulate observations of planets, comets, and moons with a variety of geometries and realistic noise models \TJ{probably someone like Geronimo should write this part}.
+Either through its web graphical interface, its public API, or the local version available through Docker, PSG allows users to simulate observations of planets, comets, and moons with a variety of geometries and realistic noise models.
 
@@
 \section{Simplified Climate Model}
+\label{sec:vspec-gcm}
 \vspec{} allows users to run simulations with any climate model they desire (so long as it can be converted into the \texttt{pypsg.PyGCM} format) and even comes with utilities to work with PSG/GlobES binary files, ExoCAM \citep{wolf2022}, WACCM \citep{marsh2013}, and ExoPlaSim \citep{paradise2022} without any custom code from the user.
@@
 
-\section{Conclusion \label{sec:conclusion}}
-\TJ{Talk about what we can do next.}
+\section{Conclusion}
+\label{sec:conclusion}
+We present a suite of codes to model observations of exoplanets with 3D atmospheres orbiting variable stars.
+The main package, \vspec{}, produces combined-light phase curves by interfacing with an ensemble of support packages whose responsibilities range from grid interpolation to interfacing with PSG.
+These codes are all open-source and available to the public on GitHub and the Python package index, PyPI.
+
+While development is ongoing, these codes are all stable.
+Future work on \vspec{} will concern improving performance so that it can be used to produce phase-curve grids for retrievals.
+A goal of \texttt{vspec-vsm} development should be to create a common API between it and more physics-based stellar models so that \vspec{} can interface with those models as well.
+GridPolator is currently in the process of adding support for more built-in stellar grids, these more complex and covering a larger range of spectral types that the current \vspec{} grid.
+\texttt{pypsg} has the most to offer the exoplanet community as a stand-alone package, but currently lacks support for much of the data that PSG is capable of producing.
+Its development so far has focused around the \vspec{} use case, so future work will center around making it more useful to the majority of PSG users.
+
+\section{Acknowledgements}
+The authors would like to thank Peter Hauschildt for generating a custom grid of PHOENIX models.
 

### 2024-08-01 src/tex/ms.tex
Editing session with Avi August 1

@@
 % \title{The \vspec{} code: Spectroscopic phase curves of 3D exoplanet models in the presence of stellar variability}
-\title{\vspec{} and Friends: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.}
+\title{The \vspec{} Collection: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.}
 
-\author{Ted Johnson}
+\author[0000-0002-1570-2203]{Ted M Johnson}
 \affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
@@
 \begin{abstract}
-    We present the Variable Star Phase Curve (\href{https://github.com/VSPEC-collab/VSPEC} \vspec{}) code,
-    a python package to simulate combined-light spectroscopic observations of exoplanets in the presence of stellar variability and inhomogeneity.
-    \vspec{} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG GlobES) application along with a custom-built
-    stellar model based on an existing grid of stellar photosphere models to produce spectroscopic light curves of the planet-host system.
-    \vspec{} can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements,
-    and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs. We additionally present
-    a set of codes developed alongside \vspec{}, including the stellar surface model \texttt{vspec-vsm}, the stellar spectral grid interpolation code
-    GridPolator, and a Python interface for PSG \texttt{pypsg}.
+    We present the Variable Star PhasE Curve (VSPEC) Collection, a set of Python packages for simulating combined-light spectroscopic observations of 3-dimensional exoplanet atmospheres in the presence of stellar variability and inhomogeneity.
+\vspec{} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG/GlobES) application along with a custom-built multi-component time-variable stellar model based on an user-defined grid of stellar photosphere models to produce spectroscopic light curves of the planet-host system.
+\vspec{} can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements, and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs.
+We additionally present a set of codes developed alongside the core \vspec{} modules, including the stellar surface model generator \texttt{vspec-vsm}, the stellar spectral grid interpolation code GridPolator, and a Python interface for PSG \texttt{pypsg}.
 \end{abstract}
@@
 flares).
-As an example, recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between atmospheric absorption by water and water-rich spots on the stellar surface, and similar effects were seen in the transit spectrum of LHS 1140b \citet{cadieux2024}.
+As an example, recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between atmospheric absorption by water and water-rich spots on the stellar surface, and similar effects were seen in the transit spectrum of LHS 1140b \citet{cadieux2024} and many other recent transit measurements \TJ{(more examples)}.
 This ``transit light source effect'' \citep[TLS,][see also \citet{apai2018,barclay2021,garcia2022,barclay2023}]{rackham2018} occurs when the region of the stellar surface occulted by the transiting planet is not representative of the disk-integrated spectrum.
 
-Stellar contamination may also cause errors in the extraction of thermal phase curve spectroscopy if the stellar spectrum changes significantly over the period of the planet's orbit; if the variations are non-linear but only linear interpolation between subsequent secondary eclipses is used to remove the stellar contribution, the results planetary flux measurements will be contaminated.
+Stellar contamination may also cause errors in the extraction of thermal phase curve spectroscopy if the stellar spectrum changes significantly over the period of the planet's orbit; if the variations are non-linear but only linear interpolation between subsequent secondary eclipses is used to remove the stellar contribution, the resulting planetary flux measurements will not be exclusively representative of the planet's emission.
 This will be particularly important for observations of planets around more variable stars such as active M-stars or later-type evolved host stars.
@@
 
-To adequately prepare for future observations and future missions it will be necessary to demonstrate a method to mitigate these effects.
+To adequately prepare for future observations and future missions, it will be necessary to demonstrate methods to mitigate these effects.
 This task requires a flexible tool that combines models of exoplanet atmospheres and stellar variability in a robust way.
-In this paper we present \vspec{}: Variable Star PhasE Curve\footnote{\url{https://github.com/VSPEC-collab/VSPEC}},
-an open-source Python 3 package to simulate observations of exoplanet systems with variable host stars.
-\vspec{} itself is merely an interface that pulls together a variable star surface model \texttt{vspec-vsm}\footnote{\url{https://github.com/VSPEC-collab/vspec-vsm}},
-a stellar spectra grid interpolation code Gridpolator\footnote{\url{https://github.com/VSPEC-collab/GridPolator}}, and a Python interface for
-the Planetary Spectrum Generator \citep[PSG,][]{villanueva2018}, \texttt{pypsg}\footnote{\url{https://github.com/tedjohnson12/pypsg}}. All of these codes together allow \vspec{} to account for
-3D planetary atmospheres, time-resolved effects of both planet and star, many geometries including transit and eclipse, and realistic noise modeling.
+In this paper we present the \vspec{} (Variable Star PhasE Curve\footnote{\url{https://github.com/orgs/VSPEC-collab/repositories}}) Collection, a collection of open-source Python 3 packages to simulate observations of exoplanet systems with variable host stars.
+The \vspec{}\footnote{\url{https://github.com/VSPEC-collab/VSPEC}} package itself is merely an interface that pulls together a variable star surface model \texttt{vspec-vsm}\footnote{\url{https://github.com/VSPEC-collab/vspec-vsm}}, a stellar spectra grid interpolation code Gridpolator\footnote{\url{https://github.com/VSPEC-collab/GridPolator}}, and a Python interface for the Planetary Spectrum Generator \citep[PSG,][]{villanueva2018}, \texttt{pypsg}\footnote{\url{https://github.com/tedjohnson12/pypsg}}.
+All of these codes together allow \vspec{} to account for 3D planetary atmospheres, time-resolved effects of both planet and star, multiple geometries including transit and secondary eclipse, and realistic noise modeling.
 
-In this paper we will describe the science behind all of these codes and demonstrate examples of their use.
+In this paper we will describe the science motivation behind all of these codes, the details of each package and module in the \vspec{} collection, and demonstrate examples of their use.
 Section \ref{sec:vspec} describes \vspec{} and goes into some detail about setup, use, and the code's internals.
@@
 We discuss the spectral model interpolation code GridPolator in Section \ref{sec:gridpolator}.
-Section \ref{sec:vspec-gcm} is dedicated to the simple climate model that \vspec{} uses to produce realistic phase curves in the absence of a sophisticated GCM.
+Section \ref{sec:vspec-gcm} is dedicated to the simple climate model that \vspec{} uses to produce realistic phase curves in the absence of a user-supplied output file from a more sophisticated Global Circulation Model (GCM).
 Section \ref{sec:examples} demonstrates a few simple \vspec{} use cases.
@@
 The star is then evolved at the output cadence, at each epoch generating a set of key-value pairs that describe the surface temperature and coverage fractions of the portion of the stellar disk visible to the observer.
-These \teff\ s and coverage fractions are fed into GridPolator to produce composite stellar spectra.
+These \teff s and coverage fractions are fed into GridPolator to produce composite stellar spectra.
 
-\subsubsection{Reflected light}
-\vspec{} also produces a composite spectrum of the portion of the star visible to the planet to accurately incorporate the variable star into the reflected light spectrum.
+\subsection{Reflected light}
+\vspec{} then produces a composite spectrum of the portion of the star visible to the planet to accurately incorporate the time-dependent stellar flux into the reflected-light spectrum.
 To compute the total reflected light in the output files, \vspec{} first divides the reflected light flux from PSG by the PSG stellar model to obtain the apparent albedo (I/F).
-This albedo is then multiplied
-by the composite spectrum (facing the planet) to obtain total reflected light flux.
+This albedo is then multiplied by the composite stellar spectrum (facing the planet) to obtain the total reflected flux.
 
-\subsubsection{Noise}
-Alongside the planetary radiance files, PSG returns a file giving the noise profile for each epoch broken down by source.
+\subsection{Simulated Noise}
+Alongside the planetary radiance files, PSG returns a file giving the analytic noise estimate for each epoch broken down by source.
 We assume that the total noise is the quadrature sum of its constituent parts and that the \texttt{source} column (i.e.
@@
 
-\subsubsection{Transit and Eclipse}
-Users must be careful when using \vspec{} to simulate transit and eclipse measurements because of the vast timescale differences between those events
-and the typical phase variations of a planet -- a sparsely sampled phase curve that happens to include one epoch of transit will not adequately sample the lightcurve accurately during a transit or eclipse because the interpolator knows no better than to assume a linear interpolation scheme.
+\subsection{Transit and Eclipse}
+Users must be careful when using \vspec{} to simulate transit and eclipse measurements because of the vast timescale differences between those events and the typical phase variations of a planet -- a sparsely sampled phase curve that happens to include one epoch of transit will not adequately sample the lightcurve accurately during a transit or eclipse because the interpolator knows no better than to assume a linear interpolation scheme.
 It is recommended that users use a high cadence for observations of transits and eclipses so that each event will be sampled by more than one point.
 
-In the case of a transiting geometry, \vspec{} computes the spectrum of the occulted portion of the star in order to properly simulate the
-transit light source effect \citep[TLS][]{rackham2018}.
+In the case of a transiting geometry, \vspec{} computes the spectrum of the occulted portion of the star in order to properly simulate the transit light source effect \citep[TLS][]{rackham2018}.
 This spectrum is the flux blocked by the planet in the case that it is a solid occulting circle, and therefore produces a flat transmission spectrum.
@@
 \end{equation}
-where $F_\lambda$ is the flux of the variable star surface occulted by the planet with the atmosphere considered. $F_{{\rm rock}, \lambda}$ is the
-flux of the surface occulted by a bare rock with the planet's radius, $F_{{\rm PSG,} \lambda}$ is the flux computed by PSG to be lost
+where $F_\lambda$ is the flux of the variable star surface occulted by the planet with the atmosphere considered.
+$F_{{\rm rock}, \lambda}$ is the flux of the surface occulted by a bare rock with the planet's radius, $F_{{\rm PSG,} \lambda}$ is the flux computed by PSG to be lost
 due to occultation and transmission, and $R_*$ and $R_p$ are the stellar and planetary radii, respectively.
 
-In the case of a total eclipse of the planet, the planetary thermal and reflected contributions are set to zero flux. However, in the case of a partial eclipse,
-\vspec{} computes the fraction of the planet that is visible to the observer and reduces both thermal and reflected flux accordingly. In this case
-\vspec{} treats the planet homogeneously; however, future work could utilize PSG/GlobES's ability to return a hypercube of spatially resolved
-spectra in order to construct planetary spectra only from regions of the planet that are not eclipsed.
+In the case of a total eclipse of the planet, the planetary thermal and reflected contributions are set to zero flux.
+However, in the case of a partial eclipse (large impact parameter), \vspec{} computes the fraction of the planet that is visible to the observer and reduces both thermal and reflected flux accordingly.
+In this case \vspec{} treats the planet homogeneously; however, future work could utilize PSG/GlobES's ability to return a hypercube of spatially resolved spectra in order to construct planetary spectra only from regions of the planet that are not eclipsed.
 
@@
 The \vspec{} stellar model is designed in a modular fashion to allow for both simple and complex behaviors.
-The model is available as the Python package \texttt{vspec-vsm}.
+The model is available as the Python package \texttt{vspec-vsm}, which is based on the earlier stellar spot model \texttt{spotty} \citep{barclay2021}.
 It is spectral-model agnostic, meaning that it relies on another package (GridPolator, see Section \ref{sec:gridpolator}) to compute the actual variable spectra.
@@
 To consider this effect in the faculae lightcurve, we compute the fraction of each facula's projected area -- the area on the disk it would occupy as seen by the observer -- that is occupied by the hot wall versus the cool floor.
-Mathematically, this is no different from finding the area of the intersection of two overlapping ellipses:
+Mathematically, this is no different from finding the area of the intersection of two overlapping ellipses, and we use this conceptualization to calculate the contributions of each portion, as described below.
 
@@
 circles).
-This problem is solved for arbitrary circles by \citet{weisstein2004}.
-The details of their derivation are not relevant to this text, but we use their result to create a function $f(d,R)$ that computes the fraction of a circle with radius $R$ and center $(d,0)$ that is inside the unit circle.
-For a facula with radius $r$, wall height $h$, and angle from disk center $theta$, we find that the effective area of the cool floor is:
+This problem is solved for arbitrary circles by \citet{weisstein2004}; the details of their derivation are not relevant to this text, but we use their result to create a function $f(d,R)$ that computes the fraction of a circle with radius $R$ and center $(d,0)$ that is inside the unit circle.
+For a facula with radius $r$, wall height $h$, and angle from disk center $\theta$, we find that the effective area of the cool floor is:
 \begin{equation}

### 2024-08-01 src/tex/ms.tex
Add more TLS citations

@@
 
+\correspondingauthor{Ted M Johnson}
+\email{ted.johnson@unlv.edu}
+
 \author[0000-0002-1570-2203]{Ted M Johnson}
@@
 \affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
-% Also CRESST, SURA, SEEC!
 
@@
 flares).
-As an example, recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between atmospheric absorption by water and water-rich spots on the stellar surface, and similar effects were seen in the transit spectrum of LHS 1140b \citet{cadieux2024} and many other recent transit measurements \TJ{(more examples)}.
+As an example, recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between atmospheric absorption by water and water-rich spots on the stellar surface, and similar effects were seen in the transit spectrum of LHS 1140b \citep{cadieux2024} and many other recent transit measurements iwth JWST \citep[e.g.,][]{lim2023,may2023,fournier-tondreau2024}.
 This ``transit light source effect'' \citep[TLS,][see also \citet{apai2018,barclay2021,garcia2022,barclay2023}]{rackham2018} occurs when the region of the stellar surface occulted by the transiting planet is not representative of the disk-integrated spectrum.
 
-Stellar contamination may also cause errors in the extraction of thermal phase curve spectroscopy if the stellar spectrum changes significantly over the period of the planet's orbit; if the variations are non-linear but only linear interpolation between subsequent secondary eclipses is used to remove the stellar contribution, the resulting planetary flux measurements will not be exclusively representative of the planet's emission.
+Stellar contamination may also cause errors in the extraction of thermal phase curve spectroscopy if the stellar spectrum changes significantly over the period of the planet's orbit; if the variations are non-linear, but only linear interpolation between subsequent secondary eclipses is used to remove the stellar contribution, the resulting planetary flux measurements will not be exclusively representative of the planet's emission.
 This will be particularly important for observations of planets around more variable stars such as active M-stars or later-type evolved host stars.

### 2024-08-01 src/scripts/cowan_fig.py
Rasterize maps

@@ -27,7 +27,7 @@ for e,l,m,c in zip(eps,label,modes,colors):
     tsurf = np.concatenate([tsurf[index:],tsurf[:index]])
     ax.plot(lons,tsurf,color=c,label=f'$\\epsilon = {l}$')
 
-ax.set_xlabel('$\\Phi$')
+ax.set_xlabel('Longitude')
 ax.set_ylabel('$T/T_0$')
 ax.set_xticks([-0.5*np.pi,0,0.5*np.pi,np.pi, 1.5*np.pi,])
 ax.set_xticklabels(['$-\\pi/2$','$0$','$\\pi/2$','$\\pi$','$3\\pi/2$'])
@@ -47,7 +47,7 @@ for a,l,c in zip(alphas,labels,colors):
     tsurf = get_curve(a,x)
     ax.plot(x,tsurf,label=f'$\\alpha = {l}$',color=c)
 
-ax.set_xlabel('$\\lambda$')
+ax.set_xlabel('Latitude')
 ax.set_ylabel('$T/T_{\\rm eq}$')
 ax.set_xticks([0,0.25*np.pi,0.5*np.pi])
 ax.set_xticklabels(['$0$','$\\pi/4$','$\\pi/2$'])

### 2024-08-01 src/scripts/cowan_gcm.py

@@ -36,7 +36,7 @@ fig = plt.figure(figsize=(4,3))
 proj = ccrs.Robinson(central_longitude=0)
 ax = fig.add_subplot(projection=proj)
 
-im = ax.pcolormesh(lons.to_value(u.deg),lats.to_value(u.deg),data.to_value(u.K),cmap='gist_heat',transform=ccrs.PlateCarree())
+im = ax.pcolormesh(lons.to_value(u.deg),lats.to_value(u.deg),data.to_value(u.K),cmap='gist_heat',transform=ccrs.PlateCarree(),rasterized=True)
 gl = ax.gridlines(crs=ccrs.PlateCarree(),draw_labels=True,
     color='grey', alpha=0.8, linestyle='--')
 gl.top_xlabels = False

### 2024-08-01 src/scripts/ld_transit.py

@@ -20,13 +20,13 @@ pypsg.docker.set_url_and_run()
 lam_model = VSPEC.ObservationModel.from_yaml(LAMBERT_CONFIG)
 sol_model = VSPEC.ObservationModel.from_yaml(SOLAR_CONFIG)
 
-lam_model._build_star()
-sol_model._build_star()
+# lam_model._build_star()
+# sol_model._build_star()
 
-# lam_model.build_planet()
-# lam_model.build_spectra()
-# sol_model.build_planet()
-# sol_model.build_spectra()
+lam_model.build_planet()
+lam_model.build_spectra()
+sol_model.build_planet()
+sol_model.build_spectra()
 
 lam_data = VSPEC.PhaseAnalyzer.from_model(lam_model)
 sol_data = VSPEC.PhaseAnalyzer.from_model(sol_model)
@@ -65,13 +65,15 @@ lam_model.star.plot_surface(0*u.deg,0*u.deg,ax=lam_ax,
                             orbit_radius=lam_model.params.planet.semimajor_axis,
                             radius=lam_model.params.planet.radius,
                             phase=179.3*u.deg,
-                            inclination=lam_model.params.system.inclination
+                            inclination=lam_model.params.system.inclination,
+                            rasterize=True
                             )
 sol_model.star.plot_surface(0*u.deg,0*u.deg,ax=sol_ax,
                             orbit_radius=sol_model.params.planet.semimajor_axis,
                             radius=sol_model.params.planet.radius,
                             phase=179.3*u.deg,
-                            inclination=sol_model.params.system.inclination
+                            inclination=sol_model.params.system.inclination,
+                            rasterize=True
                             )
 
 # inax = lam_ax.inset_axes([-0.99,-0.15,0.3,0.3],transform=ccrs.PlateCarree())

### 2024-08-01 src/scripts/surface_map_and_lc.py

@@ -47,7 +47,7 @@ pl_orbit = 0.05*u.AU
 inclination = 89.8*u.deg
 phase = 180.4*u.deg
 
-model.star.plot_surface(lat0, lon0, ax1)
+model.star.plot_surface(lat0, lon0, ax1,rasterize=True)
 
 # LC
 

### 2024-08-02 src/tex/ms.tex
Add pypsg code listing

@@
 
-\author[0000-0002-1570-2203]{Ted M Johnson}
+% \author[0000-0002-1570-2203]{Ted M Johnson}
+\author{Ted M Johnson}
 \affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
@@
 be written:
-\begin{lstlisting}[label={ls:yaml}]
+\begin{lstlisting}[label={ls:yaml},caption={\vspec{} YAML Configuration}]
 system:
@@
 parameters in Listing \ref{ls:yaml} would be written:
-\begin{lstlisting}[language=Python]
+\begin{lstlisting}[language=Python, caption={\vspec{} Python configuration}]
 from astropy import units as u
@@
 \vspec{} uses a Python interface for the PSG API system called \texttt{pypsg}, which is a general stand-alone library built for any Python-based API call to PSG.
-\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object.Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
+\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object.
+Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
 Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package.
-This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key
-could be committed to a public repository.
+This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key could be committed to a public repository.
 
@@
 \texttt{pypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.
+
+Listing \ref{ls:pypsg} shows a minimal example that uses \texttt{pypsg} to make a call to PSG/GlobES.
+Here \texttt{pypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance.
+This configuration object is then passed to an \texttt{APICall} instance.
+When the API call is made, the contents of the \texttt{PyConfig} are included in the body of the POST request in the \texttt{file} field.
+PSG reads this the same as it would if we had written those contents to a file \texttt{config.txt} and passed the \texttt{--data-urlencode file@config.txt} option to \texttt{cURL}, as are the instructions in the PSG handbook \citep[pp.
+162-163]{villanueva2022}.
+
+\begin{lstlisting}[
+    basicstyle=\footnotesize,language=Python,label={ls:pypsg},caption={\texttt{pypsg} Example}
+]
+    import pypsg
+    from astropy import units as u
+    from netCDF4 import Dataset
+
+    # This next line only needs to be run once per user.
+    # These settings are saved to `~/.pypsg/settings.json`
+    pypsg.settings.save_setting(
+        api_key='my_api_key',
+        url='https://psg.gsfc.nasa.gov/api.php'
+    )
+
+    gcm_data = Dataset('path/to/gcm.nc')
+
+    cfg = pypsg.cfg.PyConfig(
+        target=pypsg.cfg.Target(
+            object='Exoplanet',
+            name='proxima-cen-b'
+            ...
+        ),
+        geometry=pypsg.cfg.Observatory(...),
+        generator=pypsg.cfg.Generator(...),
+        telescope=pypsg.cfg.SingleTelescope(...)
+        noise=pypsg.cfg.CCD(...)
+        gcm=pypsg.globes.waccm_to_pygcm(
+            gcm_data,
+            itime=0,
+            molecules=['H2O','CO2'],
+            aerosols=['Water','WaterIce'],
+            mean_molecular_mass=48.0
+        )
+
+    )
+
+    caller = pypsg.APICall(
+        cfg=cfg,
+        app='globes',
+        output_type='rad'
+    )
+    response = caller()
+
+    print(response.rad['Wave/freq','proxima-cen-b'])
+
+    -------------------
+
+    Wave/freq  proxima-cen-b
+        um      W / (um m2) 
+    ----------- -------------
+            1.0   1.81314e-20
+           1.02    3.1926e-20
+         1.0404   4.95983e-19
+       1.061208    5.0892e-19
+     1.08243216   3.79632e-19
+            ...           ...
+     16.3163953   7.66301e-20
+     16.6427232   7.66245e-20
+    16.97557767   7.54221e-20
+    17.31508922   7.49258e-20
+    17.66139101   7.35507e-20
+    Length = 146 rows
+
+\end{lstlisting}
 

### 2024-08-02 src/tex/ms.tex
Fix periastron typo

@@
     inclination: 89.7 deg
-    phase_of_periasteron: 0 deg
+    phase_of_periastron: 0 deg
 \end{lstlisting}
@@
         inclination=89.7*u.deg,
-        phase_of_pariasteron=0*u.deg
+        phase_of_periastron=0*u.deg
     )

### 2024-08-02 src/scripts/granule_lc.py

@@ -58,7 +58,7 @@ planet = params.PlanetParameters.std(init_phase=180*u.deg,init_substellar_lon=0*
 system = params.SystemParameters(
     distance=1.3*u.pc,
     inclination=30*u.deg,
-    phase_of_periasteron=0*u.deg
+    phase_of_periastron=0*u.deg
 )
 observation = params.ObservationParameters(
     observation_time=3*u.day,

### 2024-08-02 src/scripts/kempton23.py

@@ -69,7 +69,7 @@ planet_params = params.PlanetParameters(
 system_params = params.SystemParameters(
     distance=14.6427*u.pc,
     inclination=inclination,
-    phase_of_periasteron=0*u.deg
+    phase_of_periastron=0*u.deg
 )
 
 star_dict = {

### 2024-08-02 src/scripts/mirecle.py

@@ -74,7 +74,7 @@ planet_params = params.PlanetParameters(
 system_params = params.SystemParameters(
     distance=1.3*u.pc,
     inclination=inclination,
-    phase_of_periasteron=0*u.deg
+    phase_of_periastron=0*u.deg
 )
 
 

### 2024-08-02 src/scripts/moran23.py

@@ -127,7 +127,7 @@ spotted_rock_planet = params.PlanetParameters(
 system_params = params.SystemParameters(
     distance=8.07*u.pc,
     inclination=inclination,
-    phase_of_periasteron=0*u.deg
+    phase_of_periastron=0*u.deg
 )
 
 

### 2024-08-02 src/scripts/kempton23.py
Fix cutoff ylabel in kempton figure

@@ -222,9 +222,11 @@ for edge,ax in zip(bin_edges,axes):
     ax.plot(time,(spotted_lc),c='xkcd:lavender',label='Spotted')
     ax.text(0.55,0.7,f'{w1:.1f} - {w2:.1f}',transform=ax.transAxes)
     ax.set_ylim(-100,700)
-fig.subplots_adjust(hspace=0,wspace=0)
+
 axes[0].legend(loc='upper left')
 axes[-1].set_xlabel('Time (hour)')
-_ = axes[n_ax//2].set_ylabel('Planet Flux (ppm)')
+_ = axes[n_ax//2].set_ylabel('Apparent Planet Flux (ppm)')
 
+fig.tight_layout()
+fig.subplots_adjust(hspace=0,wspace=0)
 fig.savefig(outfile)

### 2024-08-02 src/tex/ms.tex
Add line numbers and docs urls

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[twocolumn,linenumbers]{aastex631}
 \usepackage{showyourwork}
@@
 % \def\vspec#{\texttt{VSPEC}}
+
+\newcommand{\urldocsvspec}{\url{https://vspec-collab.github.io/VSPEC/0.6.0/index.html}}
+\newcommand{\urldocsvspecvsm}{\url{https://vspec-vsm.readthedocs.io/en/latest/}}
+\newcommand{\urldocspypsg}{\url{https://tedjohnson12.github.io/pypsg/}}
+\newcommand{\urldocsgridpolator}{\url{https://vspec-collab.github.io/GridPolator/}}
 
@@
 The flow of \vspec{}'s code is broken into three parts: i) read in the configuration, ii) compute planetary spectra, and iii) compute stellar spectra.
+Detailed API documentation, configuration parameter descriptions, and examples can be found on the \vspec{} documentation page\footnote{\urldocsvspec}.
 \subsection{Configuring \vspec{}}
@@
 162-163]{villanueva2022}.
+More details on use of this library including its API and examples can be found on the \texttt{pypsg} documentation page\footnote{\urldocspypsg}.
 
@@
 a container for objects representing sources of variability such as \texttt{StarSpot}, \texttt{Facula}, \texttt{StelarFlare}, and \text{Granulation}, which are described below.
+For a complete description of the code, including examples, see the \texttt{vspec-vsm} documentation page\footnote{\urldocsvspecvsm}.
 

### 2024-08-02 src/tex/ms.tex
Fix weird bug with url

@@
 \usepackage{listings}
+\usepackage{hyperref}
 
@@
 
-\newcommand{\urldocsvspec}{\url{https://vspec-collab.github.io/VSPEC/0.6.0/index.html}}
 \newcommand{\urldocsvspecvsm}{\url{https://vspec-vsm.readthedocs.io/en/latest/}}
@@
 The flow of \vspec{}'s code is broken into three parts: i) read in the configuration, ii) compute planetary spectra, and iii) compute stellar spectra.
-Detailed API documentation, configuration parameter descriptions, and examples can be found on the \vspec{} documentation page\footnote{\urldocsvspec}.
+See the \vspec{} documentation page\footnote{\href{https://vspec-collab.github.io/VSPEC/0.6.0/index.html}{https://vspec-collab.github.io/VSPEC/0.6.0/index.html}} for detailed API documentation.
 \subsection{Configuring \vspec{}}
@@
 be written:
-\begin{lstlisting}[label={ls:yaml},caption={\vspec{} YAML Configuration}]
+\begin{lstlisting}[
+    label={ls:yaml},
+    caption={\vspec{} YAML Configuration}
+]
 system:

### 2024-08-05 src/tex/ms.tex
change pypsg name

@@
 \newcommand{\urldocsvspecvsm}{\url{https://vspec-vsm.readthedocs.io/en/latest/}}
-\newcommand{\urldocspypsg}{\url{https://tedjohnson12.github.io/pypsg/}}
+\newcommand{\urldocslibpypsg}{\url{https://tedjohnson12.github.io/libpypsg/}}
 \newcommand{\urldocsgridpolator}{\url{https://vspec-collab.github.io/GridPolator/}}
@@
 \vspec{} can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements, and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs.
-We additionally present a set of codes developed alongside the core \vspec{} modules, including the stellar surface model generator \texttt{vspec-vsm}, the stellar spectral grid interpolation code GridPolator, and a Python interface for PSG \texttt{pypsg}.
+We additionally present a set of codes developed alongside the core \vspec{} modules, including the stellar surface model generator \texttt{vspec-vsm}, the stellar spectral grid interpolation code GridPolator, and a Python interface for PSG \texttt{libpypsg}.
 \end{abstract}
@@
 In this paper we present the \vspec{} (Variable Star PhasE Curve\footnote{\url{https://github.com/orgs/VSPEC-collab/repositories}}) Collection, a collection of open-source Python 3 packages to simulate observations of exoplanet systems with variable host stars.
-The \vspec{}\footnote{\url{https://github.com/VSPEC-collab/VSPEC}} package itself is merely an interface that pulls together a variable star surface model \texttt{vspec-vsm}\footnote{\url{https://github.com/VSPEC-collab/vspec-vsm}}, a stellar spectra grid interpolation code Gridpolator\footnote{\url{https://github.com/VSPEC-collab/GridPolator}}, and a Python interface for the Planetary Spectrum Generator \citep[PSG,][]{villanueva2018}, \texttt{pypsg}\footnote{\url{https://github.com/tedjohnson12/pypsg}}.
+The \vspec{}\footnote{\url{https://github.com/VSPEC-collab/VSPEC}} package itself is merely an interface that pulls together a variable star surface model \texttt{vspec-vsm}\footnote{\url{https://github.com/VSPEC-collab/vspec-vsm}}, a stellar spectra grid interpolation code Gridpolator\footnote{\url{https://github.com/VSPEC-collab/GridPolator}}, and a Python interface for the Planetary Spectrum Generator \citep[PSG,][]{villanueva2018}, \texttt{libpypsg}\footnote{\url{https://github.com/tedjohnson12/libpypsg}}.
 All of these codes together allow \vspec{} to account for 3D planetary atmospheres, time-resolved effects of both planet and star, multiple geometries including transit and secondary eclipse, and realistic noise modeling.
@@
 Section \ref{sec:vspec} describes \vspec{} and goes into some detail about setup, use, and the code's internals.
-In Section \ref{sec:pypsg} we describe \texttt{pypsg} as an interface for the Planetary Spectrum Generator.
+In Section \ref{sec:libpypsg} we describe \texttt{libpypsg} as an interface for the Planetary Spectrum Generator.
 Section \ref{sec:star} describes \vspec{}'s stellar model, \texttt{vspec-vsm}.
@@
 \vspec{} generates a phase curve though a series of API calls to the Planetary Spectrum Generator \citep[PSG][]{villanueva2018}, using
-\texttt{pypsg} (see Section \ref{sec:pypsg}) to interface with the API. Initially, configurations are sent that give PSG information about
+\texttt{libpypsg} (see Section \ref{sec:libpypsg}) to interface with the API. Initially, configurations are sent that give PSG information about
 the system that does not change with time -- that includes the 3D Global Circulation Model (GCM) and instrument parameters. \vspec{} then enters
@@
 
-\section{Interfacing with the Planetary Spectrum Generator via \texttt{pypsg}}
-\label{sec:pypsg}
+\section{Interfacing with the Planetary Spectrum Generator via \texttt{libpypsg}}
+\label{sec:libpypsg}
 
@@
 
-\vspec{} uses a Python interface for the PSG API system called \texttt{pypsg}, which is a general stand-alone library built for any Python-based API call to PSG.
-\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object.
-Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
+\vspec{} uses a Python interface for the PSG API system called \texttt{libpypsg}, which is a general stand-alone library built for any Python-based API call to PSG.
+\texttt{libpypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object.
+Fields of a \texttt{libpypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
 Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package.
@@
 
-A powerful feature of \texttt{pypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
+A powerful feature of \texttt{libpypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
 GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial.
-\texttt{pypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.
+\texttt{libpypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.
 
-Listing \ref{ls:pypsg} shows a minimal example that uses \texttt{pypsg} to make a call to PSG/GlobES.
-Here \texttt{pypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance.
+Listing \ref{ls:libpypsg} shows a minimal example that uses \texttt{libpypsg} to make a call to PSG/GlobES.
+Here \texttt{libpypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance.
 This configuration object is then passed to an \texttt{APICall} instance.
@@
 162-163]{villanueva2022}.
-More details on use of this library including its API and examples can be found on the \texttt{pypsg} documentation page\footnote{\urldocspypsg}.
+More details on use of this library including its API and examples can be found on the \texttt{libpypsg} documentation page\footnote{\urldocslibpypsg}.
 
 \begin{lstlisting}[
-    basicstyle=\footnotesize,language=Python,label={ls:pypsg},caption={\texttt{pypsg} Example}
+    basicstyle=\footnotesize,language=Python,label={ls:libpypsg},caption={\texttt{libpypsg} Example}
 ]
-    import pypsg
+    import libpypsg
     from astropy import units as u
@@
     # This next line only needs to be run once per user.
-    # These settings are saved to `~/.pypsg/settings.json`
-    pypsg.settings.save_setting(
+    #     These settings are 
+    #     saved to `~/.libpypsg/settings.json`
+    libpypsg.settings.save_setting(
         api_key='my_api_key',
@@
 
-    cfg = pypsg.cfg.PyConfig(
-        target=pypsg.cfg.Target(
+    cfg = libpypsg.cfg.PyConfig(
+        target=libpypsg.cfg.Target(
             object='Exoplanet',
@@
         ),
-        geometry=pypsg.cfg.Observatory(...),
-        generator=pypsg.cfg.Generator(...),
-        telescope=pypsg.cfg.SingleTelescope(...)
-        noise=pypsg.cfg.CCD(...)
-        gcm=pypsg.globes.waccm_to_pygcm(
+        geometry=libpypsg.cfg.Observatory(...),
+        generator=libpypsg.cfg.Generator(...),
+        telescope=libpypsg.cfg.SingleTelescope(...)
+        noise=libpypsg.cfg.CCD(...)
+        gcm=libpypsg.globes.waccm_to_pygcm(
             gcm_data,
@@
 
-    caller = pypsg.APICall(
+    caller = libpypsg.APICall(
         cfg=cfg,
@@
 \label{sec:vspec-gcm}
-\vspec{} allows users to run simulations with any climate model they desire (so long as it can be converted into the \texttt{pypsg.PyGCM} format) and even comes with utilities to work with PSG/GlobES binary files, ExoCAM \citep{wolf2022}, WACCM \citep{marsh2013}, and ExoPlaSim \citep{paradise2022} without any custom code from the user.
+\vspec{} allows users to run simulations with any climate model they desire (so long as it can be converted into the \texttt{libpypsg.PyGCM} format) and even comes with utilities to work with PSG/GlobES binary files, ExoCAM \citep{wolf2022}, WACCM \citep{marsh2013}, and ExoPlaSim \citep{paradise2022} without any custom code from the user.
 
@@
 GridPolator is currently in the process of adding support for more built-in stellar grids, these more complex and covering a larger range of spectral types that the current \vspec{} grid.
-\texttt{pypsg} has the most to offer the exoplanet community as a stand-alone package, but currently lacks support for much of the data that PSG is capable of producing.
+\texttt{libpypsg} has the most to offer the exoplanet community as a stand-alone package, but currently lacks support for much of the data that PSG is capable of producing.
 Its development so far has focused around the \vspec{} use case, so future work will center around making it more useful to the majority of PSG users.

### 2024-08-05 src/scripts/flare_freq.py

@@ -8,12 +8,12 @@ import matplotlib.pyplot as plt
 
 from vspec_vsm.flares import FlareGenerator
 from VSPEC import ObservationModel, PhaseAnalyzer
-import pypsg
+import libpypsg
 
 import paths
 
 import setup_psg;setup_psg.setup_psg()
-pypsg.docker.set_url_and_run()
+libpypsg.docker.set_url_and_run()
 
 SEED = 10
 outfile = paths.figures / 'flare_freq.pdf'

### 2024-08-05 src/scripts/granule_lc.py

@@ -7,7 +7,7 @@ https://vspec.readthedocs.io/en/latest/auto_examples/other/star/plot_granulation
 from pathlib import Path
 from astropy import units as u
 import matplotlib.pyplot as plt
-import pypsg
+import libpypsg
 
 from VSPEC import ObservationModel,PhaseAnalyzer
 from VSPEC import params
@@ -20,7 +20,7 @@ outfile = paths.figures / 'granulation_lightcurve.pdf'
 
 SEED = 32
 import setup_psg;setup_psg.setup_psg()
-pypsg.docker.set_url_and_run()
+libpypsg.docker.set_url_and_run()
 
 
 header = params.Header(

### 2024-08-05 src/scripts/kempton23.py

@@ -2,7 +2,7 @@ from pathlib import Path
 import numpy as np
 import matplotlib.pyplot as plt
 from astropy import units as u
-import pypsg
+import libpypsg
 
 from VSPEC import ObservationModel,PhaseAnalyzer
 from VSPEC import params
@@ -12,7 +12,7 @@ import paths
 
 SEED = 1214
 import setup_psg;setup_psg.setup_psg()
-pypsg.docker.set_url_and_run()
+libpypsg.docker.set_url_and_run()
 
 outfile = paths.figures / 'kempton23.pdf'
 

### 2024-08-05 src/scripts/ld_transit.py

@@ -7,7 +7,7 @@ from astropy import units as u
 import matplotlib.pyplot as plt
 import cartopy.crs as ccrs
 import VSPEC
-import pypsg
+import libpypsg
 
 import paths
 
@@ -15,7 +15,7 @@ LAMBERT_CONFIG = paths.scripts / 'ld_lc_lamb.yaml'
 SOLAR_CONFIG = paths.scripts / 'ld_lc_solar.yaml'
 OUTFILE = paths.figures / 'ld_transit.pdf'
 
-pypsg.docker.set_url_and_run()
+libpypsg.docker.set_url_and_run()
 
 lam_model = VSPEC.ObservationModel.from_yaml(LAMBERT_CONFIG)
 sol_model = VSPEC.ObservationModel.from_yaml(SOLAR_CONFIG)

### 2024-08-05 src/scripts/mirecle.py

@@ -6,7 +6,7 @@ from pathlib import Path
 import numpy as np
 import matplotlib.pyplot as plt
 from astropy import units as u
-import pypsg
+import libpypsg
 
 from VSPEC import ObservationModel,PhaseAnalyzer
 from VSPEC import params
@@ -17,7 +17,7 @@ outfile = paths.figures / 'mirecle.pdf'
 
 SEED = 10
 import setup_psg;setup_psg.setup_psg()
-pypsg.docker.set_url_and_run()
+libpypsg.docker.set_url_and_run()
 
 # Instrument
 inst = params.InstrumentParameters.mirecle()

### 2024-08-05 src/scripts/moran23.py

@@ -2,7 +2,7 @@ from pathlib import Path
 import numpy as np
 import matplotlib.pyplot as plt
 from astropy import units as u
-import pypsg
+import libpypsg
 
 from VSPEC import ObservationModel,PhaseAnalyzer
 from VSPEC import params
@@ -16,7 +16,7 @@ SEED = 10
 PRE_TRANSIT = 8
 IN_TRANSIT = 13
 import setup_psg;setup_psg.setup_psg()
-pypsg.docker.set_url_and_run()
+libpypsg.docker.set_url_and_run()
 
 # Instrument
 

### 2024-08-05 src/scripts/setup_psg.py

@@ -4,7 +4,7 @@ When running the GHA it is important that the PSG API key is set.
 
 import os
 
-import pypsg
+import libpypsg
 
 import paths
 
@@ -13,7 +13,7 @@ outfile = paths.output / 'psg_status.txt'
 def setup_psg():
     try:
         key = os.environ['PSG_API_KEY']
-        pypsg.settings.save_settings(api_key=key)
+        libpypsg.settings.save_settings(api_key=key)
         s = 'The PSG API key was set from an environment variable.'
 
     except KeyError:

### 2024-08-05 src/scripts/surface_map_and_lc.py

@@ -14,10 +14,10 @@ from vspec_vsm.flares import FlareGenerator
 from vspec_vsm.star import Star
 from vspec_vsm.config import MSH
 
-import pypsg
+import libpypsg
 
 import setup_psg;setup_psg.setup_psg()
-pypsg.docker.set_url_and_run()
+libpypsg.docker.set_url_and_run()
 
 from VSPEC import ObservationModel,PhaseAnalyzer
 

### 2024-08-07 src/tex/ms.tex
Make syw change to allow orchid

@@
 \usepackage{hyperref}
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\lstset{
+    backgroundcolor=\color{white},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\ttfamily\footnotesize,
+    breakatwhitespace=false,
+    breaklines=true,
+    captionpos=b,
+    keepspaces=true,
+    numbers=left,
+    numbersep=5pt,
+    showspaces=false,
+    showstringspaces=false,
+    showtabs=false,
+    tabsize=1
+}
 
@@
 
-% \author[0000-0002-1570-2203]{Ted M Johnson}
-\author{Ted M Johnson}
+% \author{Ted M Johnson}
+\author[0000-0002-1570-2203]{Ted M Johnson}
 \affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
@@
 human readability. For example, the \texttt{system} section, which describes the relationship between the observed planetary system and the observer, could
-be written:
-\begin{lstlisting}[
-    label={ls:yaml},
-    caption={\vspec{} YAML Configuration}
-]
+be written as shown in Listing \ref{ls:yaml}.
+\begin{lstlisting}[label={ls:yaml},caption=\vspec{} YAML Configuration]
 system:
@@
 The second input method available is to directly initialize the Python object that \vspec{} uses internally to store its model parameters.
-The top-level object, \texttt{InternalParameters}, is structured to mirror the YAML input file, with one argument per YAML section. The
-parameters in Listing \ref{ls:yaml} would be written:
-\begin{lstlisting}[language=Python, caption={\vspec{} Python configuration}]
+The top-level object, \texttt{InternalParameters}, is structured to mirror the YAML input file, with one argument per YAML section. The same
+configuration shown in Listing \ref{ls:yaml} would be written in Python as shown in Listing \ref{ls:python}.
+\begin{lstlisting}[language=Python, caption={\vspec{} Python configuration},label={ls:python}]
 from astropy import units as u
@@
 \label{sec:libpypsg}
-
 PSG\footnote{\url{https://psg.gsfc.nasa.gov/}} \citep{villanueva2018} is a powerful radiative transfer tool that is ubiquitous in the exoplanet and solar system atmosphere fields.
@@
 
-\vspec{} uses a Python interface for the PSG API system called \texttt{libpypsg}, which is a general stand-alone library built for any Python-based API call to PSG.
-\texttt{libpypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object.
-Fields of a \texttt{libpypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
-Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package.
-This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key could be committed to a public repository.
-
-A powerful feature of \texttt{libpypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
-GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial.
-\texttt{libpypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.
-
-Listing \ref{ls:libpypsg} shows a minimal example that uses \texttt{libpypsg} to make a call to PSG/GlobES.
-Here \texttt{libpypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance.
-This configuration object is then passed to an \texttt{APICall} instance.
-When the API call is made, the contents of the \texttt{PyConfig} are included in the body of the POST request in the \texttt{file} field.
-PSG reads this the same as it would if we had written those contents to a file \texttt{config.txt} and passed the \texttt{--data-urlencode file@config.txt} option to \texttt{cURL}, as are the instructions in the PSG handbook \citep[pp.
-162-163]{villanueva2022}.
-More details on use of this library including its API and examples can be found on the \texttt{libpypsg} documentation page\footnote{\urldocslibpypsg}.
-
 \begin{lstlisting}[
-    basicstyle=\footnotesize,language=Python,label={ls:libpypsg},caption={\texttt{libpypsg} Example}
+    language=Python,label={ls:libpypsg},
+    caption={\texttt{libpypsg} Example}
 ]
@@
 
-    # This next line only needs to be run once per user.
+# This next line only needs to be run
+#     once per user.
 #     These settings are 
@@
     gcm=libpypsg.globes.waccm_to_pygcm(
-            gcm_data,
-            itime=0,
+        gcm_data, itime=0,
         molecules=['H2O','CO2'],
         aerosols=['Water','WaterIce'],
-            mean_molecular_mass=48.0
-        )
-
+        mean_molecular_mass=48.0)
 )
@@
 caller = libpypsg.APICall(
-        cfg=cfg,
-        app='globes',
+    cfg=cfg, app='globes',
     output_type='rad'
@@
 response = caller()
+print(response.rad['Wave/freq','proxima-cen-b'])
+\end{lstlisting}
 
-    print(response.rad['Wave/freq','proxima-cen-b'])
+\vspec{} uses a Python interface for the PSG API system called \texttt{libpypsg}, which is a general stand-alone library built for any Python-based API call to PSG.
+\texttt{libpypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object.
+Fields of a \texttt{libpypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.
+Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package.
+This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key could be committed to a public repository.
 
-    -------------------
-
+\begin{lstlisting}[caption={\texttt{libpypsg} Output},label={ls:psgout}]
 Wave/freq  proxima-cen-b
@@
      1.0404   4.95983e-19
-       1.061208    5.0892e-19
-     1.08243216   3.79632e-19
     ...           ...
-     16.3163953   7.66301e-20
-     16.6427232   7.66245e-20
 16.97557767   7.54221e-20
@@
 \end{lstlisting}
+
+A powerful feature of \texttt{libpypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
+GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial.
+\texttt{libpypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.
+
+Listing \ref{ls:libpypsg} shows a minimal example that uses \texttt{libpypsg} to make a call to PSG/GlobES, with the output shown in Listing \ref{ls:psgout}.
+Here \texttt{libpypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance.
+This configuration object is then passed to an \texttt{APICall} instance.
+When the API call is made, the contents of the \texttt{PyConfig} are included in the body of the POST request in the \texttt{file} field.
+PSG reads this the same as it would if we had written those contents to a file \texttt{config.txt} and passed the \texttt{--data-urlencode file@config.txt} option to \texttt{cURL}, as are the instructions in the PSG handbook \citep[pp.
+162-163]{villanueva2022}.
+More details on use of this library including its API and examples can be found on the \texttt{libpypsg} documentation page\footnote{\urldocslibpypsg}.
 

### 2024-08-07 src/tex/ms.tex
Merge changes through Tuesday 8/6

@@
 
+\newcommand{\urldocsvspec}{\url{https://vspec-collab.github.io/VSPEC/0.6.0/index.html}}
 \newcommand{\urldocsvspecvsm}{\url{https://vspec-vsm.readthedocs.io/en/latest/}}
@@
 \author{Cameron Kelahan}
+\affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
+\affiliation{Center for Research and Exploration in Space Science and Technology, NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
 \affiliation{NASA Goddard Space Flight Center \\
@@
 
-\author{Tom Barclay}
+\author{Thomas Barclay}
 \affiliation{NASA Goddard Space Flight Center \\
@@
     We present the Variable Star PhasE Curve (VSPEC) Collection, a set of Python packages for simulating combined-light spectroscopic observations of 3-dimensional exoplanet atmospheres in the presence of stellar variability and inhomogeneity.
-\vspec{} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG/GlobES) application along with a custom-built multi-component time-variable stellar model based on an user-defined grid of stellar photosphere models to produce spectroscopic light curves of the planet-host system.
+\vspec{} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG/GlobES) application along with a custom-built multi-component time-variable stellar model based on a user-defined grid of stellar photosphere models to produce spectroscopic light curves of the planet-host system.
 \vspec{} can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements, and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs.
@@
 
-In the era of high-sensitivity transiting exoplanet characterization missions such as the Hubble Space Telescope (HST), the James Webb Space Telescope (JWST) and the future Atmospheric Remote-sensing Infrared Exoplanet Large-survey (ARIEL), spectral analysis of exoplanet atmospheres is increasingly sensitive to contamination due to stellar inhomogeneities (e.g.
+In the era of high-sensitivity transiting exoplanet characterization missions such as the Hubble Space Telescope (HST), the James Webb Space Telescope (JWST), and the future Atmospheric Remote-sensing Infrared Exoplanet Large-survey (ARIEL), spectral analysis of exoplanet atmospheres is increasingly sensitive to contamination due to stellar inhomogeneities (e.g.
 spots, granulation) and stochastic stellar variability (e.g.
 flares).
-As an example, recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between atmospheric absorption by water and water-rich spots on the stellar surface, and similar effects were seen in the transit spectrum of LHS 1140b \citep{cadieux2024} and many other recent transit measurements iwth JWST \citep[e.g.,][]{lim2023,may2023,fournier-tondreau2024}.
+As an example, recent analysis of the JWST/NIRSpec transit of GJ 486b by \citet{moran2023} exposed a degeneracy between atmospheric absorption by water and water-rich spots on the stellar surface, and similar effects were seen in the transit spectrum of LHS 1140b \citep{cadieux2024} and many other recent transit measurements with JWST \citep[e.g.,][]{lim2023,may2023,fournier-tondreau2024}.
 This ``transit light source effect'' \citep[TLS,][see also \citet{apai2018,barclay2021,garcia2022,barclay2023}]{rackham2018} occurs when the region of the stellar surface occulted by the transiting planet is not representative of the disk-integrated spectrum.
@@
 
-\subsection{Stellar Lightcurve}
+\subsection{Stellar light curve}
 
@@
 The star is initialized and allowed to evolve for a user-specified amount of time in order for spots and faculae to approach growth-decay equilibrium (this is also called the ``burn in'' phase).
-The star is then evolved at the output cadence, at each epoch generating a set of key-value pairs that describe the surface temperature and coverage fractions of the portion of the stellar disk visible to the observer.
+The star is then evolved at the output cadence, generating a set of key-value pairs at each epoch that describe the surface temperature and coverage fractions of the portion of the stellar disk visible to the observer.
 These \teff s and coverage fractions are fed into GridPolator to produce composite stellar spectra.
@@
 \subsection{Transit and Eclipse}
-Users must be careful when using \vspec{} to simulate transit and eclipse measurements because of the vast timescale differences between those events and the typical phase variations of a planet -- a sparsely sampled phase curve that happens to include one epoch of transit will not adequately sample the lightcurve accurately during a transit or eclipse because the interpolator knows no better than to assume a linear interpolation scheme.
+Users must be careful when using \vspec{} to simulate transit and eclipse measurements because of the vast timescale differences between those events and the typical phase variations of a planet -- a sparsely sampled phase curve that happens to include one epoch of transit will not adequately sample the light curve accurately during a transit or eclipse because the interpolator knows no better than to assume a linear interpolation scheme.
 It is recommended that users use a high cadence for observations of transits and eclipses so that each event will be sampled by more than one point.
@@
 Either through its web graphical interface, its public API, or the local version available through Docker, PSG allows users to simulate observations of planets, comets, and moons with a variety of geometries and realistic noise models.
+
+\vspec{} uses a Python interface for the PSG API system called \texttt{pypsg}, which is a general stand-alone library built for any Python-based API call to PSG.
+\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object.
+Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangeable without loss of information.
+Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package.
+This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key could be committed to a public repository.
+
+A powerful feature of \texttt{pypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
+GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial.
+\texttt{pypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.
+
+Listing \ref{ls:pypsg} shows a minimal example that uses \texttt{pypsg} to make a call to PSG/GlobES.
+Here \texttt{pypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance.
+This configuration object is then passed to an \texttt{APICall} instance.
+When the API call is made, the contents of the \texttt{PyConfig} are included in the body of the POST request in the \texttt{file} field.
+PSG reads this the same as it would if we had written those contents to a file \texttt{config.txt} and passed the \texttt{--data-urlencode file@config.txt} option to \texttt{cURL}, as are the instructions in the PSG handbook \citep[pp.
+162-163]{villanueva2022}.
+More details on use of this library including its API and examples can be found on the \texttt{pypsg} documentation page\footnote{\urldocslibpypsg}.
 
@@
 However, this can lead to unnecessary spatial oversampling at the poles, and may lead to poor resolution at the equator to keep the total number of points from being excessively high.
-To mitigate these problems \texttt{vspec-vsm} supports a ``grid'' generated using a Fibbonacci spiral to create a near-isotropic distribution of points.
+To mitigate these problems, \texttt{vspec-vsm} supports a ``grid'' generated using a Fibbonacci spiral to create a near-isotropic distribution of points.
 This is especially useful when simulating transits, giving optimal sampling at the stellar equator.
@@
     \caption{{\bf Top}: The surface map of a star with spots. The star has a photospheric temperature of 3300K, a penumbra temperature of 2700K, and an umbra temperature of 2500K.
-    {\bf Bottom}: An example lightcurve of a spotted star. Flux variations are due to varying spot coverage as a function of sub-observer point. Two rotational periods are shown.
+    {\bf Bottom}: An example light curve of a spotted star. Flux variations are due to varying spot coverage as a function of sub-observer point. Two rotational periods are shown.
     }
@@
 the hot wall is visible to the observer, and faculae appear as bright points; near the center, however, the cool floor is exposed and faculae appear dark.
-To consider this effect in the faculae lightcurve, we compute the fraction of each facula's projected area -- the area on the disk it would occupy as seen by the observer -- that is occupied by the hot wall versus the cool floor.
+To consider this effect in the faculae light curve, we compute the fraction of each facula's projected area -- the area on the disk it would occupy as seen by the observer -- that is occupied by the hot wall versus the cool floor.
 Mathematically, this is no different from finding the area of the intersection of two overlapping ellipses, and we use this conceptualization to calculate the contributions of each portion, as described below.
@@
 Flares are an important source of stellar variability on short timescales.
-\texttt{vspec-vsm}'s flare lightcurve model is based on the \texttt{xoflares} package \citep{barclay2020}, which itself is based on empirically derived lightcurve shapes from \citet{davenport2016}.
-However, \texttt{xoflares} is designed to fit lightcurves in a single spectral band, so we add additional parameters to extend it to a multiwavelength lightcurve.
+\texttt{vspec-vsm}'s flare light curve model is based on the \texttt{xoflares} package \citep{barclay2020}, which itself is based on empirically derived light curve shapes from \citet{davenport2016}.
+However, \texttt{xoflares} is designed to fit light curves in a single spectral band, so we add additional parameters to extend it to a multiwavelength light curve.
 We completely describe a flare using its temperature, total energy, full width at half maximum (FWHM), and the time of its peak.
 We model a flare as a hot, optically thin region above the photosphere that produces a blackbody spectrum.
-In this model, the temperature is constant, and the sharp rise and fall seen in the lightcurve is caused by the region's rapidly changing area.
+In this model, the temperature is constant, and the sharp rise and fall seen in the light curve is caused by the region's rapidly changing area.
 This simplification allows the total flare energy $E$ to be used as a normalization factor:
@@
 where $A$ is the time-dependent area of the flare region, $T$ is the constant flare temperature, and $\sigma$ is the Stefan-Boltzmann constant.
-This allows flare lightcurves to be produced with a fixed bolometric luminosity.
+This allows flare light curves to be produced with a fixed bolometric luminosity.
 Because the temperature is fixed, the relevant quantity for modeling the flux of a flare in a given integration is the integrated time-area $\int A\,dt$.
@@
 
-Before simulating an observation, \vspec{} asks \texttt{vspec-vsm} to pre-compute a population of flares based on a power-law frequency-energy relationship based on Kepler and TESS lightcurves \citep{gao2022}:
+Before simulating an observation, \vspec{} asks \texttt{vspec-vsm} to pre-compute a population of flares based on a power-law frequency-energy relationship based on Kepler and TESS light curves \citep{gao2022}:
 \begin{equation} \label{eq:flare_freq}
@@
 Error bars based on the square root of the number of observed flares as this is a Poisson process.
-{\bf Bottom}: Lightcurve of a star flaring with the same power-law slope as the top panel, but the intercept ($\beta$) has been increased by 0.3 for visual effect.
+{\bf Bottom}: light curve of a star flaring with the same power-law slope as the top panel, but the intercept ($\beta$) has been increased by 0.3 for visual effect.
 The mean flare temperature is $9000$ K and the mean FWHM is $3$ hours.
@@
     \caption{
-        Effect of granulation on the lightcurve of a star.
+        Effect of granulation on the light curve of a star.
 The parameters chosen in this simulation mean that, on average, 10\% of the stellar surface is covered by the cool, inter-granule region with a temperature 300K lower than the surrounding photosphere, and that this coverage varies by about 1\% on a timescale of 6 hours.
@@
     \caption{
-        {\bf Right}: Transit lightcurves at 1 ${\rm \mu m}$ for a Lambertian (i.e.
+        {\bf Right}: Transit light curves at 1 ${\rm \mu m}$ for a Lambertian (i.e.
 no limb darkening) star and one with significant limb darkening \citep[$u_1=0.5,\,u_2=0.15$,][]{espinoza2015}.
@@
 A better treatment would vary the temperature of the stellar model as a function of $\mu$; we simply apply a gray scaling factor to the flux.
-Figure \ref{fig:ld_transit} compares transit lightcurves for two different sets of limb darkening parameters.
+Figure \ref{fig:ld_transit} compares transit light curves for two different sets of limb darkening parameters.
 
@@
 Within this suite of codes, GridPolator has the added responsibility of binning spectra and producing wavelength axes that match those output by PSG.
-Wavelength sampling is done using constant resolving power $R=\frac{\lambda}{\Delta \lambda}$.
+Wavelength sampling is done using a constant resolving power $R=\frac{\lambda}{\Delta \lambda}$.
 The binning algorithm takes a pair of high-resolution flux and wavelength arrays and resamples the flux to a lower-resolution wavelength axis.
@@
         MIRI phase curves for GJ 1214b binned by wavelength. Compare to \citet{kempton2023} Extended Data Figure 1.
-        While the addition of stellar variability does have a visible effect on the lightcurve, it is much smaller than the thermal emission by the exoplanet.
+        While the addition of stellar variability does have a visible effect on the light curve, it is much smaller than the thermal emission by the exoplanet.
         }

### 2024-08-13 src/tex/ms.tex
Merge Cameron's edits

@@
 model \citep{spruit1976} where faculae are described as three-dimensional pores in the stellar surface with a hot, bright wall and a cool, dark floor, as shown in Figure \ref{fig:fac_struct}.
-Their three-dimensional structure causes faculaes' observational properties to change depending on their angle from disk-center.
+Their three-dimensional structure causes faculae's observational properties to change depending on their angle from disk-center.
 Close to the limb,
@@
 Each facula is born and dies at a radius of $e^{-2}$ of its maximum, effectively existing in the code for two lifetimes.
-\citet{hovis-afflerbach2022} suggest a typical facula lifetime to be on the order of 6 hours, with a distribution that resembles a Poisson function; however, we choose a lognormal distribution for both lifetime and maximum radius because it does not allow for these values to be 0.
+\citet{hovis-afflerbach2022} suggest a typical facula lifetime to be on the order of 6 hours, with a distribution that resembles a Poisson function; however we choose a log-normal distribution for both lifetime and maximum radius because it does not allow for these values to be 0.
 We choose to correlate lifetime and radius so that they are determined by a normalized random distribution - for each new facula, a value is randomly drawn and determines both the facula lifetime and maximum radius.
@@
 The stellar atmospheres used to create the spectra have solar metallicity and a $\log{g}$ of 5.0, with effective temperatures ranging from 2300K to 3900K in steps of 100K.
-Currently this is the only grid currently supported by \vspec{}, with support for additional grids planned in future releases.
+This is the only grid currently supported by \vspec{}, with support for additional grids planned in future releases.
 % The second grid is hosted by the Space Telescope Science Institute\footnote{\url{https://archive.stsci.edu/hlsps/reference-atlases/cdbs/grid/phoenix/}}.
@@
 The \vspec{} GCM centers around a 2D temperature map of the planet's surface. Based on \citet{cowan2011}, this map
-is determined exclusively by the incident flux from the star (parameterized by the host \teff~and radius
+is determined exclusively by the incident flux from the star (parameterized by the host \teff~and radius,
 and the planet's semimajor axis and Bond albedo) and the unitless thermal redistribution efficiency $\epsilon$.
@@
         {\bf Bottom}: Relative surface temperature as a function of latitude $\lambda$ for varying latitudinal mixing parameter $\alpha$.
-A value of 1 means the equatorial temperature as computed in the upper panel.
-Without mixing ($\alpha=0$) heat is concentrated at the Equator.
+A value of 1 means the equatorial temperature, as computed in the upper panel.
+Without mixing ($\alpha=0$), heat is concentrated at the equator.
 However, as $\alpha$ increases, the equator cools and heat is redistributed towards the poles.
@@
 
-The treatment by \citet{cowan2011} includes no mixing in the latitudinal direction, nor a vertical temperature profile; to allow for variations in pressure and inclination, we augment their treatment with several simple approximations for latitudinal and vertical temperature variations.
+The treatment by \citet{cowan2011} includes no mixing in the latitudinal direction, nor a vertical temperature profile.
+To allow for variations in pressure and inclination, we augment their treatment with several simple approximations for latitudinal and vertical temperature variations.
 
@@
 \end{equation}
-where $f(0;0)$ is the previously computed temperature at the equator (see Figure \ref{fig:cowan_curves}). Plugging this in to Equation (\ref{eq:e_out}), we see
+where $f(0;0)$ is the previously computed temperature at the equator (see Figure \ref{fig:cowan_curves}). Plugging this into Equation (\ref{eq:e_out}), we see
 \begin{equation}
@@
 
-Finally, we populate our atmosphere with molecules. The user specifies the species and abundance and atmosphere is filled accordingly with a constant molecular suite.
+Finally, we populate our atmosphere with molecules. The user specifies the species and abundance, and the atmosphere is filled accordingly with a constant molecular suite.
 
@@
 Future work on \vspec{} will concern improving performance so that it can be used to produce phase-curve grids for retrievals.
-A goal of \texttt{vspec-vsm} development should be to create a common API between it and more physics-based stellar models so that \vspec{} can interface with those models as well.
+A goal of \texttt{vspec-vsm} development should be to create a common API between it and more physics-based stellar models, so that \vspec{} can interface with those models as well.
 GridPolator is currently in the process of adding support for more built-in stellar grids, these more complex and covering a larger range of spectral types that the current \vspec{} grid.

### 2024-08-14 src/tex/ms.tex
Add conclusions and acknowledgements

@@
 \vspec{} utilizes a library of pre-computed stellar models for producing its multi-component composite stellar spectrum; detailed stellar spectral models are computationally expensive to generate, so it is more efficient to interpolate over a discretized grid.
-This interpolation is handled by the VSPEC-related Python package GridPolator, whose main class \texttt{GridSpectra} acts as a wrapper for the SciPy \citep{virtanen2020} or JAX \citep{jamesbradbury2018} \texttt{RegularGridInterpolator} with a specialized API.
+This interpolation is handled by the VSPEC-related Python package GridPolator, whose main class \texttt{GridSpectra} acts as a wrapper for the SciPy \citep{virtanen2020} or JAX \citep{bradbury2018} \texttt{RegularGridInterpolator} with a specialized API.
 This setup accomodates spectral grids of arbitrary dimension, and treats the wavelength axis independently of the other parameters so that each wavelength channel is independent.
@@
 \subsection{Phase Curve}
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/kempton23.pdf}
+    \caption{
+        MIRI phase curves for GJ 1214b binned by wavelength. Compare to \citet{kempton2023} Extended Data Figure 1.
+        While the addition of stellar variability does have a visible effect on the light curve, it is much smaller than the thermal emission by the exoplanet.
+        }
+    \script{kempton23.py}
+    \label{fig:gj1214b}
+\end{figure}
+
 Analysis of JWST MIRI-LRS phase curves of GJ 1214b by \citet{kempton2023} did not consider stellar
@@
 We assume the star behaves linearly between eclipse measurements and attribute any deviation to the planet.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{figures/kempton23.pdf}
-    \caption{
-        MIRI phase curves for GJ 1214b binned by wavelength. Compare to \citet{kempton2023} Extended Data Figure 1.
-        While the addition of stellar variability does have a visible effect on the light curve, it is much smaller than the thermal emission by the exoplanet.
-        }
-    \script{kempton23.py}
-    \label{fig:gj1214b}
-\end{figure}
 
@@
 
-While development is ongoing, these codes are all stable.
+While all of these codes are stable, development is ongoing.
 Future work on \vspec{} will concern improving performance so that it can be used to produce phase-curve grids for retrievals.
+In general, a phase curve retrieval will not require the stellar portion of VSPEC's output, and if there is no need for both thermal and reflected spectra, then only one PSG calculation is required per epoch.
+This could reduce \vspec{}'s computational cost by approximately a factor of 4 purely by skipping unnecessary steps.
+
 A goal of \texttt{vspec-vsm} development should be to create a common API between it and more physics-based stellar models, so that \vspec{} can interface with those models as well.
 GridPolator is currently in the process of adding support for more built-in stellar grids, these more complex and covering a larger range of spectral types that the current \vspec{} grid.
+However, as the size of these grids grows, it becomes less efficient to build the grid prior to each run.
+Future versions will support caching of grids locally using standard Python object serialization libraries such as \texttt{pickle}\footnote{\url{https://docs.python.org/3/library/pickle.html}}.
+This method would allow interpolator objects to be recalled nearly instantly (limited by hard drive read speeds) rather than the expensive process of re-computation.
+
 \texttt{libpypsg} has the most to offer the exoplanet community as a stand-alone package, but currently lacks support for much of the data that PSG is capable of producing.
@@
 \section{Acknowledgements}
-The authors would like to thank Peter Hauschildt for generating a custom grid of PHOENIX models.
+The authors would like to thank Peter Hauschildt for generating a custom grid of PHOENIX stellar models.
+
+This manuscript was typeset using the open-science software \showyourwork, which makes the entire manuscript, including each figure, completely reproducible by anyone with access to the git repository at \url{https://github.com/VSPEC-Collab/vspec-paper}.
+
+This work made use of Astropy:\footnote{\url{http://www.astropy.org}} a community-developed core Python package and an ecosystem of tools and resources for astronomy \citep{astropycollaboration2013,astropycollaboration2018,astropycollaboration2022}.
+
+The figures in this article were created using the Matplotlib \citep{hunter2007}, NumPy \citep{harris2020}, and Cartopy \citep{metoffice2010} libraries.
 

### 2024-08-14 src/scripts/kempton23.py

@@ -212,7 +212,7 @@ def get_lc(data:PhaseAnalyzer,w1:u.Quantity,w2:u.Quantity):
 
 bin_edges = np.arange(5.0,12.0,0.5)
 n_ax = len(bin_edges)
-fig,axes = plt.subplots(n_ax,1,figsize=(4,10),sharex=True)
+fig,axes = plt.subplots(n_ax,1,figsize=(4,8.5),sharex=True)
 for edge,ax in zip(bin_edges,axes):
     w1,w2 = edge*u.um,(edge+0.5)*u.um
     quiet_lc = get_lc(data_quiet,w1,w2)

### 2024-08-14 src/tex/ms.tex
Clean up psg section

@@
 
-\vspec{} uses a Python interface for the PSG API system called \texttt{pypsg}, which is a general stand-alone library built for any Python-based API call to PSG.
-\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object.
-Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangeable without loss of information.
-Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package.
-This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key could be committed to a public repository.
-
-A powerful feature of \texttt{pypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
-GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial.
-\texttt{pypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.
-
-Listing \ref{ls:pypsg} shows a minimal example that uses \texttt{pypsg} to make a call to PSG/GlobES.
-Here \texttt{pypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance.
-This configuration object is then passed to an \texttt{APICall} instance.
-When the API call is made, the contents of the \texttt{PyConfig} are included in the body of the POST request in the \texttt{file} field.
-PSG reads this the same as it would if we had written those contents to a file \texttt{config.txt} and passed the \texttt{--data-urlencode file@config.txt} option to \texttt{cURL}, as are the instructions in the PSG handbook \citep[pp.
-162-163]{villanueva2022}.
-More details on use of this library including its API and examples can be found on the \texttt{pypsg} documentation page\footnote{\urldocslibpypsg}.
-
 \begin{lstlisting}[
@@
 Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package.
-This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key could be committed to a public repository.
+This dedicated caller is useful because it reads a user's settings (e.g.
+an API key or the URL of the PSG API endpoint) from a file on their hard drive.
+This makes it much simpler to publish code that utilizes PSG because 1) it eliminates the danger of accidentally publishing an API key, and 2) it means that users can run the same code regardless of whether they have a local installation of PSG.
 
-\begin{lstlisting}[caption={\texttt{libpypsg} Output},label={ls:psgout}]
+\begin{lstlisting}[caption={\texttt{libpypsg} Example Output},label={ls:psgout}]
 Wave/freq  proxima-cen-b
@@
 
-A powerful feature of \texttt{libpypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
+A powerful feature of \texttt{libpypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application \TJ{(Fauchez et al., 2024, in review; see also Kofman et al., 2024, accepted)}.
 GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial.
-\texttt{libpypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.
+\texttt{libpypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022}, the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013}, and ExoPlaSim \citep{paradise2022} to the correct format.
 

### 2024-08-14 src/tex/ms.tex
Maybe the last editing session?

@@
 \begin{abstract}
-    We present the Variable Star PhasE Curve (VSPEC) Collection, a set of Python packages for simulating combined-light spectroscopic observations of 3-dimensional exoplanet atmospheres in the presence of stellar variability and inhomogeneity.
-\vspec{} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG/GlobES) application along with a custom-built multi-component time-variable stellar model based on a user-defined grid of stellar photosphere models to produce spectroscopic light curves of the planet-host system.
-\vspec{} can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements, and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs.
-We additionally present a set of codes developed alongside the core \vspec{} modules, including the stellar surface model generator \texttt{vspec-vsm}, the stellar spectral grid interpolation code GridPolator, and a Python interface for PSG \texttt{libpypsg}.
+    We present the Variable Star PhasE Curve (\texttt{VSPEC}) Collection, a set of Python packages for simulating combined-light spectroscopic observations of 3-dimensional exoplanet atmospheres in the presence of stellar variability and inhomogeneity.
+\texttt{VSPEC} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG/GlobES) application along with a custom-built multi-component time-variable stellar model based on a user-defined grid of stellar photosphere models to produce spectroscopic light curves of the planet-host system.
+\texttt{VSPEC} can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements, and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs.
+We additionally present a set of codes developed alongside the core \texttt{VSPEC} modules, including the stellar surface model generator \texttt{vspec-vsm}, the stellar spectral grid interpolation code GridPolator, and a Python interface for PSG, \texttt{libpypsg}.
 \end{abstract}
@@
     language=Python,label={ls:libpypsg},
-    caption={\texttt{libpypsg} Example}
+    caption={\texttt{libpypsg} Example Input}
 ]
@@
 
-In this section we will simulate the transit of a bare rock across a spotted star to demonstrate how an inhomogeneous
+In this section we will simulate the transit of a atmosphere-less occulting disk across a spotted star to demonstrate how an inhomogeneous
 stellar surface can lead to a false atmospheric signal -- a phenomenon known as the ``transit light source effect'' \citep{rackham2018}.
-During transit, the planetary atmosphere is illuminated by the portion of the star directly behind the planet. The depth of the transit
-should be measured with respect for the spectrum of this region of the stellar surface. However, the disk-integrated stellar spectrum is often used instead.
-In the case that the stellar surface is not homogeneous, the transit signal is contaminated by the star.
-We will reproduce results by \citet{moran2023}, who observed this effect with JWST/NIRSpec during a transit of super-Earth GJ 486b.
 
-Whenever possible, we draw parameter values from \citet{moran2023} and the NExScI Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/GJ1214b}}.
-We use the JWST NIRSpec/G395H instrument setup but set $R=200$ to account for the resolution of the reduction.
-Additionally we observe for 3.53 hours with an 8 minute cadence. Mid-transit is reached exactly halfway through the observation.
+During primary transit of a planet with an atmosphere, the atmosphere is illuminated by the portion of the star directly behind the planet.
+Ideally, the depth of the transit should be measured with respect to the spectrum of the region of the stellar surface covered by the planet.
+However, the disk-integrated stellar spectrum (as measured before an after the transit) is used as the baseline.
+In the case that the stellar surface is not homogeneous, the transit signal is modulated by the uncorrected stellar inhomogeneities.
+In this example, we will reproduce the TLS results by \citet{moran2023}, who contended with this degeneracy with JWST/NIRSpec observations during a transit of super-Earth GJ 486b.
 
-We will do two \vspec{} simulations:
+Whenever possible, we draw parameter values from \citet{moran2023} and the NExScI rchive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/GJ486b}}.
+We use the JWST NIRSpec/G395H instrument setup from PSG with an observing time of 3.53 hours with an 8 minute cadence to match \citet{moran2023}, but set $R=200$ to account for the resolution of their reduction.
+
+We performed two \vspec{} simulations based on the two scenarios described in \citet{moran2023}:
 \begin{enumerate}
@@
 \end{enumerate}
+
 In Case \ref{case:1barh2o} we expect to see an absorption spectrum characteristic of \ce{H2O}.
@@
     \caption{
-        The results from our spotted transit experiment. We find that the addition of
-        spots to a star's surface can give an otherwise flat transit spectrum features that appear to be molecular bands.
+        The results from our spotted transit experiment.
+We find that the addition of spots to a star's surface can produce spectral features in what would otherwise be a flat transit spectrum due to a lack of atmosphere or other mitigating issues such as planetary aerosols.
+These features mimic those that would be produced due to molecular absorption in a planetary atmosphere.
 This is in agreement with the results of \citet{moran2023}.
@@
 
-We find that, as expected, the addition of spots changes the transit depth in a way that depends on wavelength -- mimicking
-light lost to absorbers in a planetary atmosphere. Figure \ref{fig:moran_transit} shows us that, in this case, we observe false absorption;
-this is due to a temporary increase in the spotted fraction of the surface relative to the quiet photosphere. However, if a spot was blocked by
-the planet we might see similar features inverted to mimic emission.
+We find that, as expected, the addition of spots changes the transit depth in a way that depends on wavelength -- mimicking light that would be lost to absorbers in a planetary atmosphere.
+Figure \ref{fig:moran_transit} shows us that, in this case, we observe false absorption signals; this is due to a temporary increase in the spotted fraction of the surface relative to the quiet photosphere.
+However, if a spot was blocked by the planet we might see similar features inverted to mimic emission.
 
@@
     \caption{
-        MIRI phase curves for GJ 1214b binned by wavelength. Compare to \citet{kempton2023} Extended Data Figure 1.
-        While the addition of stellar variability does have a visible effect on the light curve, it is much smaller than the thermal emission by the exoplanet.
+        A simulation of a MIRI phase curve for GJ 1214b, binned by wavelength; compare to real data published by \citet{kempton2023}(Extended Data Figure 1).
+While the addition of stellar variability does have a visible effect on the light curve, it is much smaller than the thermal emission produced by the exoplanet.
         }
@@
 
-Analysis of JWST MIRI-LRS phase curves of GJ 1214b by \citet{kempton2023} did not consider stellar
-contamination because of the slow rotation rate of the host star \citep[approximately 1/80$^{\text{th}}$ the orbital frequency,][]{cloutier2021}.
-In this example we simulate this system using \vspec{} and find that this is a reasonable assumption.
+Analysis of JWST MIRI-LRS phase curves of GJ 1214b by \citet{kempton2023} did not consider stellar contamination because of the slow rotation rate of the host star \citep[approximately 1/80$^{\text{th}}$ the orbital frequency,][]{cloutier2021}.
+In this example, we simulate these observations using \vspec{} and determine whether this is a reasonable assumption.
 
-We initialize \vspec{} to run a 41.0 hour observation with a cadence of 15 minutes. We use stellar and planetary parameters
-from the NExScI Exoplanet Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/GJ1214b}}.
-The GCM used has a thermal inertia of $\epsilon = 6$ and a 1 bar \ce{CO2} atmosphere.
+We initialized \vspec{} to simulate an observation of the system using the same instrument setup and observation parameters as \citet{kempton2023}.
+We used stellar and planetary parameters from the NExScI Exoplanet Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/GJ1214b}}.
+The VSPEC GCM (see Section \ref{sec:vspec-gcm}) used has a thermal inertia of $\epsilon = 6$ and a 1 bar \ce{CO2} atmosphere, appropriate for a high-metallicity \ce{CO2}-rich atmosphere scenario.
 
 We initialized two stellar models: one with nothing more than a 3250 K surface, and another with 20\% coverage by 2700 K spots.
-
-After running the \vspec{} simulations, we bin into 0.5 $\mu m$ wavelength channels and normalize each channel to the two eclipses.
-We assume the star behaves linearly between eclipse measurements and attribute any deviation to the planet.
+After running the \vspec{} simulations, we binned the simulated data into 0.5 $\mu m$ wavelength channels and normalized each channel to the two eclipses, similar to standard data analysis techniques; this assumes that the star behaves linearly between eclipse measurements and therefore attributes any deviation to the planet.
 
 We show in Figure \ref{fig:gj1214b} that the stellar properties assumed in this example do not generate enough contamination to
-significantly change the observed phase curve of GJ 1214b. At it's worst (i.e. near transit) the observed spectrum varies by 50 ppm due to stellar
-contamination -- compared to the 700 ppm relative flux of the planet at $8.5 {\mu m}$. Similar studies could be done to assess the potential for
-stellar contamination in other exoplanetary systems in order to judge fitness as a target.
+significantly change the observed phase curve of GJ 1214b. At worst (i.e. near transit), the observed spectrum varies by 50 ppm due to stellar
+contamination -- compared to the 700 ppm relative flux of the planet at $8.5 {\mu m}$.
 
-% \subsection{MIRECLE Phase Curve}
-% In this example we will look at the best-case target for the Mid-IR Exoplanet CLimate Explorer \citep[MIRECLE,][]{mandell2022}
-% mission concept: Proxima Centauri b with no stellar variability. We will instead examine the expected noise produced by this observation
-% and assess the detectability of the planet in a single orbit. \vspec{} uses the noise models built into PSG, allowing native support for
-% noise due to the instrument, source, and astrophysical background.
-
-% We use \vspec{}'s built-in MIRECLE instrument template, which includes a 2 m aperture and observes from 1-18 $\mu m$ with $R=50$.
-% We draw stellar and bulk planetary parameters from the NExScI Exoplanet Archive\footnote{\url{https://exoplanetarchive.ipac.caltech.edu/overview/proxima\%20cen}}
-% and assume $i=85^\circ$ and Earth density. The GCM we use has a Bond albedo of 0.3, $\epsilon=1.5$, $P_{\rm surf} = 1 \text{bar}$ and a 100 ppm \ce{CO2} atmosphere with a \ce{N2} background.
-% We ran the simulation for one orbit with an 24 hour cadence. Because this is a best-case scenario, we assumed perfect knowledge of the stellar spectrum.
-% We are left with the planetary spectrum and photon noise from the star.
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=0.5\textwidth]{figures/mirecle.pdf}
-%     \caption{
-%         Simulated spectra of Proxima Centauri b observed by MIRECLE. Noise and error bars based on 24 hour integrations with a 2 m aperture MIRECLE
-%         space telescope.
-%         }
-%     \script{mirecle.py}
-%     \label{fig:mirecle}
-% \end{figure}
+Similar studies could be done to assess the potential for
+stellar contamination in other exoplanetary systems in order to judge fitness as a target; in particular, faster stellar rotation should produce much larger contamination signals that may approach or dwarf the planetary signal.
 
@@
 The main package, \vspec{}, produces combined-light phase curves by interfacing with an ensemble of support packages whose responsibilities range from grid interpolation to interfacing with PSG.
-These codes are all open-source and available to the public on GitHub and the Python package index, PyPI.
+These codes are all open-source and available to the public on GitHub and the Python Package Index (PyPI).
 
 While all of these codes are stable, development is ongoing.
-Future work on \vspec{} will concern improving performance so that it can be used to produce phase-curve grids for retrievals.
+Future work on \vspec{} will focus on improving performance so that it can be used to produce phase-curve grids for retrievals.
 In general, a phase curve retrieval will not require the stellar portion of VSPEC's output, and if there is no need for both thermal and reflected spectra, then only one PSG calculation is required per epoch.
@@
 
-A goal of \texttt{vspec-vsm} development should be to create a common API between it and more physics-based stellar models, so that \vspec{} can interface with those models as well.
-GridPolator is currently in the process of adding support for more built-in stellar grids, these more complex and covering a larger range of spectral types that the current \vspec{} grid.
+A future goal of \texttt{vspec-vsm} development should be to create a common API for externally-developed physics-based variable star surface models, so that \vspec{} can interface with those models as well.
+GridPolator is currently in the process of adding support for more built-in stellar grids, which are more complex and cover a larger range of spectral types than the current \vspec{} grid.
 However, as the size of these grids grows, it becomes less efficient to build the grid prior to each run.
 Future versions will support caching of grids locally using standard Python object serialization libraries such as \texttt{pickle}\footnote{\url{https://docs.python.org/3/library/pickle.html}}.
-This method would allow interpolator objects to be recalled nearly instantly (limited by hard drive read speeds) rather than the expensive process of re-computation.
+This method would allow interpolator-oriented Python objects to be recalled nearly instantly (limited by hard drive read-write speeds) rather than the expensive process of re-initialization.
 
@@
 The authors would like to thank Peter Hauschildt for generating a custom grid of PHOENIX stellar models.
+
+T.J. acknowledges support from the Exoplanet Modeling and Analysis Center \citep[EMAC, ][]{renaud2023}.
+
+The authors would like to thank the Sellers Exoplanet Environments Collaboration (SEEC) and Exoplanet Spectroscopy Technologies (ExoSpec) ISFM teams at NASA's Goddard Space Flight Center for their consistent financial and collaborative support.
 

### 2024-08-14 src/tex/ms.tex
Add missing citations

@@
 \usepackage{hyperref}
+\usepackage{natbib}
 
@@
 
-A powerful feature of \texttt{libpypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application \TJ{(Fauchez et al., 2024, in review; see also Kofman et al., 2024, accepted)}.
+A powerful feature of \texttt{libpypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application (\citealt{fauchez2024}, in prep; see also \citealt{kofman2024}, accepted)
 GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial.
@@
 
-T.J. acknowledges support from the Exoplanet Modeling and Analysis Center \citep[EMAC, ][]{renaud2023}.
+T.J. acknowledges support from the Exoplanet Modeling and Analysis Center \citep[EMAC, ][]{renaud2022}.
 

### 2024-08-15 src/tex/ms.tex
change to elsarticle format

@@
 % Define document class
-\documentclass[twocolumn,linenumbers]{aastex631}
+\documentclass[linenumbers,preprint,authoryear]{elsarticle}
 \usepackage{showyourwork}
@@
 
-\shorttitle{\vspec{}}
-\shortauthors{Johnson et al.}
+% \shorttitle{\vspec{}}
+% \shortauthors{Johnson et al.}
 
@@
 % Title
-% \title{The \vspec{} code: Spectroscopic phase curves of 3D exoplanet models in the presence of stellar variability}
 \title{The \vspec{} Collection: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.}
 
-\correspondingauthor{Ted M Johnson}
-\email{ted.johnson@unlv.edu}
+\author[1,2,3,4,5]{Ted M Johnson\corref{cor1}}
+\ead{ted.johnson@unlv.edu}
 
-% \author{Ted M Johnson}
-\author[0000-0002-1570-2203]{Ted M Johnson}
-\affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
-\affiliation{Center for Research and Exploration in Space Science and Technology, NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
-\affiliation{NASA Goddard Space Flight Center \\
-8800 Greenbelt Rd \\
-Greenbelt, MD 20771, USA}
-\affiliation{Nevada Center for Astrophysics, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
-\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
+\author[1,2,3]{Cameron Kelahan}
 
-\author{Cameron Kelahan}
-\affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
-\affiliation{Center for Research and Exploration in Space Science and Technology, NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
-\affiliation{NASA Goddard Space Flight Center \\
-8800 Greenbelt Rd \\
-Greenbelt, MD 20771, USA}
+\author[3]{Avi M. Mandell}
 
-\author{Avi M. Mandell}
-\affiliation{NASA Goddard Space Flight Center \\
-8800 Greenbelt Rd \\
-Greenbelt, MD 20771, USA}
+\author[3]{Ashraf Dhahbi}
 
-\author{Ashraf Dhahbi}
-\affiliation{NASA Goddard Space Flight Center \\
-8800 Greenbelt Rd \\
-Greenbelt, MD 20771, USA}
+\author[3]{Tobi Hammond}
 
-\author{Tobi Hammond}
-\affiliation{NASA Goddard Space Flight Center \\
-8800 Greenbelt Rd \\
-Greenbelt, MD 20771, USA}
+\author[3]{Thomas Barclay}
 
-\author{Thomas Barclay}
-\affiliation{NASA Goddard Space Flight Center \\
-8800 Greenbelt Rd \\
-Greenbelt, MD 20771, USA}
+\author[3]{Veselin B. Kostov}
+\author[3]{Geronimo L. Villanueva}
 
-\author{Veselin B. Kostov}
-\affiliation{NASA Goddard Space Flight Center \\
-8800 Greenbelt Rd \\
-Greenbelt, MD 20771, USA}
+\cortext[cor1]{Corresponding Author}
 
-\author{Geronimo L. Villanueva}
-\affiliation{NASA Goddard Space Flight Center \\
-8800 Greenbelt Rd \\
-Greenbelt, MD 20771, USA}
+\affiliation[1]{
+    organization={Southeastern Universities Research Association},
+    addressline={1201 New York Avenue NW, Suite 430},
+    city={Washington, DC},
+    postcode={20055},
+    country={USA}
+}
+\affiliation[2]{
+    organization={Center for Research and Exploration in Space Science and Technology, NASA Goddard Space Flight Center},
+    addressline={8800 Greenbelt Road},
+    city={Greenbelt, MD},
+    postcode={20771},
+    country={USA}
+}
+\affiliation[3]{
+    organization={NASA Goddard Space Flight Center},
+    addressline={8800 Greenbelt Road},
+    city={Greenbelt, MD},
+    postcode={20771},
+    country={USA}
+}
+\affiliation[4]{
+    organization={Nevada Center for Astrophysics, University of Nevada, Las Vegas},
+    addressline={4505 South Maryland Parkway},
+    city={Las Vegas, NV},
+    postcode={89154},
+    country={USA}
+}
+\affiliation[5]{
+    organization={Department of Physics and Astronomy, University of Nevada, Las Vegas},
+    addressline={4505 South Maryland Parkway},
+    city={Las Vegas, NV},
+    postcode={89154},
+    country={USA}
+} 
+\affiliation[6]{
+    organization={University of Maryland, Baltimore County},
+    addressline={1000 Hilltop Cir},
+    city={Baltimore, MD},
+    postcode={21250},
+    country={USA}
+}
 
-% Abstract with filler text
 \begin{abstract}
@@
 
-\keywords{Exoplanet, M dwarf}
+\begin{keyword}
+    Exoplanet Atmospheres, Stellar Variability, Spectroscopic Phase Curve, Open Source, Planetary Spectrum Generator
+\end{keyword}
 
-% Main body with filler text
+% Highlights:
+% Instruments that characterize exoplanet atmospheres are sensitive to the effects of stellar inhomogeneities.
+% We present VSPEC, a code to simulate observations of 3D planets in the presence of stellar variability and inhomogeneity.
+% VSPEC, and the codes that support it, are open source, well documented, and available to the public.
+
+\maketitle
+
 \section{Introduction}
@@
     \centering
-    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_diagram.pdf}}
-    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_depth.pdf}}
+    \includegraphics[width=0.5\textwidth]{figures/facula_diagram.pdf}
+    \includegraphics[width=0.5\textwidth]{figures/facula_depth.pdf}
     \script{facula_depth.py}
@@
 
+\bibliographystyle{elsarticle-harv}
 \bibliography{syw,VSPEC}

### 2024-08-15 src/tex/ms.tex
remove all inline \frac

@@
 Within this suite of codes, GridPolator has the added responsibility of binning spectra and producing wavelength axes that match those output by PSG.
-Wavelength sampling is done using a constant resolving power $R=\frac{\lambda}{\Delta \lambda}$.
+Wavelength sampling is done using a constant resolving power $R=\lambda/{\Delta \lambda}$.
 The binning algorithm takes a pair of high-resolution flux and wavelength arrays and resamples the flux to a lower-resolution wavelength axis.
-PSG specifies the central wavelength of each channel, and so this algorithm defines bin edges to be $\lambda_{\rm cen} \pm \frac{\lambda_{\rm cen}}{R}$.
+PSG specifies the central wavelength of each channel, and so this algorithm defines bin edges to be $\lambda_{\rm cen} \pm {\lambda_{\rm cen}}/{R}$.
 

### 2024-08-15 src/tex/ms.tex
change e in formulas to exp Note that $e^{-1}$ and other simple powers of e are excluded from this
Add author contributions

@@
     \begin{array}{lr}
-        A_0 e^{(t-t_0)/\tau}, & \text{if } t \leq t_0 \\
+        A_0 \exp{((t-t_0)/\tau)}, & \text{if } t \leq t_0 \\
         A_0 - W(t-t_0), & \text{if } t > t_0
@@
 
+\section{Author Contributions} % see https://www.elsevier.com/researcher/author/policies-and-guidelines/credit-author-statement
+{\bf Ted M Johnson:} Conceptualization, Methodology, Software, Writing - Original Draft
+{\bf Cameron Kelahan:} Conceptualization, Methodology, Software, Writing - Review \& Editing
+{\bf Avi M. Mandell:} Conceptualization, Methodology, Writing - Original Draft, Writing - Review \& Editing, Supervision, Project administration, Funding acquisition
+{\bf Ashraf Dhahbi:} Software
+{\bf Tobi Hammond:} Software
+{\bf Thomas Barclay:} Methodology, Software
+{\bf Veselin B. Kostov:} Methodology
+{\bf Geronimo L. Villanueva:} Resources, Writing - Review \& Editing, Supervision
+
 \bibliographystyle{elsarticle-harv}

### 2024-08-20 src/tex/ms.tex
Clean up article
Move figs around, add inkscape acknowledgement
add orcid IDS
Add orcid IDS
Merge response to Veselin's comments

@@
 % Define document class
-\documentclass[linenumbers,preprint,authoryear]{elsarticle}
+\documentclass[linenumbers,5p,twocolumn,authoryear]{elsarticle}
 \usepackage{showyourwork}
@@
 \usepackage{natbib}
+\usepackage{xcolor}
+\usepackage{orcidlink}
 
@@
 
-\author[1,2,3,4,5]{Ted M Johnson\corref{cor1}}
+\author[1,2,3,4,5]{Ted M Johnson\,\orcidlink{0000-0002-1570-2203}\corref{cor1}}
 \ead{ted.johnson@unlv.edu}
 
-\author[1,2,3]{Cameron Kelahan}
+\author[1,2,3]{Cameron Kelahan\,\orcidlink{0009-0002-0075-8041}}
 
-\author[3]{Avi M. Mandell}
+\author[3]{Avi M. Mandell\,\orcidlink{0000-0002-8119-3355}}
 
-\author[3]{Ashraf Dhahbi}
+\author[3]{Ashraf Dhahbi\,\orcidlink{0009-0009-0200-6821}}
 
-\author[3]{Tobi Hammond}
+\author[3]{Tobi Hammond\,\orcidlink{0009-0002-5756-9778}}
 
-\author[3]{Thomas Barclay}
+\author[3]{Thomas Barclay\,\orcidlink{0000-0001-7139-2724}}
 
-\author[3]{Veselin B. Kostov}
-\author[3]{Geronimo L. Villanueva}
+\author[3]{Veselin B. Kostov\,\orcidlink{0000-0001-9786-1031}}
+\author[3]{Geronimo L. Villanueva\,\orcidlink{0000-0002-2662-5776}}
 
@@
 
-\begin{figure*}
+\begin{figure*}[h]
     \centering
@@
 In the case of a transiting geometry, \texttt{vspec-vsm} is responsible for computing the properties of the portion of the stellar disk that is occulted.
-It does this by first projecting both the transiting planet and the visible part of the surface orthographically as they would be viewed by a distant observer.
-It then selects all the points that are near enough to the transit to be occulted.
-The code then iterates through each of those points and computes the fraction of it covered by the occulting circle using a 2D numerical integral.
-The weight of each point is determined by the geometric projected area as well as a quadratic limb darkening law \citep[see][]{espinoza2015}.
+It does this by first projecting every visible point (i.e.
+those for which $\mu > 0$, $\mu$ is the cosine of the angle from disk-center) orthographically onto an $x$-$y$ plane where the stellar disk is represented by the unit circle.
+The planet, in this projection, is represented by a circle centered at $(a/R_*\,\sin{\phi},~ a/R_*\,\cos{\phi}\,\cos{i})$ with a radius of $R_p/R_*$, for semi-major axis $a$, stellar radius $R_*$, planetary radius $R_p$, phase past midtransit $\phi$, and inclination $i$.
+It then iterates through each point on the surface, computing the distance in projected coordinates to the center of the occulting disk.
+Also, for each point, an effective radius is calculated in the projected coordinates
+\begin{equation}
+    r_{\rm eff} = 2\mu\sqrt{\Omega/(4\pi\,\text{[sr]})}
+\end{equation}
+where $\Omega$ is the solid angle of each point.
+Any point for which the distance is less than $R_p/R_* + 2r_{\rm eff}$ from the occulting disk's center is flagged as possibly interacting with the planet.
+The majority of these points are completely covered by the planet, and those that are less than $R_p/R_* - 2r_{\rm eff}$ from the center of the occultation are immediately marked as such.
+The remaining points -- those near the edge of the occultation -- are subdivided into a $100\times 100$ point meshgrid.
+A 2D numerical integral is performed over this grid to compute the fraction of each point that is inside the occulting disk.
+
+ When computing the disk-integrated flux (including flux removed via a transit) the weight of each point is determined by the geometric projected area as well as a quadratic limb darkening law \citep[see][]{espinoza2015}.
+We treat limb darkening as a weight factor on the contribution of each point to the disk-integrated spectrum that depends on $\mu$.
+Points that are part of spots and faculae are treated the same as the photospheric points -- their respective weights are increased if they lie near the center and decreased if they lie on the limb.
 This treatment allows the user to specify two parameters $u_1$ and $u_2$ that describe the flux variation as a function of angle from disk-center.
@@
 
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
 
-The \vspec{} GCM centers around a 2D temperature map of the planet's surface. Based on \citet{cowan2011}, this map
-is determined exclusively by the incident flux from the star (parameterized by the host \teff~and radius,
-and the planet's semimajor axis and Bond albedo) and the unitless thermal redistribution efficiency $\epsilon$.
-This efficiency is 0 for a planet that re-radiates instantly and $\gg 1$ for an atmosphere that efficiently redistributes heat to its night side.
-Figure \ref{fig:thermal_inertia} shows a temperature map example for an Earth-like planet with $\epsilon = 2\pi$.
-These temperature maps are validated by the energy balance of the planet; \vspec{} computes the incident and emergent flux across the planet and issues a warning if they are not in agreement.
-
 \begin{figure}
@@
 
-To compute the surface temperature we first compute the temperature at the equator as a function of longitude. This is done by integrating Equation (10) of \citet{cowan2011}
+The \vspec{} GCM centers around a 2D temperature map of the planet's surface. Based on \citet{cowan2011}, this map
+is determined exclusively by the incident flux from the star (parameterized by the host \teff~and radius,
+and the planet's semimajor axis and Bond albedo) and the unitless thermal redistribution efficiency $\epsilon$.
+This efficiency is 0 for a planet that re-radiates instantly and $\gg 1$ for an atmosphere that efficiently redistributes heat to its night side.
+Figure \ref{fig:thermal_inertia} shows a temperature map example for an Earth-like planet with $\epsilon = 2\pi$.
+These temperature maps are validated by the energy balance of the planet; \vspec{} computes the incident and emergent flux across the planet and issues a warning if they are not in agreement.
+
+To compute the surface temperature we first compute the temperature at the equator as a function of longitude. This is done by integrating \citet{cowan2011} Equation (10).
 \begin{equation}
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/moran23.pdf}
+    \caption{
+        The results from our spotted transit experiment.
+We find that the addition of spots to a star's surface can produce spectral features in what would otherwise be a flat transit spectrum due to a lack of atmosphere or other mitigating issues such as planetary aerosols.
+These features mimic those that would be produced due to molecular absorption in a planetary atmosphere.
+This is in agreement with the results of \citet{moran2023}.
+        }
+    \script{moran23.py}
+    \label{fig:moran_transit}
+\end{figure}
+
 In this section we will simulate the transit of a atmosphere-less occulting disk across a spotted star to demonstrate how an inhomogeneous
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{figures/moran23.pdf}
-    \caption{
-        The results from our spotted transit experiment.
-We find that the addition of spots to a star's surface can produce spectral features in what would otherwise be a flat transit spectrum due to a lack of atmosphere or other mitigating issues such as planetary aerosols.
-These features mimic those that would be produced due to molecular absorption in a planetary atmosphere.
-This is in agreement with the results of \citet{moran2023}.
-        }
-    \script{moran23.py}
-    \label{fig:moran_transit}
-\end{figure}
-
 We find that, as expected, the addition of spots changes the transit depth in a way that depends on wavelength -- mimicking light that would be lost to absorbers in a planetary atmosphere.
@@
 \subsection{Phase Curve}
-\begin{figure}[t]
+\begin{figure}[h]
     \centering
@@
 The figures in this article were created using the Matplotlib \citep{hunter2007}, NumPy \citep{harris2020}, and Cartopy \citep{metoffice2010} libraries.
+The top panel of Figure \ref{fig:fac_struct} was created using the open-source vector graphics software Inkscape\footnote{\url{https://inkscape.org/}}.
 

### 2024-08-20 src/tex/ms.tex
Add a spot crossing for Veselin

@@
         {\bf Right}: Transit light curves at 1 ${\rm \mu m}$ for a Lambertian (i.e.
-no limb darkening) star and one with significant limb darkening \citep[$u_1=0.5,\,u_2=0.15$,][]{espinoza2015}.
+no limb darkening) star, one with significant limb darkening \citep[$u_1=0.5,\,u_2=0.15$,][]{espinoza2015}, and one with that same limb darkening law, but with the addition of spots.
 In each case a 1 $R_e$ planet transits a 0.12 $R_\odot$ ($T_{\rm eff} = 3000$ K) star with an observation cadence of 1 minute.
@@
 The Lambertian star (top) does not exhibit any limb darkening, and its disk has constant surface brightness.
-However, the limb-darkened star (bottom) is much brighter at disk-center than it is near the limb.
+However, the limb-darkened stars (middle and bottom) are much brighter at disk-center than near the limb.
 The dark circle on the leftmost limb of each star is the region occulted by the planet when it is $0.7^\circ$ from mid-transit.
+In the spotted star (bottom) the planet crosses over several spots as it moves from left to right.
+Note in the light curves that while no spot is covered, the transit depth is larger, and the effective radius of the planet appears higher.
+However, when a spot is crossed, the transit depth decreases dramatically.
     }

### 2024-08-20 src/scripts/ld_transit.py

@@ -13,28 +13,34 @@ import paths
 
 LAMBERT_CONFIG = paths.scripts / 'ld_lc_lamb.yaml'
 SOLAR_CONFIG = paths.scripts / 'ld_lc_solar.yaml'
+SPOT_CONFIG = paths.scripts / 'ld_lc_spot.yaml'
 OUTFILE = paths.figures / 'ld_transit.pdf'
 
 libpypsg.docker.set_url_and_run()
 
 lam_model = VSPEC.ObservationModel.from_yaml(LAMBERT_CONFIG)
 sol_model = VSPEC.ObservationModel.from_yaml(SOLAR_CONFIG)
+spt_model = VSPEC.ObservationModel.from_yaml(SPOT_CONFIG)
 
 # lam_model._build_star()
 # sol_model._build_star()
+# spt_model._build_star()
 
 lam_model.build_planet()
 lam_model.build_spectra()
 sol_model.build_planet()
 sol_model.build_spectra()
+spt_model.build_planet()
+spt_model.build_spectra()
 
 lam_data = VSPEC.PhaseAnalyzer.from_model(lam_model)
 sol_data = VSPEC.PhaseAnalyzer.from_model(sol_model)
+spt_data = VSPEC.PhaseAnalyzer.from_model(spt_model)
 
-fig = plt.figure(figsize=(8, 6))
-gs = fig.add_gridspec(5, 2)
+fig = plt.figure(figsize=(7, 6.5))
+gs = fig.add_gridspec(17, 2)
 fig.subplots_adjust(hspace=0,wspace=0.5)
-ax1: plt.Axes = fig.add_subplot(gs[1:4,1])
+ax1: plt.Axes = fig.add_subplot(gs[4:13,1])
 
 wl_index = np.argmin(np.abs(lam_data.wavelength - 1*u.um))
 
@@ -46,9 +52,12 @@ ax1.plot(sol_data.time.to_value(u.hr), sol_data.lightcurve('total',wl_index,norm
 obs = sol_data.lightcurve('total',wl_index)
 obs = obs/obs[0] + np.random.normal(scale=sol_data.lightcurve('noise',wl_index)/obs[0])
 ax1.scatter(sol_data.time.to_value(u.hr), obs, c='k', marker='.',s=1,alpha=1)
+ax1.plot(spt_data.time.to_value(u.hr), spt_data.lightcurve('total',wl_index,normalize=0), 'b', label='Spotted')
[…]

### 2024-08-20 src/scripts/ld_transit.py
clarify that the no LD case is a lambertian radiator

@@ -44,7 +44,7 @@ ax1: plt.Axes = fig.add_subplot(gs[4:13,1])
 
 wl_index = np.argmin(np.abs(lam_data.wavelength - 1*u.um))
 
-ax1.plot(lam_data.time.to_value(u.hr), lam_data.lightcurve('total',wl_index,normalize=0), 'r', label='Lambertian')
+ax1.plot(lam_data.time.to_value(u.hr), lam_data.lightcurve('total',wl_index,normalize=0), 'r', label='Lambertian rad.')
 obs = lam_data.lightcurve('total',wl_index)
 obs = obs/obs[0] + np.random.normal(scale=lam_data.lightcurve('noise',wl_index)/obs[0])
 ax1.scatter(lam_data.time.to_value(u.hr), obs, c='r',marker='.',s=1,alpha=1)

### 2024-08-20 src/scripts/ld_transit.py
Make LD light curves a more pleasant color

@@ -44,18 +44,18 @@ ax1: plt.Axes = fig.add_subplot(gs[4:13,1])
 
 wl_index = np.argmin(np.abs(lam_data.wavelength - 1*u.um))
 
-ax1.plot(lam_data.time.to_value(u.hr), lam_data.lightcurve('total',wl_index,normalize=0), 'r', label='Lambertian rad.')
+ax1.plot(lam_data.time.to_value(u.hr), lam_data.lightcurve('total',wl_index,normalize=0), '#544F43', label='Lambertian rad.')
 obs = lam_data.lightcurve('total',wl_index)
 obs = obs/obs[0] + np.random.normal(scale=lam_data.lightcurve('noise',wl_index)/obs[0])
-ax1.scatter(lam_data.time.to_value(u.hr), obs, c='r',marker='.',s=1,alpha=1)
-ax1.plot(sol_data.time.to_value(u.hr), sol_data.lightcurve('total',wl_index,normalize=0), 'k', label='$u_1=0.5,~u_2=0.15$')
+ax1.scatter(lam_data.time.to_value(u.hr), obs, c='#544F43',marker='.',s=1,alpha=1)
+ax1.plot(sol_data.time.to_value(u.hr), sol_data.lightcurve('total',wl_index,normalize=0), '#E28221', label='$u_1=0.5,~u_2=0.15$')
 obs = sol_data.lightcurve('total',wl_index)
 obs = obs/obs[0] + np.random.normal(scale=sol_data.lightcurve('noise',wl_index)/obs[0])
-ax1.scatter(sol_data.time.to_value(u.hr), obs, c='k', marker='.',s=1,alpha=1)
-ax1.plot(spt_data.time.to_value(u.hr), spt_data.lightcurve('total',wl_index,normalize=0), 'b', label='Spotted')
+ax1.scatter(sol_data.time.to_value(u.hr), obs, c='#E28221', marker='.',s=1,alpha=1)
+ax1.plot(spt_data.time.to_value(u.hr), spt_data.lightcurve('total',wl_index,normalize=0), '#54B295', label='Spotted')
 obs = spt_data.lightcurve('total',wl_index)
 obs = obs/obs[0] + np.random.normal(scale=spt_data.lightcurve('noise',wl_index)/obs[0])
-ax1.scatter(spt_data.time.to_value(u.hr), obs, c='b', marker='.',s=1,alpha=1)
[…]

### 2024-09-22 src/tex/ms.tex
Treat figure placement better
[reviewer] Add language expaining (lack of) transit fitting

@@
 
-\begin{figure*}[h]
+\begin{figure*}[!htpb]
     \centering
@@
 We use the JWST NIRSpec/G395H instrument setup from PSG with an observing time of 3.53 hours with an 8 minute cadence to match \citet{moran2023}, but set $R=200$ to account for the resolution of their reduction.
+For simplicity, we do not include limb darkening or stellar rotation.
 
@@
 these simulations because the effect was not considered in the original analysis of GJ 486b.
+This is a simplified and highly controlled experiment, so there is no need to use standard transit light curve fitting techniques.
+We are able to compute the transit depth as $(f_\text{pre}-f_\text{tr})/f_\text{pre}$, where $f_\text{pre}$ is the total spectrum before the transit, at an epoch chosen by visual inspection, and $f_\text{tr}$ is the total spectrum during the transit, again chosen by visual inspection.
+A light curve fitting code would be needed if we had introduced stellar rotation or limb darkening.
 
@@
 \subsection{Phase Curve}
-\begin{figure}[h]
+\begin{figure}[!htbp]
     \centering

### 2024-09-26 src/tex/ms.tex
[reviewer] add vspec diagram

@@
 
-The flow of \vspec{}'s code is broken into three parts: i) read in the configuration, ii) compute planetary spectra, and iii) compute stellar spectra.
-See the \vspec{} documentation page\footnote{\href{https://vspec-collab.github.io/VSPEC/0.6.0/index.html}{https://vspec-collab.github.io/VSPEC/0.6.0/index.html}} for detailed API documentation.
+\begin{figure*}
+    \includegraphics[width=\textwidth]{figures/vspec-diagram.pdf}
+    \caption{
+        Flowchart diagram outlining the typical workflow of a \vspec{} simulation.
+        {\bf Upper right}: Processes representing the main \vspec{} stages alongside Python code snippets.
+Nearly all \vspec{} use cases involve the code \texttt{model.build\_planet(); model.build\_spectra()}, though the input file and the particulars of the analysis can be very different.
+        {\bf Upper left}: \texttt{build\_planet()} algorithm.
+This portion of the code is responsible for interfacing with PSG (via \texttt{libpypsg}, see Section \ref{sec:libpypsg}) to generate a phase curve.
+The decision labeled ``Update GCM?'' evaluates to true when a time-dependent climate model is specified in the input file, or when the PSG configuration returned by the latest API call is longer than 1500 lines, indicating that PSG should be reset.
+For convenience, the PSG GlobES application appends new configuration files to those from previous API calls, but stops reading after 2000 lines for security reasons.
+Therefore, the internal configuration file must be reset on occasion.
+        {\bf Lower left}: \texttt{build\_spectra()} algorithm.
+This portion of the code is responsible for adding stellar variability into our results.
+We use the outputs from the \texttt{build\_planet()} step to create interpolator instances (implemented as \texttt{GridPolator.Gridspectra} objects, see Section \ref{sec:gridpolator}) for the time- and wavelength-dependent values of relevant quantities (e.g., thermal emission flux, transit depth, various types of noise).
+We then step through each epoch in our time series and evaluate the variable star model (using \texttt{vspec-vsm}, see Section \ref{sec:star}).
+We replace the stellar spectrum from PSG with that of our model, scaling the noise appropriately.
+Finally, we evolve the star to the time of the next observation.
+        {\bf Lower right}: Diagram key. Note that, in general, all the steps marked as \texttt{GridPolator} processes also involve \texttt{vspec-vsm}.
+    }
+    \label{fig:vspec-diagram}
+    \script{copy_static_figs.py}
+\end{figure*}
+
+The typical flow of \vspec{}'s code is shown in Figure \ref{fig:vspec-diagram}, as well as described in the following sections.
+It is generally broken into four parts: i) read in the configuration, ii) compute planetary spectra, iii) compute stellar spectra, iv) analyze, with steps (i) and (iv) varying widely by use case.
+See the \vspec{} documentation page\footnote{\urldocsvspec} for detailed API documentation.
 \subsection{Configuring \vspec{}}

### 2024-09-26 src/scripts/copy_static_figs.py

@@ -1,0 +1,14 @@
+"""
+Copy stuff from static
+"""
+import shutil
+import paths
+
+FILENAMES = [
+    'vspec-diagram.pdf'
+]
+
+for filename in FILENAMES:
+    static_path = paths.static / filename
+    target_path = paths.figures / filename
+    shutil.copy(static_path, target_path)

### 2024-09-26 src/tex/ms.tex
[reviewer] add a brief overview at beginning of section 2

@@
 It is generally broken into four parts: i) read in the configuration, ii) compute planetary spectra, iii) compute stellar spectra, iv) analyze, with steps (i) and (iv) varying widely by use case.
-See the \vspec{} documentation page\footnote{\urldocsvspec} for detailed API documentation.
+
+Most users of \vspec{} will interact exclusively with two classes: \texttt{VSPEC.ObservationModel} and \texttt{VSPEC.PhaseAnalyzer}.
+The bulk of the work is done by the \texttt{ObservationModel} class, as it reads in a user's inputs, runs the model, and writes output files.
+\vspec{} produces many output files, which themselves can be read easily using tools such as Astropy \citep{astropycollaboration2013,astropycollaboration2018,astropycollaboration2022}.
+However, the \texttt{PhaseAnalyzer} class makes it easy to read them as an ensemble, as well as quickly pull out portions of the data that are of interest.
+
+Section \ref{subsec:config} describes initializing the \texttt{ObservationModel}.
+Section \ref{subsec:phase-curve} describes computing the planetary phase curve using the \texttt{ObservationModel.build\_planet()} method.
+The next step, calling \texttt{ObservationModel.build\_spectra()}, recalls the data produced by PSG previously, but combines it with a spectral model of a variable star, scaling reflected light and noise appropriately.
+This step, along with detailed descriptions of various substeps, is described in Sections \ref{subsec:stellar-lc}, \ref{subsec:reflected}, \ref{subsec:noise}, and \ref{subsec:transit}.
+
+Use of the \texttt{PhaseAnalyzer} object is not described in detail as it mostly acts as a container, but Section \ref{sec:examples} includes many examples of its use.
+Additionally, the \vspec{} documentation page\footnote{\urldocsvspec} contains detailed API documentation.
 \subsection{Configuring \vspec{}}
@@
 However the user decides to provide model parameters, they are read into the main \vspec{} object: the \texttt{ObservationModel}.
-This is the class that runs the model and handles files. Upon initialization a local working directory
-(by convention \texttt{.vspec}) is created to store simulation outputs and intermediate files.
+Upon initialization a local working directory (by convention \texttt{.vspec/}) is created to store simulation outputs and intermediate files.
 
@@
 \label{subsec:phase-curve}
-\vspec{} generates a phase curve though a series of API calls to the Planetary Spectrum Generator \citep[PSG][]{villanueva2018}, using
-\texttt{libpypsg} (see Section \ref{sec:libpypsg}) to interface with the API. Initially, configurations are sent that give PSG information about
-the system that does not change with time -- that includes the 3D Global Circulation Model (GCM) and instrument parameters. \vspec{} then enters
-the main loop, iterating through each observation epoch (e.g.integration or combination of integrations) and making a pair of API calls: one to get a combined light spectrum and one to get the thermal flux only.
+\vspec{} generates a phase curve when the \texttt{ObservationModel.build\_planet()} method is called.
+It does this though a series of API calls to the Planetary Spectrum Generator \citep[PSG][]{villanueva2018}, using \texttt{libpypsg} (see Section \ref{sec:libpypsg}) to interface with the API.
+Initially, configurations are sent that give PSG information about the system that does not change with time -- that includes the 3D Global Circulation Model (GCM) and instrument parameters.
+\vspec{} then enters the main loop, iterating through each observation epoch (e.g.integration or combination of integrations) and making a pair of API calls: one to get a combined light spectrum and one to get the thermal flux only.
 Results from PSG in each epoch are stored locally in the \texttt{.fits} format.
@@
 \subsection{Stellar light curve}
+\label{subsec:stellar-lc}
 
-The last step is to replace the stellar spectra used by PSG with \vspec{}'s stellar model. In this step \vspec{} acts as a bridge between
+The last step is to replace the stellar spectra used by PSG with \vspec{}'s stellar model by calling the \texttt{ObservationModel.build\_spectra()} method.
+In this step \vspec{} acts as a bridge between
 the stellar surface model \texttt{vspec-vsm} (which models spatial changes to the star with time, see Section \ref{sec:star}) and a grid of
@@
 \subsection{Reflected light}
+\label{subsec:reflected}
 \vspec{} then produces a composite spectrum of the portion of the star visible to the planet to accurately incorporate the time-dependent stellar flux into the reflected-light spectrum.
@@
 \subsection{Simulated Noise}
+\label{subsec:noise}
 Alongside the planetary radiance files, PSG returns a file giving the analytic noise estimate for each epoch broken down by source.
@@
 \subsection{Transit and Eclipse}
+\label{subsec:transit}
 Users must be careful when using \vspec{} to simulate transit and eclipse measurements because of the vast timescale differences between those events and the typical phase variations of a planet -- a sparsely sampled phase curve that happens to include one epoch of transit will not adequately sample the light curve accurately during a transit or eclipse because the interpolator knows no better than to assume a linear interpolation scheme.

### 2024-09-26 src/tex/ms.tex
[reviewer] change symbol for spot lifetime

@@
 \begin{equation}
-    N(\Delta t) = \frac{4 \pi R_*^2 f_{\rm spot}}{A_{\rm mean}} \frac{\Delta t}{T}
+    N(\Delta t) = \frac{4 \pi R_*^2 f_{\rm spot}}{A_{\rm mean}} \frac{\Delta t}{\tau_\text{spot}}
 \end{equation}
-where $R_*$ is the stellar radius, $f_{\rm spot}$ is the fraction of the surface covered by spots in growth-decay equilibrium, $A_{\rm mean}$ is the lifetime-averaged mean spot area, and $T$ is the spot lifetime, defined as:
+where $R_*$ is the stellar radius, $f_{\rm spot}$ is the fraction of the surface covered by spots in growth-decay equilibrium, $A_{\rm mean}$ is the lifetime-averaged mean spot area, and $\tau_\text{spot}$ is the spot lifetime, defined as:
 \begin{equation}
-    T = \frac{A_0}{W} - \frac{1}{\tau} \ln{\left (\frac{A_{\rm min}}{A_0}\right )}
+    \tau_\text{spot} = \frac{A_0}{W} - \frac{1}{\tau} \ln{\left (\frac{A_{\rm min}}{A_0}\right )}
 \end{equation}

### 2024-09-26 src/tex/ms.tex
[reviewer] add language about discontinuity at the poles

@@
 To allow for variations in pressure and inclination, we augment their treatment with several simple approximations for latitudinal and vertical temperature variations.
+However, it should be noted that these assumptions produce a discontinuity at the poles, which is non-physical.
+A more physically-motivated model of latitudinal mixing (e.g., including a diffusion parameter in the heat transport equation) would avoid this caveat, but would also be more expensive to produce.
 

### 2024-09-26 src/tex/ms.tex
[reviewer] warn that the adiabatic atmosphere assumption violates the energy balance

@@
 \begin{equation}
+    \label{eq:tz}
     T_z = T_{\rm surf}\,\left( \frac{P_z}{P_{\rm surf}} \right)^{1-\frac{1}{\gamma}}
 \end{equation}
+
+It should be noted that varying temperature in the z direction via Equation (\ref{eq:tz}) violates one of the assumptions of the \citet{cowan2011} prescription -- that the gas parcel modeled is several scale heights thick.
+It also does not allow for heat exchange between atmosphere layers, which could have significant impacts on the atmosphere's energy balance.
 

### 2024-09-26 src/scripts/moran23.py
[reviewer] move eureka data to the back

@@ -278,7 +278,7 @@ for data,label,color in zip(
     ax.plot(data.wavelength,transit_depth*1e6,label=label,color=color)
 
 ax.errorbar(moran_x,moran_y,yerr=moran_yerr,
-    fmt='o',color='k',label='Moran+23 Eureka!',markersize=3)
+    fmt='o',color='k',label='Moran+23 Eureka!',markersize=3,zorder=-100)
 
 ax.set_xlabel('Wavelength (um)')
 ax.set_ylabel('Transit depth (ppm)')

### 2024-09-26 src/tex/ms.tex
Clarify the tz profile badness

--- 2024-09-26 src/tex/ms.tex
@@
 \begin{equation}
+    \label{eq:tz}
     T_z = T_{\rm surf}\,\left( \frac{P_z}{P_{\rm surf}} \right)^{1-\frac{1}{\gamma}}
 \end{equation}
+
+Prescribing the temperature profile via Equation (\ref{eq:tz}) is, of course, an oversimplification.
+The profile does not consider the energy transport between layers of the atmosphere, and will violate the energy balance of the planet.
+We emphasize that this model is meant to approximate an atmosphere quickly, not be a self-consistent climate model.
 
